\chapter{THE HYPOTHESIS OF ORGANIC EVOLUTION—1}

FROM whence came Man? What is his destiny? In the revelations of the Lord we are taught
that man is the offspring of God. His destiny is to gain immortality through the atonement of
Jesus Christ and having passed through death and the resurrection, to live forever. If he will
be obedient to all of the commandments of the Lord and abide in his covenants he will gain
eternal life, which is the greatest gift of God. Eternal life means to become a son of God, a
joint heir with Jesus Christ, receiving the fulness of the Father's kingdom. It is written:

For if ye live after the flesh, ye shall die: but if ye through the Spirit do mortify the deeds of
the body, ye shall live.

For as many as are led by the Spirit of God, they are the sons of God.

For ye have not received the spirit of bondage again to fear; but ye have received the Spirit of
adoption, whereby we cry, Abba, Father.

The Spirit itself beareth witness with our spirit, that we are the children of God:

And if children, then heirs; heirs of God, and joint-heirs with Christ; if so that we suffer with
him, that we may be also glorified together. 1

Behold, what manner of love the Father hath bestowed upon us, that we should be called the
sons of God: therefore the world knoweth us not, because it knew him not.

Beloved, now are we the sons of God, and it doth not yet appear what we shall be: but we
know that, when he shall appear, we shall be like him; for we shall see him as he is.

And every man that hath this hope in him purifieth himself, even as he is pure. 2

He that overcometh shall inherit all things; and and I will be his God, and he shall be my son.
3

And he [Adam] heard a voice out of heaven, saying: Thou art baptized with fire, and with the
Holy Ghost. This is the record of the Father, and the Son, from henceforth and forever;

And thou art after the order of him who was without beginning of days or end of years, from
all eternity to all eternity.

Behold, thou art one in me, a son of God; and thus may all become my sons. Amen. 4

And these are the words of Jesus after his resurrection when speaking to Mary:

Jesus saith unto her, Touch me not; for I am not yet ascended to \textit{my} Father: but go to \textit{my
brethren}, and say unto them, I ascend unto \textit{my} Father, and \textit{your} Father; and to \textit{my} God, and
\textit{your} God. 5 (italics mine.)

To all who have faith in God and in Jesus Christ, and who believe in the efficacy of the
atonement, these are glorious principles. We have reached the point, however, in the history
of the world, when many of the learned who are steeped in the philosophies of the world,
reject all of this and in their place have substituted another doctrine and way of life. In the
nineteenth century there were many good Christian people who believed in the scriptures as
far as they understood them. They accepted the atonement of Jesus Christ. Their ministers
preached that he was the Redeemer and accepted the doctrine of the resurrection of the dead.
Today these doctrines are looked upon as being antiquated, and those believing them are
looked upon by many scientists as ignorant of the beginning of things and are guilty of
accepting the false philosophies, myths and legends coming from the ancient ``Chaldeans and
Babylonians.'' 6 Then came these days of so-called scientific research to discover the origin
of man, for the Bible revelations failed to satisfy those who were no longer guided by the
Light of Christ, and the teachings became prevalent that the scriptures were filled with
``Christian errors.'' Dr. Andrew D. White in his relentless determination to destroy man's faith
in the Bible stories of creation and all things miraculous, declared that ``Biblical theology
continued to spin its own webs out of its own bowels, and all the lesser theological flies
continued to be entangled in them; yet here and there stronger thinkers broke loose from this
entanglement and helped somewhat to disentangle others.'' 7

So now, in the twentieth century, the doctrines of the critics of the Bible and the teachings of
the organic evolutionists, have gained the ascendency in the scientific world. It is true that in
former years we lived in a Christian nation, the fact persists that now many Christian
ministers, so-called, have been caught in the web of modernism and organic evolution and
have rejected the fundamental doctrines of Christianity; and they, like the Christians in the
days of Rome, have mingled their religious views with these modern (pagan) teachings.
Because of the influence of destructive criticism and these theories of the descent of man,
many ministers have rejected the fall of Adam, the atonement of Jesus Christ, and the
resurrection of the dead. In fact they have come to the point where they have discarded the
doctrine of the resurrection of Jesus Christ, and that he is the Only Begotten Son of God.
Their Christianity, filled with abundant errors before, has sunk to a lower level. These
advocates of modernism and evolutionary teachings, glory in the fact that their influence has
helped to eliminate from Christianity, the ``dogma of Adam's fall,'' 8 and the ``legendary
husks and rinds of our sacred books.'' 9 One day, when they come to the judgment, they will
have to give an accounting for all this mischief they have done. It may be imagined how they
will feel, when they are forced to confront the thousands who have been turned away from
faith in God and acceptance of his divine plan of salvation, because these enemies of truth
were eager to destroy the scriptures and the mission of Jesus Christ. If great joy will be felt
by the individual who has, through his humble effort, saved one soul, then how great must be
the remorse of these learned men when they discover that their efforts have been the means
of destroying thousands of souls? 10

This brings us to the discussion of what I believe to be the most pernicious doctrine ever
entering the mind of man: the theory that man evolved from the lower forms of life. For its
source we must go beyond the activities and research of mortal man to the author of evil,
who has been an enemy of truth from the beginning before the earth was formed. This
doctrine is not new in this modern world; it is merely its pernicious application that has been
developed during this and the preceding century. It was natural for people in former times to
think that life developed spontaneously. Did they not see what appeared to them countless
forms of life spontaneously appearing? Such life appeared in cheese, in stagnant pools,
decayed matter or most anything that stood long enough in the warmth of the sun. Even men
who professed Christian convictions accepted this to be the fact. Crocodiles were generated
from the slime of the Nile, bees and flies from decomposed flesh, and some writers
endeavored to draw a line between animals which were generated in carion and those coming
from earth and water. In fact there was scarcely a limit in the minds of many as to the extent
of spontaneous generation. This belief continued down to the middle of the nineteenth
century especially in relation to insects and bacteria and lower forms of life, and it was the
scientists who followed the ideas of Darwin, Wallace and others, who were the keenest in the
search to discover if this apparent notion were true. When careful research was made the
whole theory exploded; and it was these scientists who were forced to admit that such a thing
as spontaneous generation is not true. Such men as Huxley, Tyndall, Spencer, Pasteur,
Haeckel and many others experimented in this field. It was Pasteur who showed that this
fallacy could not be maintained. They had to admit that every living thing from germ to
elephant comes from antecedent life. Writing on this subject Dr. C. W. Saleeby has said:

Now, the remarkable fact—one of the most striking in the history of science is that the time
honored belief in spontaneous generation should have been attacked, and attacked with
apparent success, just at the very time when it would otherwise have begun to assume real
philosophical importance. For ages it had been accepted, taken as a matter of course, and not
regarded as having any particular bearing upon the supreme question. Then there came the
time when this belief would have been an all-important link, without which the chain of
evolution could not be completed, a link without which we were left to contemplate a perfect
chain of inorganic evolution—the history of life upon the earth, with an abyss between the
two that could not be bridged, for how came life where was no life? A series of experiments
were made—experiments in which, strikingly enough, some of the greatest evolutionists of
the day took a leading part, and these seemed to upset, just when it was most wanted by
themselves for the establishment of their new doctrine, the belief which had gone without
question for so many ages.

Now, some may be inclined to wonder how it should be that certain pioneers of the new
doctrine of evolution, such as Tyndall and Huxley, should devote themselves with such
persistence and labor and force to the overthrow of a doctrine which was so necessary for the
complete establishment of their own case—so much so, that when they had overthrown it,
they found themselves, as regards their own doctrine of evolution, placed in a difficulty from
which they did not live to emerge. . . .

It is well worth noting that the common doctrine of spontaneous generation was always held
in reference to organic materials, such as the slime of the Nile—not the dry sand of the
desert. The reader may be inclined to say that men's belief on this subject in the past
generation makes very confused reading, and, indeed, that is true. But the fact is that their
beliefs were most confused. The work of Darwin had staggered everybody, and
straightforward, systematic, unprejudiced thinking was very nearly impossible in the welter
of controversy. Nevertheless, something apparently definite was done. The doctrine of the
beginning of life upon the earth was almost undiscussed, and the accepted notion of the
nature of matter—a notion which to us who know radium, seems purile—was left
unchallenged in all its falsity. But the work of the great French chemist Pasteur led to a close
examination of the belief that humble forms of life are daily produced from lifeless organic
materials, and the conclusion was reached that no such spontaneous generation occurs.
This conclusion is of great importance in the history of modern thought, and it was
proclaimed with much rejoicing and vigor as a great achievement of science, whilst some of
the chief advocates seemed at times to forget the extreme awkwardness of the inferences
which had to be made from it. The doctrine may be stated in Latin in the form of the familiar
dogma, \textit{Omne vivum ex. vivo:} Every living thing from a living thing. . . . For every creature,
microbe, or mammoth or man, we must trace back in imagination a series of living ancestors,
different perhaps in various characters, but always living. This series must be traced back and
back and back until . . .?

And there the difficulty arose. 11

Notwithstanding the great discovery of Pasteur, Darwin and his followers were not retarded
in their search to find the beginning of life and to prove that all things have developed from
spontaneous life. This question has never been answered successfully other than the account
in the scriptures: If spontaneous generation cannot be created now, how could it be possible
several million or billion years ago? Conditions, according to the teachings of science, are
more favorable now than they possibly could have been in the far distant past. To get a
beginning these advocates must \textit{assume} some starting point, notwithstanding there is no
evidence that will support it. All evidence points to the contrary. Now a word in relation to
this assumed starting point.

According to one scientist:

The doctrine explains that the animal kingdom, past and present, is comparable to a highly
developed tree. The trunk is the original amoeba or something similar to it. 12

According to Charles Darwin:

The most ancient progenitors in the kingdom of Vertebrata, at which we are able to obtain an
\textit{obscure glance, apparently} (My italics) consisted of a group of marine animals, resembling
the \textit{larvae} of existing Ascidians. These animals \textit{probably} (My italics) gave rise to a group of
fishes, as lowly organized as the lancelet; and from these the Ganoids, and other fishes like
the Lepidosiren, must have been developed. From such fish a very small advance would
carry us on to the Amphibians.

From this beginning Mr. Darwin argues that while it is a very long step, it is ``a pedigree of
prodigious length, but not, it may be said, of noble quality.'' 13

It is hardly necessary to go into detail about the evolutionary theory of descent, more than to
say, it teaches that from this small beginning has sprung all living creatures on the face of the
earth. We will let one of the advocates of this theory state the case:

Organic evolution is the branch of evolution that deals with the development of plant and
animal life. According to its teachings all plants and animals have come up through a long
line of successive changes from a common ancestry. 14

It makes little difference whether the evolutionist decides on the amoeba or on the larvae of
an ascidian for his beginning, the insuperable difficulties are just the same. It requires a vast
stretch of the imagination, far beyond the realms of reason, for one to accept such a theory as
this. Let us suppose for the sake of argument, that the first speck of life was an amoeba. We
can suppose—for that is in keeping with the entire doctrine of organic evolution, for its entire
structure is based on supposition, and cannot be based on anything else—so, we will
suppose, that back several millions or billions of years—no one was there to watch the
process by which this speck of life came spontaneously into existence—the amoeba suddenly
appeared and multiplied, as the amoeba will do, and after millions of years, it, or one of its
descendants began to develop fins, then a head and then a tail and after several more millions
of years it became a fish, or a tadpole, or a brachiopod, or a trilobite, or a snail, even a
worm—it makes no difference which, one guess is as good as another—and becoming tired
of the water it came out upon the land, leaving its companions to develop into acquatic
animals, while it dug itself in the soil and became a plant, a fern, a rose bush or a tree. Then
another, discovering that the land was pleasant, also came forth from the water and became a
frog, a toad, a lizard or a snake and in course of time its descendant became a tiger, a lion, a
bear, an elephant, dinosaur or a little timid lamb; perhaps it took to the air as a dove, a robin,
a hawk or an eagle. Why go any farther? Does it not all sound extremely ridiculous? Well, so
it is! Yet it is this kind of rubbish that is put forth apparently in all seriousness. Books are
written about it; lectures are given in class rooms, from pulpits and platforms, and thousands
of well meaning people say they believe it! Then again the question arises: Each of these
animals had to have a companion, and we find ourselves in a quandary to discover just why
and how both males and females came into existence, both in the animal world and among
trees and other vegetation. So we find ourselves floundering in the depths of an unfathomable
hypothesis about which no one has ever been able to do more than to make an uncertain
guess. Others of this amoeba's descendants became a bee, a wasp or a grasshopper, a gnat or
a fly. Among these wonderful mutations there also came forth a monkey, then a baboon, a
gorilla and then man!

My dear friends, cannot you see how utterly foolish it all is? Why is it that thousands of
intelligent looking human beings are willing to accept these stupid teachings? Frankly it is
because Satan has deceived them and they love darkness rather than the light. Surely the day
has come prophesied by Paul and written in his second Epistle to the Thessalonians:

And then shall that Wicked be revealed, whom the Lord shall consume with the spirit of his
mouth, and shall destroy with the brightness of his coming:

Even him, whose coming is after the working of Satan with all power and signs and lying
wonders,

And with all deceivableness of unrighteousness in them that perish; because they receive not
the love of the truth, that they might be saved.

\textit{And for this cause God shall send them strong delusion, that they should believe a lie:}

That they all might be damned who believe not the truth, but had pleasure in
unrighteousness. Ch. 2:8-12. (My italics).

Permit me to quote a few paragraphs from the writings of Hon. William Jennings Bryan:

Having given Darwin's conclusions as to man's ancestry, I shall quote him to prove that his
hypothesis is not only groundless, but absurd and harmful to society. It is groundless because
there is not a single fact in the universe that can be cited to prove that man is descended from
the lower animals. Darwin does not use facts; he uses conclusions drawn from similarities.
He builds upon assumptions, probabilities and inferences, and asks the acceptance of his
hypothesis notwithstanding the fact that connecting links have not hitherto been discovered.
He advances an hypothesis which, if true, would find support on every foot of the earth's
surface, but which, as a matter of fact, finds support nowhere. There are myriads of living
creatures about us, from insects too small to be seen with the naked eye to the largest
mammals, and yet, not one is in transition from one species to another; every one is perfect.
It is strange that slight similarities could make him ignore gigantic differences. The remains
of nearly one hundred species of vertebrate life have been found in the rocks, of which more
than one-half are found living today, and none of the survivors show material change. The
word hypothesis is a synonym used by scientists for the word guess; it is more dignified in
sound and more imposing to the sight, but it has the same meaning as the old-fashioned every
day word, guess. If Darwin had described his doctrine as a guess instead of calling it an
hypothesis, it would not have lived a year. Probably nothing impresses Darwin more than the
fact that at an early stage the fetus of a child cannot be distinguished from the fetus of an ape,
but why should such a similarity in the beginning impress him more than the difference at
birth and the immeasurable gulf between the two at forty? If science cannot detect a
difference, \textit{known to exist}, between the fetus of an ape and the fetus of a child, it should not
ask us to substitute the inferences, the presumptions and the probabilities of science for the
word of God.

Science has rendered invaluable service to society; her achievements are innumerable—and
the hypothesis of scientists should be carefully examined and their arguments fairly weighed,
but the scientist cannot compel acceptance of any argument he advances, except as, judged
upon its merits, it is convincing. Man is infinitely more than science; science, as well as the
Sabbath, was made for man. It must be remembered, also, that not all sciences are of equal
importance. . . . 15

And yet I read only a few weeks ago, on page 124, of a little book recently issued by a
prominent New York minister, the following:

``Man has grown up in this universe gradually developing his powers and functions as
responses to his environment. If he has eyes, so the biologists assure us, it is because light
waves played upon the skin and eyes came out in answer; if he has ears it is because the \textit{air
waves} were there first and the ears came out to hear. Man never yet, \textit{according to the
evolutionist}, has developed any power save as a reality called it into being. There would be
no fins if there were no water, no wings, if there were no air, no legs if there were no land.''

You see I only called your attention to forty percent of the absurdities; he speaks of eyes,
ears, fins, wings, and legs—five. I only called attention to eyes and legs—two. The
evolutionist guesses himself away from God, but he only makes matters worse. How long did
the ``light waves'' have to play on the skin before the eyes came out? The evolutionist is very
deliberate; he is long on time. He would certainly give the eye thousands of years, if not
millions, in which to develop; but how could he be sure that the light waves played all the
time in one place or played in the same place generation after generation until the
development was complete? And why did the light waves quit playing when two eyes were
perfected? Why did they not keep on playing until there were eyes all over the body? Why do
they not play today, so that we may see eyes in process of development? And if the light
waves created the eyes, why did they not create them strong enough to bear the light? Why
did the light waves make eyes and then make eye-lids to keep the light out of the eyes?

And so with ears. They must have gone \textit{in} ``to hear,'' instead of out, and wasn't it lucky that
they happened to go in on opposite sides of the head instead of cate-cornered or at random?
Is it not easier to believe in a God who can make the eye, the ear, the fin, the wing, and the
leg, as well as the light, the sound, the air, the water and the land?

There is such an abundance of ludicrous material that it is hard to resist the temptation to
continue illustrations indefinitely, but a few more will be sufficient in order that you may be
prepared to ridicule these pseudo-scientists who come to you with guesses instead of facts,
let me give you three recent bits of evolutionary lore.

Last November I was passing through Philadelphia and read in the afternoon paper a report
of an address delivered in that city by a college professor employed in extension work. Here
is an extract from the paper's account of the speech: ``Evidence that early man climbed trees
with their feet lies in the way we wear the heels of our shoes—more at the outside. A baby
can wiggle its big toe without wiggling its other toes—an indication that it once used its big
toes in climbing trees.'' What a consolation it must be to mothers to know that the baby is not
to be blamed for wiggling the big toe without wiggling the other toes. It cannot help it, poor
little thing; it is an inheritance from ``the tree man,'' so the evolutionist tells us.

And here is another: ``We often dream of falling. Those who fell out of the trees some fifty
thousand years ago and were killed, of course, had no descendants. So those who fell and
were \textit{not} hurt, of course, lived, and so we are never hurt in our dreams of fallings.'' Of course,
if we were actually descended from the inhabitants of trees, it would seem quite likely that
we descended from those that were \textit{not} killed in falling. But they must have been badly
frightened if the impression made upon their feeble minds could have lasted for fifty
thousand years and still be vivid enough to scare us.

If the Bible said anything so idiotic as these guessers put forth in the name of science,
scientists would have a great time ridiculing the sacred pages, but men who scoff at the
recorded interpretation of dreams by Joseph and Daniel seem to be able to swallow the
amusing interpretations offered by the Pennsylvania professor.

A few months ago the Sunday School Times quoted a professor in an Illinois University as
saying that the great day in history was the day when a water puppy crawled up on the land
and, deciding to be a land animal, became man's progenitor. If these scientific speculators
can agree upon the day they will probably insist on our abandoning Washington's birthday,
the Fourth of July, and even Christmas, in order to join with the whole world in celebrating
``Water Puppy Day.'' 16

We have all read such silly things in supposedly scientific articles. I remember one such
article in a book written by an advocate of organic evolution who attached letters as symbols
following his name indicating that he held scholastic degrees. He declared that the reason a
man when drowning will clutch at a straw was because his ape ancestors swung from limb to
limb in trees and for the same reason boys delighted to swing on trapeze and horizontal bars
and climb trees; and that the reason why babies crawled before they learned to walk is
because in the far distant past their ancestors went on all fours. Surely an hypothesis is hard
put to it to have to resort to such foolish things as evidence. However, Mr. Charles Darwin,
the king-bolt of the entire group, was guilty of saying things just as foolish as any of these.
Here are a few examples. I quote from Darwin's \textit{Descent of Man}, published by D. Appleton,
edition of 1897:

Rudiments of various muscles have been observed in many parts of the human body; and not
a few muscles, which are regularly present in some of the lower animals can occasionally be
detected in man in a greatly reduced condition. Every one must have noticed the power
which many animals, especially horses, possess of moving and twitching their skin; and this
is effected by the \textit{panniculus carnosus}. Remnants of this muscle in an efficient state are
found in various parts of our bodies; for instance the muscle of the forehead, by which the
eyebrows are raised. The \textit{platysma myoides}, which are well developed on the neck, belong to
this system. (Pages 12-13.)

According to this argument, all of us who are able to wrinkle our foreheads and raise our
eyebrows give evidence that we are related to the horse. A wonderful stretch of the
imagination is it not? Here is another from page 19:

I am informed by Sir James Paget that often several members of a family have a few hairs in
their eyebrows much longer than the others; so that even this slight peculiarity seems to be
inherited. These hairs, too, seem to have their representatives; for in the chimpanzee, and in
certain species of Macacus, there are scattered hairs of considerable length rising from the
naked skin above the eyes, and corresponding to our eyebrows; similar long hairs project
from the hairy covering of the superciliary ridges in some baboons. (Page 19.)

Here again we have the argument that all of us—and as President Heber J. Grant at times said
facetiously ``of whom I am which''—who have a few hairs in our eyebrows that refuse to
remain uniformly of the same length, give evidence that we have descended from the
chimpanzee or baboon. Let us carry it further. Because we have a nose, eyes and a mouth, an
ear and a tongue, therefore we are related to every other creature that possesses these same
``rudiments'' and therefore we have the same common ancestor. Wonderful reasoning!

Again:

With respect to development, we can clearly understand, on the principle of variations
supervening at a rather late embryonic period, and being inherited at a corresponding period,
how it is that the embryos of wonderfully different forms should still retain, more or less
perfectly, the structure of their common progenitor. No other explanation has ever been given
of the marvelous fact that the embryos of a man, dog, seal, bat, reptile, etc., can at first hardly
be distinguished from each other. In order to understand the existence of rudimentary organs,
we have only to suppose—

(And may I add that throughout this work and in all of this author's works, there is nothing
but speculation and supposition. Not one single fact linking man to any other form of life.)

that a former progenitor possessed the parts in question in a perfect state, and under changed
habits of life they became greatly reduced, either from simple disuse, or through the natural
selection of those individuals which were least encumbered with a superfluous part, aided by
the other means previously indicated. (Pages 24-25.)

And so we find Mr. Darwin's works filled with such comparisons which get us nowhere for
there are no means in existence by which this descent can be shown or proved. No one denies
that there are comparisons, conditions that are similar to be found in living beings. The fish,
the fowl, the beast and man, all have blood, the life-giving medium by which they exist. They
have internal organs, muscles and bones; they eat, they smell, they see and they hear. The
Almighty who created them built them, naturally, in many parts on a similar plan. Why
should it be otherwise? But there is nothing in any of these things that in the slightest
indicates or proves a common ancestry. It is quite evident that these advocates are hard put to
it, when they have to resort to the embryos of ``man, dog, seal, bat, reptile, etc.,'' but as far as
history of man goes back, the human offspring has always come into the world, as has every
other creature, \textit{``after his kind.''} Mr. Darwin states that ``Man is developed from an ovule,
about 125th of an inch in diameter, which differs in no respect from the ovules of other
animals.'' ``In its early stages it can hardly be distinguished from other animals,'' of the
vertebrate kingdom. He assumes that \textit{all} vertebrates constitute and belong to one kingdom.
So this confusing story goes on. Following the lead of Darwin and a few earlier advocates of
this theory, thousands of text-books have been written, steeped in the essence of organic
evolution.

The advocates of this doctrine have sent forth expeditions to all parts of the earth seeking
fossils from the rocks and bowels of the earth. More particularly, in later years, these
researches have gone forth with the hopeful desire of finding in the fossils evidences of
creatures formerly existing in all the stages from the simple vertebrates up through
continuous processions till they could come to \textit{homo sapiens}, or modern man. These
researches carried on with the greatest zeal, have not resulted in any discoveries to prove the
theory of evolution. These disappointments have forced the advocates to resort to
deception—deception of the most reprehensible character. In the American Museum of
Natural History, New York City, in the ``Hall of the Age of Man'' are found a number of
busts prepared under the direction of Professor Henry Fairfield Osborn which are supposed
to be ``restorations'' showing the descent of man through ape ancestors. The professor who
``restored'' these ``missing links,'' J. H. McGregor, certainly had a vivid imagination. Pictures
of these busts of the ``Trinil Ape Man,'' the ``Neanderthal Man'' and the ``Cro-Magnon Man,''
are to be found in guide series No. 52. Professor Osborne describes these three as follows:
``These three restorations of pre-historic man form a progressive series, from left to right, is
evident not only by the general form and the appearance of relative intelligence appreciated
by the most casual observer, but specially by definite anatomical characters such as increase
prominence of the chin, reduction of the eyebrow ridges, reduction of the prominence of the
lower face as a whole, increased size of skull and of brain capacity (brain capacity of the
three races from left to right: 858-900cc.; 1408cc.; 1550-1800cc.)''

The history in relation to the creation of these three busts is rather interesting and should be
generally known. The story of the first, the substance out of which the bust of the Trinil Ape-
Man was constructed consists of a small section of a brain pan, two molar teeth and a piece
of thigh bone gathered by a Dutch military surgeon named Eugene Dubois. He was a friend
of Ernest Haeckel, notorious for his deception practices, not only on the public, but also on
his scientific brethren. Mr. Dubois found a small piece of the upper part of a skull and a
molar tooth on the left bank of the Solo River near Trinil, in central Java. One year later, a
left femur bone was found about 45 or 50 feet from the place where the skull bone was found
the year before. At another point a second molar tooth was found and later a third tooth.
Previously other bones of animals had been excavated in this place and more subsequently.
The evidence that the cap of the skull and the femur and the teeth found at different places,
were all parts of the same human skeletal frame, requires a great stretch of the imagination.

Nevertheless this great stretch was made and from these a ``reconstructed'' man, known in
some scientific circles as Pithecanthropus Erectus. The first of these bones were found in the
year 1891. In the eagerness of these researches to prove the descent of man from the apes, it
became a simple matter to ``reconstruct'' such a creature as an ape-man out of the meagre
bones notwithstanding the distance which separated them when they were discovered. These
scientists have constructed a skull representing a very low order and have estimated the brain
capacity as being only 858-900cc. And from this meagre showing they have developed an
entire race.

Professor George McCready Price speaking of this imaginary being says in his \textit{The New
Geology}:

The average human brain has a capacity of about 49 ounces, and no normal brain is less than
30 ounces. The Java skull has been variously pronounced as that of an idiot, that of a sub-
man or an ape-man, or that of a large Hylobates, or gibbon.

When these remains were first brought to Europe, they were shown before the third
International Congress of Zoologists in Leiden, in September 1895. Dr. Rudolf Virchow, for
thirty years president of the Anthropological Society of Berlin, and by odds the most
prominent archaeologist and pathologist of his time, was president of this congress at Leiden;
and at the close of the remarks of Dr. Dubois, he criticized the latter's report by saying there
was no certainty that all these bones were really parts of the same individual. He further
declared that it could not be known positively whether they were the remains of a man or an
ape. Later, after he had had an opportunity of examining the remains more closely, he
expressed his decided opinion that the skull was that of a large gibbon. The teeth, he said,
were more ape-like than human; while the thigh bone, though much like that of a man, would
also pass for that of a gibbon, as the latter is the only one of the apes which habitually walks
in an upright position, and there is a very marked resemblance between the femur and that of
man.

This opinion that these bones really represent a gibbon is concurred in by some of the
foremost scientists. For example, Richard Hertwig says: ``The opinion that is most probably
correct is that the fragments belonged to an anthropomorphic ape of extraordinary size and an
abnormal cranial capacity, and with a relatively large brain.''

Only the large size seems to stand in the way of anyone's regarding it as the remains of an
ape. Macnamara, after careful study of the specimen, declared that Pithecanthropus was a
true ape of rather extraordinary size. He says, ``The cranium of the average adult male
chimpanzee and the Java cranium are so closely related that I believe them to belong to the
same family of animals, that is, to the true apes.'' 17

Dr. W. H. Ballou in an article in the North American Review (April 1922) expresses doubt
upon the \textit{bona fides} of the discoverer of these Java remains because of the fact that Dr.
Dubois sealed up these remains shortly following the discovery, with other ``finds'' and has
never allowed any scientist to examine them. 18

Alfred Watterson McCann says:

Professor Osborn's own witnesses, Klaatsch, Schwalbe, and Alsberg declare that the Trinil
Ape-Man not only does not belong to the pedigree of man but that it does belong to the
pedigree not of any extinct ape or fossil ape, but to the pedigree of the modern apes,
wherefore, ``He ceases to be a witness in support of the theory of man's descent from beasts.''
Professor Osborne admits that no living ape belongs to the pedigree of man. He also admits
that no fossil ape belongs to the pedigree of man. His witnesses Schwalbe and Klaatsch admit
that the Trinil monster does not belong to the pedigree of the modern apes, but as modern
man and modern apes are admitted to have no relation to each other, Professor Osborne can't
bring in the Trinil monster without bringing in the modern apes, and precisely that he
confesses he cannot do. 19

The ``Neanderthal Man,'' second in the list, according to the manufactured busts, shows a
little more resemblance to modern man. The story of this wonderful ``discovery'' is just as
interesting, and just as misleading as Pithecanthropus, or the Trinil Ape-Man. In August
1856, two laborers who were digging in a small cave at the entrance of the Neanderthal
gorge, near Elberfield, Germany, discovered a piece of skull. Dr. Fuhlrott, a physician of
Elberfield, became interested in this find and under his direction other fragments of bone
were found in the same cave. He collected a human thigh bone, well preserved, several
human arm bones not well preserved, a piece of fore-arm bone, a human right shoulder blade,
a part of a collar bone and five broken pieces of ribs. The following year Professor D.
Schaffhausen of Bonn made a preliminary report on these bones at the meeting of the Lower
Rhine Medical and Natural History Society of Bonn. Dr. Fuhlrott also made a report in June
1857, on the same bone fragments at the general meeting of the Natural History Society of
Prussian Rhineland and Westphalia. Dr. Fuhlroot stated that these bones might be ``fossils.''
Dr. Ales Hrdlicka said that by ``fossils'' Dr. Fuhlroot evidently ``meant belonging to a form of
humanity no more existing.'' Professor Huxley considered them the ``most apelike of any
skull discovered up to that time.'' He compared them to the flat-headed natives of Australia
and thought they were of great antiquity. There was no evidence of human implements found
with these bones. ``From all that could be learned of the circumstances attending their
finding, it was entirely uncertain whether the bones had been buried in this cave by friends in
prehistoric times, or whether they had been washed into this place from the surface of the
ground outside. As for the geological 'age' represented, 'one man's guess is as good as
another's.'

``Regarding the type of skull here being dealt with, Virchow unhesitatingly pronounced it a
pathologic specimen; and we must remember that Virchow was the founder of the science of
pathology. He further declared that he had often seen upon the streets men with just such
shaped skulls as that of the Neanderthal Man. Others, however, have strongly emphasized the
prominent ridges above the eyes, and the very long, almost flat form of the head.'' 20

Alfred Watterson McCann speaking of the receding forehead of the Neanderthal skull says:

All lovers of the theatre who admired the intellectual achievements of Sir Henry Irving
marveled over his very low forehead which sloped markedly backward, though not so much
so as the forehead of Marquis Lafayette of revolutionary fame. The writer possesses an
autographed photograph given to him in 1900 by Sir Henry himself, and is not surprised that
the materialistic evolutionist completely ignore its profile in their comparative studies. The
Henry Irving skull in some respects would confound them. One could distort its description
in support of any weird theory under discussion. 21

He further says:

Describing the original Neanderthal skull-cap, Dr. Hrdlicka says, page 30, ``The Most
Ancient Skeletal Remains of Man,''—``The forehead is very low and also slopes markedly
backward, nevertheless it presents a moderately defined convexity. The sagittal region is oval
from side to side, much like that in man of today.'' The description could be forced to fit Sir
Henry's brain pan with respect to the outlines of the sagittal median curve. It would fit Sir
Lafayette's brain pan nicely. And Rudyard Kipling's!

The internal capacity of this skull was fixed very low, for obvious reasons, by Professor
Schaffhausen. He wanted to get it, like the Piltdown skull of a later date, as close to the brute
as possible, so he declared that it had a cc. capacity of 1033. The highest form of ape stops at
600 cc.

Even Professor Huxley was forced to correct this estimate by giving it a cc. capacity of 1230.
Professor Schwalbe confirmed the Huxley measurements by giving it a cc. capacity of 1234,
which is very close to the cranial capacity of the modern school teacher's skull. 22

The imaginary Cro-Magnon man is not a bad looking fellow. Of course he could not be
because in his day, according to the theory, man had advanced to a respectable stage of
intelligence with a brain capacity of 1550-1880 cc. He had learned to make fire, to
manufacture crude implements of various kinds and weapons with which to defend himself
and family. Three skeletons were found at Cro-Magnon, Dordogne, France, and were well
developed and rather splendid specimens. Many noted scientists have declared that they are
not very old. ``The old man of Cro-Magnon was over six feet tall with a skull which
authorities say was equal to that of Bismarck.'' 23 Sir Arthur Keith declared that this race of
men ``was the finest the world has ever seen.'' 24 Macnamara added his testimony by saying
that the tradition about ``a race of giants in far distant times was no myth.'' 25 However, the
skeletal remains of this man, answers well the imaginary development of man from lower
forms of life.

Perhaps a few words should be said about the ``Piltdown Man.'' About the year 1910, Mr.
Charles Dawson walking along a country road close to Piltdown Common, in Sussex,
England, was attracted by the material with which the road bed had been repaired, he found
two men digging gravel and asked them if they had found any fossils. They answered no, but
at his request they said they would look for some. About one year later he again visited this
place and was handed a small portion of a skull showing the brain case to be rather thick, but
with a forehead as high as in ``modern'' man and not receding although somewhat narrow.
The brain capacity was estimated to be about 43 ounces or almost equal to the average
female brain which is stated to be about 44 ounces. In another visit some months later, he
was rewarded by another piece of skull. A systematic search of the gravel bed was made and
resulted in the finding of a canine tooth, a piece of jaw bone, and with these fragments found
in different parts of the gravel bed many feet apart, the entire find being, as Alfred W.
McCann has said, ``with these fragments, which a juggler could conceal in the palm of one
hand, the scientists 'reconstructed' the Piltdown Man, and at once proclaimed it to be a new
genus which they proceeded to call Eoanthropus, or 'Dawn Man,' naming the species
'Dawsoni' in honor of the discoverer.'' He further states: ``The original reconstructors not only
wanted a near-ape skull which as now, alas, vanished in their hands, but they also wanted an
ape-like face and jaw. So they put their solitary canine tooth on the right side of the lower
jaw at an angle suggestive of the ape. This also suited the requirements exactly. 26

Later investigations revealed that the tooth was not a lower tooth, but an upper tooth coming
from the left side of some jaw. Further investigation convinced some of the scientists that
these bones did not belong to the same individual, and in \textit{Nature}, (November 13, 1913, p.
319) Professor Waterson published a scientific paper in which he said:

To refer the jaw bone and the cranium of the Piltdown remains to the same individual would
be exactly equivalent to articulating a chimpanzee foot with the bones of a human thigh and
leg.

The outlines of the Piltdown jaw are identical with those of a chimpanzee jaw. The molar
teeth (of the jaw) are identical with the ape form. The cranial fragments on the other hand are
in practically all their details essentially human.

Professor D. W. Matthew of the American Museum of Natural History, January 21, 1916,
published a paper in \textit{Science}, in which he declared that Professor Gerrit S. Miller had made a
report stating that the absolute identity of the Piltdown jaw as that of a chimpanzee was
``convincing and irrefutable.'' George Grant Macurdy of Yale University in an article in
\textit{Science}, February 18, 1916, p. 228-231, made the following comment in relation to this
hoax:

Regarding the Piltdown specimens we have at last reached a position that is untenable. The
cranium is human, as was recognized by all in the beginning. On the other hand, the
mandible and the canine tooth are those of a fossil chimpanzee. This means that in place of
Eoanthropus Dawsoni we have two individuals belonging to different genera.

In \textit{Current History} (October 1927) an article appeared under the caption: \textit{Darwinism
Reaffirmed By Latest Evidence}. This was written by Watson Davis:

Science is at its best when the British Association for the Advancement of Science meets.
Not alone England but all the world attends and speaks. To its sessions many important
announcements have been made. Its program is a barometer of scientific progress. It is,
therefore, interesting that Darwin's theory of man's descent was chosen by this year's
president, Sir Arthur Keith, famous anatomist, for his principal address at the Leeds meeting.
(The full text is printed at the end of this article.)

In 1860, at an Oxford meeting of the British Association, the great Huxley had his
spectacular fight with the Bishop of Oxford and emerged victorious. But in that day it was
the usual conviction that man had appeared on earth by a special act of creation. Today the
situation is changed. There are no fights upon evolution at scientific meetings. Occasionally
far from the laboratory and the study, on the frontiers of the spread of knowledge, there are
those who wish to suppress the truth by vain laws. Even the spectacle of Dayton has had no
encore. Those who listened to Sir Arthur Keith's summary of the recent discoveries which
tend to support Darwin's ideas did not have to be convinced; the theories which they already
held were only strengthened, not altered.

The ``recent discoveries'' referred to by Mr. Davis have reference to the ``finding'' of the
notorious Piltdown Man. In defense of his position Sir Arthur Keith had this to say:

Our searches have shown us that man's evolution has not proceeded in this orderly manner,
[That is, through a single race, but through ``broken up races, and numerous separate
species.''] In some extinct races, while one part of the body has moved forward another part
has lagged behind. Let me illustrate this point, because it is important. We now know that, as
Darwin sat in his study at Down, there lay hidden at Piltdown, in Sussex, not thirty miles
distant from him, sealed up in a bed of gravel, a fossil human skull, and jaw. In 1912, thirty
years after Darwin's death, Charles Dawson discovered this skull and my friend, Sir Arthur
Smith Woodward, described it and rightly recognized that skull and jaw were parts of the
same individual, and that this individual had lived, as was determined by geological and
other evidence, in the opening phase of the Pleistocene period.

We may confidently presume that this individual was representative of the people who
inhabited England at this remote date. The skull, although deeply mineralized and thick-
walled, might well have been the rude forerunner of a modern skull, but the lower jaw was so
ape-like that some experts denied that it went with the human fossil skull at all and supposed
it to be the lower jaw of some extinct kind of chimpanzee. This mistake would never have
been made if those concerned had studied the comparative anatomy of anthropoid apes. Such
a study would have prepared them to meet with the discordances of evolution. (\textit{Current
History}, Oct. 1927, pp. 100-101.)

This ``latest evidence'' was the ``Piltdown Man'' which Sir Arthur Keith presents as an
infallable evidence of man's descent from apes. He and Sir Arthur Smith Woodward,
according to Sir Arthur's statement, declare that the jaw definitely belong to the skull and this
therefore affirms the claims of Mr. Darwin, without any doubt. It is too bad that Sir Arthur
Keith could not see the discoveries recently made throwing further light on the [in]famous
Piltdown Man. In the press dispatches of November 21, 1953, came the following copied
from the \textit{Salt Lake Tribune} of the morning of November 22nd:

``LONDON'S FACE RED ON 'OLDEST SKULL'.''

``LONDON, Nov. 1—A keeper in the British Museum of Natural History Saturday shame-
facedly took from its case of honor the supposed skull of a primitive man which for 40 years
had been called the oldest ever found in Europe.

``A card in the case explained that this plaster cast assembly represented the remains, locked
in a safe nearby, of the famous Piltdown man, an early Briton who strolled the barren south
some 600,000 years ago.

``Scientists disclosed Saturday that someone had made a monkey out of them. After a new
test with modern techniques, they declared the skull an 'elaborate hoax.'

``The Piltdown cranium is genuine, the scientists said, but the head's jaw and molars are those
of a modern ape, stained and filed down to simulate the fossil specimens of early man.

``The faking has 'no parallel' in the history of palaeontology, two anatomy professors of
Oxford University and a leading geologist of the British Museum said in a joint report.

``A British lawyer, Charles Dawson, found the preice age brain-box in a gravel pit on his
estate at Piltdown, East Sussex, in 1912, after workmen had dug up what they described as a
'queer looking coconut.'

``A blow from a pickaxe had broken the bone, splintering it many times. Dawson, an
antiquarian, by hobby, began rummaging in the pit and found the other pieces. Eminent
scientists excitedly pronounced the find the oldest ever of its kind. The lower jaw was found
later in the gravel pit. Scientists linked it with the cranium.

``Modern fluorine test has now revealed the jaw bone had been stained with bichromate of
potash and iron and the teeth artificially pared down. The jaw and teeth, the scientists say,
must have been 'planted' in the gravel pit but they emphasized that the brain case is still
regarded as a genuine fossil.''

The following day other remarks appeared in the press, and one geologist, Alvan Marston,
who had long claimed the Piltdown jaw came from an ape, said: ``The blame for this
Piltdown hoax does not lie at the door of Mr. Dawson. He was meticulous and honest.''

Perhaps the Trinil Ape-Man, if the bones could be examined, might reveal something just as
startling, but Dr. Dubois locked them up and they have not been subject to later examination.

There have been numerous other ``finds'' from which ``primitive men'' have been
manufactured, such as the La Quina Lady, the Heidelberg Man, the Moustier Man, the Pekin
Man, and others needless to mention. The fact remains that they have all been manufactured
from a few scattered fragments of bones with no certain evidence that these fragments were
parts of the same skeletal remains. However the work has gone on with a zeal unworthy of a
true scientific cause, and these scattered fragments have been palmed off on the world with a
brazen disregard of the facts that is appalling. Most of the bones thus discovered have been
found in positions many yards apart and there is no proper evidence that they belonged to the
same individual. Moreover, the bones gathered from isolated points, or deposits, have been
taken as a criterion and in imagination made to represent an entire race of imaginary people.
It has been shown in some instances that some of the bones thus gathered are not of human
origin. Even the experts have disagreed. Like the imaginary ancestor of the horse, the
complete remains of these specimens which are placed on exhibition have never been
discovered, but from a few scattered bones gathered under most uncertain conditions have
been used to deceive the public. I repeat that if the doctrine of organic evolution were based
in truth, there would be no need to search the world over for the remains of ``missing links.''
The world would be full of them. They would be walking our streets, and as William
Jennings Bryan has said, the hypothesis of evolution ``would find support on every foot of the
earth's surface.'' As it is, the advocates of this pernicious theory go to the most ridiculous
lengths and resort to the most absurd conclusions based on imaginary discoveries and fables.
They are possessed with imaginary minds and when the facts fail them, as the facts always
do, they can create species and groups and supply missing parts which in their imaginations
disappeared millions of years ago.

It is unnecessary to continue this phase of the subject more than to say that in all of these
``finds'' the wish has been father to the thought, so overly anxious have these ``discoverers''
been to find some connecting links between man and the lower animals that would give
evidence of a common origin. These ``missing links'' have not been forthcoming and the
plotters have been forced to resort to fraud and deception to bolster up their futile attempts to
prove a Satan-inspired cause, the real purpose being to destroy faith in God. Verily the words
of our Lord are true and made manifest in this last dispensation with confirmation in the
actions of men: ``And this is the condemnation, that light is come into the world, and men
loved darkness rather than light, because their deeds were evil. For everyone that doeth evil
hateth the light, neither cometh to the light, lest his deed should be reproved. But he that
doeth truth cometh to the light, that his deeds may be made manifest, that they are wrought in
God.'' 27

\newpage
REFERENCES—CHAPTER SEVEN

Footnotes

1. Romans 8:13-17.

2. John 3:1-3.

3. Rev. 21:7.

4. Moses 6:65-68.

5. John 20:17.

6. White, Dr. A. D., \textit{History of the Warfare of Science with Christian Theology}, Vol. 1, pp.
20-237.

7. \textit{Ibid.}, p. 56.

8. \textit{Ibid.}, p. 222.

9. \textit{Ibid.}, p. 223; Draper, Dr. John W., \textit{Conflict Between Religion and Science}, pp. 199-200;
Millikan, Dr. R. A., \textit{Science and Religion}, p. 12.

10. D. \& C. 18:10-16.

11. \textit{Man and The Universe}, Art. The Beginning of Life on the Earth.

12. Pack, Dr. F. J., \textit{Science and Belief in God}, p. 84.

13. Darwin, Charles, \textit{Descent of Man}, pp. 164-165, 1897 edition.

14. Pack, Dr. F. J., \textit{Science and Belief in God}, p. 83.

15. Bryan, W. J., \textit{In His Image}, pp. 91-93, 1922 edition.

16. \textit{Ibid.}, pp. 97-101.

17. Price, G. M., \textit{The New Geology}, pp. 695-697.

18. \textit{Ibid.}, p. 697.

19. McCann, A. W., \textit{God—or Gorilla}, pp 28-29.

20. Price, G. M., \textit{The New Geology}, p. 700.

21. McCann, A. W., \textit{God—or Gorilla}, p. 36.

22. \textit{Ibid.}, pp. 36-37.

23. Price, G. M., \textit{The New Geology}, p. 702.

24. \textit{Ibid.}, p. 702.

25. \textit{Ibid.}, p. 702.

26. McCann, A. W., \textit{God—or Gorilla}, pp 6-7.

27. John 3:19-21.

