\chapter{ADAM'S PLACE IN THE EARTH'S DESTINY—1}

IN the study of spiritual things and the plan of salvation members of the Church should not
turn for guidance to the philosophical and scientific theories of men. Revelation is superior to
science. The evolutionary writers have to attack revelation and any communication coming
from God or forsake their theories. The word of the Lord through his servants the prophets is
the only safe guide that man has. As I have said before, Satan dominates the thinking of the
world today and he has put into the hearts of men ideas and deductions that are diametrically
opposed to the revealed word that has come to us through the scriptures. Naturally the
advocates of these theories will attack the revelations as recorded in the Bible, and the other
standard works of the Church which sustain the teachings of the Bible. Latter-day Saints
should have abiding faith in the revelations coming through the Prophet Joseph Smith. This
is our duty by commandment from the Lord. We are not left in the most helpless position that
our friends of the so-called Christian denominations find themselves. Not only do we have
the accounts of creation and the historical records of the prophets, but we have the renewed
accounts given us by the Lord in our own dispensation. The Lord has by revelation in our day
confirmed the story of Moses, of Abraham and the Israelite prophets so that we have the
double assurance. Moreover having received the gift of the Holy Ghost it becomes our duty
not only to believe these things, but to \textit{know} them.

We have the Book of Mormon which comes fully attested by witnesses in fulfilment of the
law the Lord has given concerning witnesses. We are taught that the ways of the Lord are not
man's ways. He gave the law that all things should be established in the mouths of witnesses.
Two may be sufficient, more will be better. When the Savior was called in question by the
Pharisees as an impostor because he claimed to be the Son of God, the Lord referred to this
law and applied it. It is as follows:

The Pharisees therefore said unto him, Thou bearest record of thyself; thy record is not true.

Jesus answered and said unto them, Though I bear record of myself, yet my record is true: for
I know whence I came, and whither I go; but ye cannot tell whence I came, and whither I go.

Ye judge after the flesh; I judge no man.

And yet if I judge, my judgment is true: for I am not alone, but \textit{I and the Father that sent me.}

It is also written in your law, that the testimony of two men is true.

I am one that bear witness of myself, and the Father that sent me beareth witness of me. 1

Here the Lord called attention to the law of witnesses. Paul also writing to the Corinthian
saints said:

This is the third time I am coming to you. In the mouth of two or three witnesses shall every
word be established. 2

And in the coming forth of the Book of Mormon the Lord said to Nephi:

Wherefore, at that day when the book shall be delivered unto the man of whom I have
spoken, the book shall be hid from the eyes of the world, that the eyes of none shall behold it
save it be that three witnesses shall behold it, by the power of God, besides him to whom the
book shall be delivered; and they shall testify to the truth of the book and the things therein.

And there is none other which shall view it, save it be a few according to the will of God, to
bear testimony of his word unto the children of men; for the Lord God hath said that the
words of the faithful should speak as if it were from the dead.

Wherefore, the Lord God will proceed to bring forth the words of the book; \textit{and in the mouth
of as many witnesses as seemeth him good will he establish his word; and wo be unto him
that rejecteth the word of God.} 3

In fulfilment of this promise the Lord raised up four witnesses to bear testimony to the truth
of the Book of Mormon; men who had a hand in its coming forth. Three of them are called
``special witnesses,'' Oliver Cowdery, David Whitmer and Martin Harris. These men had the
privilege of seeing the plates of Mormon. They examined them and this in the presence of the
angel. They also heard the voice of God from the heavens ``commanding'' them that they
should bear record of it; ``wherefore,'' said they, ``to be obedient unto the commandments of
God, we bear testimony of these things.'' This makes four special witnesses. Then there were
eight others, who were privileged to see and handle these plates and they likewise were
commanded to bear testimony to what they had seen. This fully establishes the authenticity
of the Book of Mormon. Then again, in the Book of Mormon is the promise made by Moroni
(Moroni 10:4.):

And when ye shall receive these things, I would exhort you that ye would ask God, the
Eternal Father, in the name of Christ, if these things are not true; and if ye shall ask with a
sincere heart, with real intent, having faith in Christ, he will manifest the truth of it unto you,
by the power of the Holy Ghost.

There are legions now living who have received this testimony and know that the record is
true. There are others who have also read the Book of Mormon, but without the sincere desire
to know the truth, but with the object in view of opposing it. All such are under heavy
condemnation and will answer for it at the judgment seat of God. Nephi warned them (2
Nephi 33.), and likewise Moroni, who sealed the record, who said:

And I exhort you to remember these things; for the time speedily cometh that ye shall know
that I lie not, for ye shall see me at the bar of God; and the Lord God will say unto you: Did I
not declare my words unto you, which were written by this man, like as one crying from the
dead, yea, even as one speaking out of the dust?

I declare these things unto the fulfilling of the prophecies. And behold, they shall proceed
forth out of the mouth of the everlasting God; and his word shall hiss forth from generation
to generation.

And God shall show unto you, that that which I have written is true. (Moroni 10:27-29.)

Then again, whenever the Lord revealed the Priesthood, or the keys of the Priesthood, in this
dispensation, there were always two witnesses to receive them, thus complying with the
divine law. The testimonies of these men therefore become binding upon the world, and
justly the Lord will say: ``Wo be unto him that rejecteth the word of God.'' Having said this I
will present some of the testimony concerning Father Adam, his existence and mission in the
world, which testimony comes from these established witnesses. Before doing so, however,
let me refer to some of the views that have grown out of organic evolution and geology
particularly, and which are taught in most of the courses on science and by those who attempt
to write on ancient history. It is natural for these advocates to deny that Adam existed, and to
mock at the stories concerning him found in the Bible. They cannot admit that there was an
Adam. They must reject the fall, and consequently they are forced to reject the atonement of
Jesus Christ and his resurrection. It is fatal to their theories to admit any of these things.
Therefore they reject the Bible and say that the Old Testament is based on Chaldean-
Babylonian myths. 4 Such scientists as Darwin, Wallace, Huxley, Sir Arthur Keith, Andrew
D. White, John W. Draper and Robert A. Millikan, and scores of others, have turned from
Christianity to Darwinism and lost their faith; but none have been more bitter than Andrew
Dickson White.

Dr. White treats the story of Adam and the fall with bitter contempt. Dr. John William
Draper with less bitterness declares just as emphatically that it is absurd. Dr. Millikan,
without bitterness, speaks of them as having originated from the ``crudest beginnings.'' Sir
David Keith treats the story of Adam as an inconsistent one, overthrown by Darwin. So the
stories go, and we find this doctrine in our public schools and universities; and thus faith in
the Gospel of Jesus Christ is threatened. Therefore it is necessary to protect our youth by
proclaiming the truth and speaking against these theories. We, therefore, proclaim the truth
and speak against these false theories and confirm the teachings of the Bible by the
revelations of the present dispensation.

Nearly one century ago, President Brigham Young addressing the members of the Church
made this prediction:

We take this book, the Bible, which I expect to be voted out of the so-called Christian world
very soon, they are coming to it as fast as possible; I say we take this book for our guide, for
our rule of action; we take it as the foundation of our faith. It points the way to salvation like
a finger-board pointing to a city, or a map which designates the locality of mountains, rivers,
or the latitude and longitude of any place on the face of the earth that we desire to find, as we
have no better sense than to believe it; hence, I say that the Latter-day Saints have the most
natural faith and belief of any people on the face of the earth. 5

The Christian world is actually coming to the point that they will dismiss the Bible from their
schools; and by and by they will dismiss it from their pulpits and get one to suit themselves.
They will hew out for themselves cisterns that will hold no water. 6

President Wilford Woodruff in January 1871, added his testimony as follows:

We have been true and faithful in our testimony to the inhabitants of the earth; and as the
world generally has rejected our testimony the Lord has withdrawn his spirit from the people
in a great measure; and the religion they once enjoyed is as nothing to many of them.
Infidelity prevails throughout the world; very few, either priests or people, believe in a literal
fulfilment of the Bible. They have a theory, but as to believing in a real fulfilment of
prophecy, or that the Lord meant what he said and said what he meant, that is out of the
question—very few believe it. 7

Let us, therefore, confirm the Biblical account of the origin of man from the revelations in
our other Standard Works. First from the Book of Mormon. When Nephi and his brothers
returned with the brass plates to the tent of their father, Lehi took them and searched them
carefully, and rejoiced when he discovered what they contained; for they consisted of the five
books of Moses, which destructive critics of the Bible say were not written until the time of
the Exile, or later, and an account of the creation. The point concerning us now is as follows:

And after they had given thanks unto the God of Israel, my father, Lehi, took the records
which were engraven upon the plates of brass, and he did search them from the beginning.

And he beheld that they did contain the five books of Moses, which gave an account of the
creation of the world, and also of Adam and Eve, who were our first parents. 8

Commenting on what he discovered, Lehi wrote:

Wherefore, the Lord God gave unto man that he should act for himself. Wherefore, man
could not act for himself save it should be that he was enticed by the one or the other.

And I, Lehi, according to the things which I have read, must needs suppose that an angel of
God, according to that which is written, had fallen from heaven; wherefore, he became a
devil, having sought that which was evil before God.

And because he had fallen from heaven, and had become miserable forever, he sought also
the misery of all mankind. Wherefore, he said unto Eve, yea, even that old serpent, who is the
devil, who is the father of all lies, wherefore he said: Partake of the forbidden fruit, and ye
shall not die, but ye shall be as God, knowing good and evil.

And after Adam and Eve had partaken of the forbidden fruit they were driven out of the
garden of Eden, to till the earth.

And they have brought forth children; yea, \textit{even the family of all the earth.}

And the days of the children of men were prolonged, according to the will of God, that they
might repent while in the flesh; wherefore, their state became a state of probation, and their
time was lengthened, according to the commandments which the Lord God gave unto the
children of men. For he gave commandment that all men must repent; for he showed unto all
men that they were lost, because of the transgression of their parents. 9

Alma in his instruction to his son Corianton also testifies of this beginning:

Now behold, my son, I will explain this thing unto thee. For behold, after the Lord sent our
first parents forth from the garden of Eden, to till the ground, from whence they were taken—
yea, he drew out the man, and he placed at the east end of the garden of Eden, cherubim, and
a flaming sword which turned every way, to keep the tree of life—

Now, we see that the man had become as God, knowing good and evil; and lest he should put
forth his hand, and take also of the tree of life, and eat and live forever, the Lord God placed
cherubim and the flaming sword, that he should not partake of the fruit—

And thus we see, that there was a time granted unto man to repent, yea, a probationary time,
a time to repent and serve God.

For behold, if Adam had put forth his hand immediately, and partaken of the tree of life, he
would have lived forever, according to the word of God, having no space for repentance; yea,
and also the word of God would have been void, and the great plan of salvation would have
been frustrated.

But behold, it was appointed unto man to die—therefore, as they were cut off from the tree of
life they should be cut off from the face of the earth—and man became lost forever, yea, they
became fallen man.

And now, ye see by this that our first parents were cut off both temporally and spiritually
from the presence of the Lord; and thus we see they became subjects to follow after their
own will.

Now behold, it was not expedient that man should be reclaimed from this temporal death, for
that would destroy the great plan of happiness.

Therefore, as the soul could never die, and the fall had brought upon all mankind a spiritual
death as well as a temporal, that is, they were cut off from the presence of the Lord, it was
expedient that mankind should be reclaimed from this spiritual death.

Therefore, as they had become carnal, sensual, and devilish, by nature, this probationary state
became a state for them to prepare; it became a preparatory state.

And now remember, my son, if it were not for the plan of redemption, (laying it aside) as
soon as they were dead their souls were miserable, being cut off from the presence of the
Lord. 10

\textit{The words of Jesus Christ to Mormon:}

Listen to the words of Christ, your Redeemer, your Lord and your God. Behold, I came into
the world not to call the righteous but sinners to repentance; the whole need no physician, but
they that are sick; wherefore, little children are whole, for they are not capable of committing
sin; wherefore the curse of Adam is taken from them in me, that it hath no power over them;
and the law of circumcision is done away in me. 11

\textit{The word of the Lord to Moroni:}

Behold, I would exhort you that when ye shall read these things, if it be wisdom in God that
ye should read them, that ye would remember how merciful the Lord hath been unto the
children of men, from the creation of Adam even down until the time that ye shall receive
these things, and ponder it in your hearts. 12

\textit{Evidence from the Pearl of Great Price:}

And I, God, said unto mine Only Begotten, which was with me from the beginning: Let us
make man in our image, after our likeness; and it was so. And I, God, said: Let them have
dominion over the fishes of the sea, and over the fowl of the air, and over the cattle, and over
all the earth, and over every creeping thing that creepeth upon the earth.

And I, God, created man in mine own image, in the image of mine Only Begotten created I
him; male and female created I them. 13

And I, the Lord God, formed man from the dust of the ground, and breathed into his nostrils
the breath of life; and man became a living soul, the \textit{first flesh upon the earth, the first man
also}; nevertheless, all things were before created; but spiritually were they created and made
according to my word. 14

These passages are not taken from the Bible account but from the revelation to the Prophet
Joseph Smith. They confirm what is written in Genesis which the ``wise'' and ``prudent'' of the
present age treat with scorn because they conflict with their theories. Perhaps a word of
explanation is necessary here in relation to the expression, \textit{``the first flesh upon the earth.''}
We are taught in the accounts of creation in the Bible, the Book of Mormon and in the Book
of Abraham, as well as in the Book of Moses, from which I quote, that man was the last of
the creations placed upon the earth after all other things were prepared for him. This order of
sequence is historically correct. Verse five of chapter three in Moses' account is but an
explanation of the fact that all things were created in the spirit before they were naturally on
the earth. The same is true of verse five in chapter two of Genesis. These expressions about
the spirit creation are not intended to change the order of creation as placed physically on the
earth as recorded in the first chapter of Genesis.

The thought that man was the first living thing upon the earth and that he was placed here in
a state of desolation, before there was any vegetation or animal life on the land, in the air, or
in the sea, does violence to the entire account of creation as well as to reason. Such an idea
naturally causes such thoughts as these: What was the length of this desolation when he was
alone? What purpose could there have been in placing him here before the earth was
prepared for him? If the desolation was long or short, it must have been a very comfortless
place. Certainly it was not the place the Lord called ``good'' and ``finished.'' The expression,
``the first flesh upon the earth,'' is simply a statement of the fact that Adam—\textit{the first man on
the earth}—was, by reason of his transgression the first to partake of mortality. That is to say,
the fall brought the mortal as well as the spiritual death. ``Mortality'' and ``flesh'' are often
used as synonymous terms. The account in Genesis of creation informs us that Adam and
Eve were the \textit{first} to become subject to the mortal, or temporal, death. This fall brought death
into the world. We will say more of this at another time. The mortal condition was extended
so that it eventually embraces every creature on the land, in the air, and in the sea. In this way
all creatures became ``flesh,'' but not until \textit{after} man had so become. Here are a few
references on this point:

Thus saith the Lord; Cursed be the man that trusteth in man, and maketh flesh his arm, and
whose heart departeth from the Lord. (Jer. 17:5.)

Therefore, brethren, we are debtors, not to the flesh, to live after the flesh.

For if ye live after the flesh, ye shall die: but if ye through the Spirit do mortify the deeds of
the body, ye shall live. (Romans 8:12.)

Ye judge after the flesh; I judge no man. (Jesus in John 8:15.)

That he no longer should live the rest of his time in the flesh to the lusts of men, but to the
will of God. (1 Peter 4:2.)

And the days of the children of men were prolonged, according to the will of God, that they
might repent while in the flesh; wherefore, their state became a state of probation. . . . (2
Nephi 2:21.)

For all flesh is corrupted before me; and the powers of darkness prevail upon the earth,
among the children of men, in the presence of all the hosts of heaven. (D. \& C. 38:11.)

So Adam became the ``first flesh on the earth,'' through the fall, not that he was placed here
before any other creatures.

But God hath made known unto our fathers that all men must repent.

And he called upon our father Adam by his own voice, saying: I am God; I made the world,
and men before they were in the flesh.

And he also said unto him: If thou wilt turn unto me, and hearken unto my voice, and
believe, and repent of all thy transgressions, and be baptized, even in water, in the name of
mine Only Begotten Son, who is full of grace and truth, which is Jesus Christ, the only name
which shall be given under heaven, whereby salvation shall come unto the children of men,
ye shall receive the gift of the Holy Ghost, asking all things in his name, and whatsoever ye
shall ask, it shall be given you.

And our father Adam spake unto the Lord, and said: Why it is that men must repent and be
baptized in water? And the Lord said unto Adam: Behold I have forgiven thee thy
transgression in the Garden of Eden.

Hence came the saying abroad among the people, That the Son of God hath atoned for
orignal guilt, wherein the sins of the parents cannot be answered upon the heads of the
children, for they are whole from the foundation of the world.

And the Lord spake unto Adam, saying: Inasmuch as thy children are conceived in sin, even
so when they begin to grow up, sin conceiveth in their hearts, and they taste the bitter, that
they may know to prize the good.

And it is given unto them to know good from evil; wherefore they are agents unto
themselves, and I have given unto you another law and commandment. 15

Let us now consider some of the revelations in the Doctrine and Covenants in relation to
Adam's place in the destiny of the earth. In August 1830, the Lord gave a revelation to the
Prophet Joseph Smith in regard to the sacrament. The circumstances were as follows: In
preparation for a religious service it was discovered that the little group had no wine for the
sacrament. Thereupon the Prophet Joseph Smith set out to buy some. He was met by an angel
and received a revelation in which instruction was given that they should not purchase wine
from their enemies and the Lord said to the Prophet: ``. . . it mattereth not what ye shall eat or
what ye shall drink when ye partake of the sacrament, if it so be that ye do it with an eye
single to my glory—remembering unto the Father my body which was laid down for you,
and my blood which was shed for the remission of your sins. Wherefore, a commandment I
give unto you, that you shall not purchase wine neither strong drink of your enemies;
Wherefore, you shall partake of none except it is made new among you; yea, in this my
Father's kingdom which shall be built up on the earth. 16 Then the Lord continued and said
that he would partake of ``the fruit of the vine'' when the ``hour cometh,'' with Moroni, Elias,
John, the son of Zacharias and with Elijah and other ancient prophets, and also with
``Michael, or Adam, the father of all, the prince of all, the ancient of days.'' (D. \& C. 27:2-
11.)

Here Michael is identified as Adam, and he is called the ancient of days, the prince of all,
indicating that he stands at the head of the human family. In the same year (September 1830)
the Lord gave another revelation in which Michael, or Adam, is mentioned and his status
further discussed in the following passage:

But, behold, verily I say unto you, before the earth shall pass away, Michael, mine archangel,
shall sound his trump, and then shall all the dead awake, for their graves shall be opened, and
they shall come forth—yea, even all. 17

Then the Lord continues his instruction and refers to Adam's fall and the results thereof as
follows:

Wherefore, verily I say unto you that all things unto me are spiritual, and not at any time
have I given unto you a law which was temporal; neither any man, nor the children of men;
neither Adam, your father, whom I created.

Behold, I gave unto him that he should be an agent unto himself; and I gave unto him
commandment, but no temporal commandment gave I unto him, for my commandments are
spiritual; they are not natural nor temporal, neither carnal nor sensual.

And it came to pass that Adam, being tempted of the devil—for, behold, the devil was before
Adam, for he rebelled against me, saying, Give me thine honor, which is my power; and also
a third part of the hosts of heaven turned he away from me because of their agency;

And they were thrust down, and thus came the devil and his angels;

And, behold, there is a place prepared for them from the beginning, which place is hell.

And it must needs be that the devil should tempt the children of men, or they could not be
agents unto themselves; for if they never should have bitter they could not know the sweet—

Wherefore, it came to pass that the devil tempted Adam, and he partook of the forbidden fruit
and transgressed the commandment, wherein he became subject to the will of the devil,
because he yielded unto temptation.

Wherefore, I, the Lord God, caused that he should be cast out from the Garden of Eden, from
my presence, because of his transgression, wherein he became spiritually dead, which is the
first death, even that same death which is the last death, which is spiritual, which shall be
pronounced upon the wicked when I shall say: Depart, ye cursed.

But, behold, I say unto you that I, the Lord God, gave unto Adam and unto his seed, that they
should not die as to the temporal death, until I, the Lord God, should send forth angels to
declare unto them repentance and redemption, through faith on the name of mine Only
Begotten Son.

And thus did I, the Lord God, appoint unto man the days of his probation—that by his natural
death he might be raised in immortality unto eternal life, even as many as would believe;

And they that believe not unto eternal damnation; for they cannot be redeemed from their
spiritual fall, because they repent not. 18

Here we have the word of the Lord in our day stating that Michael is Adam, that he was in
the Garden of Eden and because of temptation he was driven out and partook of the two
deaths—spiritual and mortal. This came because he yielded to the enticings of the devil who
was in the beginning and because of rebellion was cast out of heaven and took with him one
third of the hosts of heaven because of their agency and rebellion. All of this confirms the
stories in the Bible which our friends who follow the lead of Darwin ridicule and claim came
from the ancient myths and legends borrowed by Israel from the Chaldeans and Babylonians.
These men who are without spiritual guidance who have been taken captive by the devil have
promulgated these stories and are largely responsible in leading tens of thousands away from
the path of truth and from God.

Again in Section 84, verse sixteen, the Lord informs us that the Priesthood came down to
Moses, from Adam, \textit{who was the first man}. In Section 107, we have further light in relation
to Adam which is important enough for a later consideration. In Section 128, we learn further
that Adam, among other prophets who held keys of Priesthood, revealed his keys to the
Prophet Joseph Smith which was essential to the fulfilling of the promises in relation to the
restoration of all things.

The fact revealed that Adam holds authority under Jesus Christ and that he is Michael, the
archangel, we must conclude, establishes him has a most worthy character. I appeal to all
members of the Church not to permit these uninspired men who are trained in the
philosophies of the world to lead them astray. The Lord has promised that when he comes he
will reveal all things. 19 In that day all these theories which are in conflict with the revealed
word of the Lord shall perish, for nothing can remain except eternal truth and that will be
established by the revelations of God.

\newpage
REFERENCES—CHAPTER FIFTEEN

Footnotes

1. John 8:13-18.

2. 2 Cor. 13:1.

3. 2 Nephi 27:12-14.

4. White, Dr. A. D., \textit{Warfare of Science and Theology}, Vol. 1, pp. 41, 96; Draper, Dr. J. W.,
\textit{Conflict Between Religion and Science}, pp. 195-6; Millikan, Dr. R. A., \textit{Evolution in Science
and Religion}, p. 66.

5. \textit{Journal of Discourses}, Vol. 13, p. 236.

6. \textit{Ibid.}, Vol. 13, p. 213.

7. \textit{Journal History}, Jan. 1, 1871.

8. 1 Nephi 5:10-11.

9. 2 Nephi 2:16-21.

10. Alma 42:2-11.

11. Moroni 8:8.

12. Ibid., 10:3.

13. Moses 2:26-27.

14. \textit{Ibid.}, 3:7.

15. \textit{Ibid.}, 6:50-56.

16. D. \& C. 27:2-3.

17. \textit{Ibid.}, 29:26.

18. \textit{Ibid.}, 29:34-44.

19. \textit{Ibid.}, 101:32.

