\chapter{ADAM'S PLACE IN THE EARTH'S DESTINY—3}

THERE are other testimonies coming from our brethren, who were trained under the
teachings of the Prophet Joseph Smith and who are able to speak by virtue of the inspiration
of the Spirit of the Lord by which they have been led. Their teachings should be heeded by
the members of the Church. The following article by the First Presidency was written for the
purpose of directing our Church members in the light of revealed truth and to fortify them
against the false doctrines and philosophies so prevalent in the world.

THE ORIGIN OF MAN By THE FIRST PRESIDENCY OF THE CHURCH

\textit{"God created man in his own image."}

Inquiries arise from time to time respecting the attitude of the Church of Jesus Christ of
Latter-day Saints upon questions which, though not vital from a doctrinal standpoint, are
closely connected with the fundamental principles of salvation. The latest inquiry of this kind
that has reached us is in relation to the origin of man. It is believed that a statement of the
position held by the Church upon this important subject will be timely and productive of
good.

In presenting the statement that follows we are not conscious of putting forth anything
essentially new; neither is it our desire so to do. Truth is what we wish to present, and truth—
eternal truth—is fundamentally old. A restatement of the original attitude of the Church
relative to this matter is all that will be attempted here. To tell the truth as God has revealed
it, and commend it to the acceptance of those who need to conform their opinions thereto, is
the sole purpose of this presentation.

"God created man in his own image, in the image of God created he him; male and female
created he them." In these plain and pointed words the inspired author of the book of Genesis
made known to the world the truth concerning the origin of the human family. Moses, the
prophet-historian, "learned," as we are told, "in all the wisdom of the Egyptians," when
making this important announcement, was not voicing a mere opinion, a theory derived from
his researches into the occult lore of that ancient people. He was speaking as the mouthpiece
of God, and his solemn declaration was for all time and for all people. No subsequent
revelator of the truth has contradicted the great leader and law-giver of Israel. All who have
since spoken by divine authority upon this theme have confirmed his simple and sublime
proclamation. Nor could it be otherwise. Truth has but one source, and all revelations from
heaven are harmonious with each other. The omnipotent Creator, the maker of heaven and
earth—had shown unto Moses everything pertaining to this planet, including the facts
relating to man's origin, and the authoritative pronouncement of that mighty prophet and seer
to the house of Israel, and through Israel to the whole world, is couched in the simplest
clause: "God created man in his own Image." (Genesis 1:27; Pearl of Great Price—Book of
Moses, 1:27-41.)

The creation was two-fold—first spiritual, secondly temporal. This truth, also, Moses plainly
taught—much more plainly than it has come down to us in the imperfect translations of theBible that are now in use. Therein the fact of a spiritual creation, antedating the temporal
creation, is strongly implied, but the proof of it is not so clear and conclusive as in other
records held by the Latter-day Saints to be of equal authority with the Jewish scriptures. The
partial obscurity of the latter upon the point in question is owing, no doubt, to the loss of
those "plain and precious" parts of sacred writ, which, as the Book of Mormon informs us,
have been taken away from the Bible during its passage down the centuries. (1 Nephi 13:24-
29.) Some of these missing parts the Prophet Joseph Smith undertook to restore when he
revised those scriptures by the spirit of revelation, the result being that more complete
account of the creation which is found in the book of Moses, previously cited. Note the
following passages:

"And now, behold I say unto you, that these are the generations of the heaven and of the
earth, when they were created, in the day that I, the Lord God, made the heaven and the
earth:

"And every plant of the field before it was in the earth, and every herb of the field before it
grew. For I, the Lord God, created all things, of which I have spoken, spiritually, before they
were naturally upon the face of the earth. For I, the Lord God, had not caused it to rain upon
the face of the earth. And I, the Lord God, had created all the children of men, and not yet a
man to till the ground: for in heaven created I them; and there was not yet flesh upon the
earth, neither in the water, neither in the air:

"But I, the Lord God, spake, and there went up a mist from the earth, and watered the whole
face of the ground.

"And I, the Lord God, formed man from the dust of the ground, and breathed into his nostrils
the breath of life; and man became a living soul, the first flesh upon the earth, the first man
also; nevertheless, all things were before created; but spiritually were they created and made
according to my word." (Pearl of Great Price—Book of Moses, 3:4-7. See also chapters 1
and 2, and compare with Genesis 1 and 2.)

These two points being established, namely, the creation of man in the image of God, and the
two-fold character of the creation, let us now inquire: What was the form of man, in the spirit
and in the body, as originally created? In a general way the answer is given in the words
chosen as the text of this treatise. "God created man in his own image." It is more explicitly
rendered in the Book of Mormon thus: "All men were created in the beginning after mine
own image." (Ether 3:15.) It is the Father who is speaking. If, therefore, we can ascertain the
form of the "Father of spirits," "The God of the spirits of all flesh," we shall be able to
discover the form of the original man.

Jesus Christ, the Son of God, is "the express image" of his Father's person. (Hebrews 1:3.) He
walked the earth as a human being, as a perfect man, and said, in answer to a question put to
him: "He that hath seen me hath seen the Father." (John 14:9.) This alone ought to solve the
problem to the satisfaction of every thoughtful, reverent mind. The conclusion is irresistible,
that if the Son of God be the express image (that is, likeness) of his Father's person, then his
Father is in the form of man; for that was the form of the Son of God, not only during his
mortal life, but before his mortal birth, and after his resurrection. It was in this form that the
Father and the Son, as two personages, appeared to Joseph Smith, when, as a boy of fourteen
years, he received his first vision. Then if God made man—the first man—in his own imageand likeness, he must have made him like unto Christ, and consequently like unto men of
Christ's time and of the present day. That man was made in the image of Christ, is positively
stated in the book of Moses: "And I, God, said unto mine Only Begotten, which was with me
from the beginning: Let us make man in our image, after our likeness; and it was so. . . . And
I, God, created man in mine own image, in the image of mine Only Begotten created I him;
male and female created I them." (2:26, 27.)

The Father of Jesus is our Father also, Jesus himself taught this truth, when he instructed his
disciples how to pray: "Our Father which art in heaven," etc. Jesus, however, is the Firstborn
among all the sons of God—the first begotten in the spirit, and the Only Begotten in the
flesh. He is our elder brother, and we, like him, are in the image of God. All men and women
are in the similitude of the universal Father and Mother, and are literally the sons and
daughters of Deity.

"God created man in his own image." This is just as true of the spirit as it is of the body,
which is only the clothing of the spirit, its complement; the two together constituting the
soul. The spirit of man is in the form of man, and the spirits of all creatures are in the likeness
of their bodies. This was plainly taught by the Prophet Joseph Smith. (Doctrine and
Covenants, 77:2.)

Here is further evidence of the fact. More than seven hundred years before Moses was shown
the things pertaining to this earth, another great prophet, known to us as the brother of Jared,
was similarly favored by the Lord. He was even permitted to behold the spirit-body of the
foreordained Savior, prior to his incarnation; and so like the body of a man was his spirit in
form and appearance, that the prophet thought he was gazing upon a being of flesh and
blood. He first saw the finger and then the entire body of the Lord—all in the spirit. The
Book of Mormon says of this wonderful manifestation:

"And it came to pass that when the brother of Jared had said these words, behold, the Lord
stretched forth his hand and touched the stones one by one with his finger. And the veil was
taken from off the eyes of the brother of Jared, and he saw the finger of the Lord; and it was
as the finger of a man, like unto flesh and blood; and the brother of Jared fell down before the
Lord, for he was struck with fear.

"And the Lord saw that the brother of Jared had fallen to the earth; and the Lord said unto
him: Arise, why hast thou fallen?

"And he saith unto the Lord: I saw the finger of the Lord, and I feared lest he should smite
me; for I knew not that the Lord had flesh and blood.

"And the Lord said unto him: Because of thy faith thou hast seen that I shall take upon me
flesh and blood; and never has man come before me with such exceeding faith as thou hast;
for were it not so, ye could not have seen my finger. Sawest thou more than this?

"And he answered: Nay; Lord, show thyself unto me.

"And the Lord said unto him: Believest thou the words which I shall speak?

"And he answered: Yea, Lord, I know that thou speakest the truth, for thou art a God of truth,
and canst not lie.

"And when he had said these words, behold, the Lord showed himself unto him, and said:
Because thou knowest these things ye are redeemed from the fall; therefore ye are brought
back into my presence; therefore I show myself unto you.

"Behold, I am he who was prepared from the foundation of the world to redeem my people.
Behold, I am Jesus Christ. I am the Father and the Son. In me shall all mankind have light,
and that eternally, even they who shall believe on my name; and they shall become my sons
and my daughters.

"And never have I showed myself unto man whom I have created, for never has man
believed in me as thou hast. Seest thou that ye are created after mine own image? Yea, even
all men were created in the beginning after mine own image.

"Behold, this body, which ye now behold, is the body of my spirit; and man have I created
after the body of my spirit; and even as I appear unto thee to be in the spirit will I appear unto
my people in the flesh." (Ether 3:6-16.)

What more is needed to convince us that man, both in spirit and in body, is the image and
likeness of God, and that God himself is in the form of man?

When the divine Being whose spirit-body the brother of Jared beheld, took upon him flesh
and blood, he appeared as a man, having "body, parts and passions," like other men, though
vastly superior to all others, because he was God, even the Son of God, the Word made flesh:
in him "dwelt the fulness of the Godhead bodily." And why should he not appear as a man?
That was the form of his spirit, and it must needs have an appropriate covering, a suitable
tabernacle. He came into the world as he had promised to come (3 Nephi 1:13) taking an
infant tabernacle, and developing it gradually to the fulness of his spirit stature. He came as
man had been coming for ages, and as man has continued to come ever since. Jesus,
however, as shown, was the Only Begotten of God in the flesh.9

"Adam, our great progenitor, "the first man," was, like Christ, a pre-existent spirit, and like
Christ he took upon him an appropriate body, the body of a man, and so became a "living
soul." The doctrine of the pre-existence,—revealed so plainly, particularly in latter days,
pours a wonderful flood of light upon the otherwise mysterious problem of man's origin. It
shows that man, as a spirit, was begotten and born of heavenly parents, and reared to
maturity in the eternal mansions of the Father, prior to coming upon the earth in a temporal
body to undergo an experience in mortality. It teaches that all men existed in the spirit before
any man existed in the flesh, and that all who have inhabited the earth since Adam have taken
bodies and become souls in like manner.

It is held by some that Adam was not the first man upon this earth, and that the original
human being was a development from lower orders of the animal creation. These, however,
are the theories of men. The word of the Lord declares that Adam was "the first man of all
men" (Moses 1:34), and we are therefore in duty bound to regard him as the primal parent of
the race. It was shown to the brother of Jared that all men were created in the beginning after
the image of God; and whether we take this to mean the spirit or the body, or both, itcommits us to the same conclusion: Man began life as a human being, in the likeness of our
heavenly Father.

True it is that the body of man enters upon its career as a tiny germ or embryo, which
becomes an infant, quickened at a certain stage by the spirit whose tabernacle it is, and the
child, after being born, develops into a man. There is nothing in this, however, to indicate
that the original man, the first of our race, began life as anything less than a man, or less than
the human germ or embryo that becomes a man.

Man, by searching, cannot find out God. Never, unaided, will he discover the truth about the
beginning of human life. The Lord must reveal himself, or remain unrevealed; and the same
is true of the facts relating to the origin of Adam's race—God alone can reveal them. Some of
these facts, however, are already known, and what has been made known it is our duty to
receive and retain.

The Church of Jesus Christ of Latter-day Saints, basing its belief on divine revelation,
ancient and modern, proclaims man to be the direct and lineal offspring of Deity. God
himself is an exalted man, perfected, enthroned, and supreme. By his almighty power he
organized the earth, and all that it contains, from spirit and element, which exist co-eternally
with himself. He formed every plant that grows, and every animal that breathes, each after its
own kind, spiritually and temporally—"that which is spiritual being in the likeness of that
which is temporal, and that which is temporal in the likeness of that which is spiritual." He
made the tadpole and the ape, the lion and the elephant; but he did not make them in his own
image, nor endow them with Godlike reason and intelligence. Nevertheless, the whole animal
creation will be perfected and perpetuated in the hereafter, each class in its "distinct order or
sphere," and will enjoy "eternal felicity." That fact has been made plain in this dispensation.
(Doctrine and Covenants 77:3.)

Man is the child of God, formed in the divine image and endowed with divine attributes, and
even as the infant son of an earthly father and mother is capable in due time of becoming a
man, so the undeveloped offspring of celestial parentage is capable, by experience through
ages and aeons, of evolving into a God.

Joseph F. Smith

John R. Winder

Anthon H. Lund,

First Presidency of the Church of Jesus Christ of Latter-day Saints. 1

The word of the Lord in the Pearl of Great Price, the Book of Abraham, and the Doctrine and
Covenants, should carry enough weight with members of the Church to satisfy them and give
them a firm foundation on which to stand. To all the members who have received the
testimony through the Holy Ghost, these teachings supporting the Old Testament, will
suffice. We do have in the Church, however, a great many members who do not have that
abiding testimony, unfortunately. These are readily disturbed by the philosophies and
theories taught in the colleges and other schools and it is difficult for them to see that the
philosophical doctrines can be false. Many of the theories are proclaimed with such positivefinality that those weak in the faith are confused or perhaps inclined to accept the deductions
of these teachers and think that the revelations must be wrong. This is a step towards
apostasy. As the Lord declared, we cannot serve two masters. We cannot accept the
hypotheses of science which are in conflict with that which is here set forth in clearness, at
the same time. They are diametrically opposite to each other. Therefore it seems the part of
wisdom to present more testimony than is obtainable in the direct word of the Lord.
Therefore I shall proceed to present some of the writings of others of the members of the
Church who were schooled under the Prophet Joseph Smith. Here is a saying by President
Brigham Young:

It is a true saying of the Savior's . . . He came for the express purpose of dividing the
righteous from the wicked. This formed as much a part of his holy ministry as any other part
of the will of the Father.

We see this principle verified from days of old. It was demonstrated in the very
commencement of the peopling of the earth. How soon an opposition was introduced in the
morning of creation, when righteousness was proclaimed, when truth was revealed, when the
light and knowledge of eternity shone with lustrous beauty upon Adam and his children. Cain
must rise up and slay his brother while they were walking with the Lord. . . .

It is very true, had not sin entered into the world, and opposition been introduced, \textit{death
would not have entered.} (My italics.) From that time to this, death, opposition, selfishness,
malice, anger, pride, darkness of every description that could be invented by the children of
men, as they have multiplied and spread abroad in the earth, have increased. 2

How did Adam and Eve sin? Did they come out in direct opposition to God and to his
government? No. But they transgressed a command of the Lord, and through that
transgression sin came into the world. . . . Then came the curse upon the fruit, upon the
vegetables, and upon our mother earth; and it came upon the creeping things, upon the grain
in the field, the fish in the sea, and upon all things pertaining to this earth, through Man's
transgression. 3

\newpage
REFERENCES—CHAPTER SEVENTEEN

Footnotes

1. \textit{Improvement Era}, Vol. 13, pp. 75-81.

2. \textit{Journal of Discourses}, Vol. 1, pp. 234-5

3. \textit{Ibid.}, Vol. 10, p. 312.

