\chapter{PRE-EXISTENCE}

TO begin with the origin and destiny of man we must go back to the time before the
foundation of the earth was laid. Our evidence for this lies solely in the source of divine
revelation. For no man can remember his pre-existence. By the decree of our Eternal Father
all that we knew and all that we did in the spirit existence was taken from our memory. That
we did exist as spirits we have been informed by revelation. In that world we walked by
sight, for we were in the presence of God. The Lord has made it known that we have seen
him, 1 and there are passages in the Bible that infer this. Why our memories of the spirit
existence was blanked out, is that it was necessary in the mortal probation that we walk by
faith, and not by sight. It was decreed that we should be tried to see if we would keep the
commandments by faith and thus prove ourselves for a place of exaltation by obedience to
commandments the Lord would give us when we were not in his presence. However, we
were not sent into this world to walk blindly, for our Eternal Father sent angels to give us
commandments and make known the eternal plan of salvation. He also raised up prophets
unto whom he spoke and revealed his word from time to time. In the Meridian of Time, God
sent his Only Begotten Son to show mankind the way and he said: "I am the way, the truth,
and the life; no man cometh to the Father, but by me." In this remark he had reference to his
divine mission as Redeemer of the world and the Savior of all who believe on him and keep
his commandments. So mankind has not been left without necessary guidance, yet men have
their agency and may choose, as an ancient prophet (Alma) said: ". . . I know that he granteth
unto men according to their desire, whether it be unto death or unto life; yea, I know that he
allotteth unto men according to their wills, whether they be unto salvation or unto
destruction. Yea and I know that good and evil have come before all men; he that knoweth
not good from evil is blameless; but he that knoweth good and evil, to him it is given
according to his desires, whether he desireth good or evil, life or death, joy or remorse of
conscience." (Alma 29:4-5.)

The Bible reveals in several passages the preexistence of man in the spirit. The first reference
is in Genesis where we read that after all things were finished the Lord "saw every thing that
he had made, and behold, it was very good."

And on the seventh day God ended his work which he had made; and he rested on the
seventh day from all his work which he had made.

And God blessed the seventh day, and sanctified it: because that in it he had rested from all
his work which God created and made.

These are the generations of the heavens and of the earth when they were created, in the day
that the Lord God made the earth and the heavens.

All of this is said in relation to the creation of our physical earth on which we stand, and then
follows this statement which is not generally understood by Bible readers as follows:

And every plant of the field before it was in the earth, and every herb of the field before it
grew: for the Lord God had not caused it to rain upon the earth, and there was not a man to
till the ground. . . .

And the Lord God formed man [that is his physical body] of the dust of the ground, and
breathed into his nostrils the breath of life; and man became a living soul.

And the Lord God planted a garden eastward in Eden; and there he put the man whom he had
formed.

And out of the ground made the Lord to grow every tree that is pleasant to the sight, and
good for food; the tree of life also in the midst of the garden, and the tree of knowledge of
good and evil. 2

Now we are informed that the Lord formed every plant of the field and every herb of the field
\textit{before} they were placed in the earth. If we had what was originally written all of this would
be perfectly clear; and this original writing we do have as the Lord revealed it to the Prophet
Joseph Smith as Moses recorded it before scribes and translators altered it. Not only were the
plants and herbs formed before they were in the earth, but also man and every creature. Let
us read what was written in the beginning as it is given in the Book of Moses, in the Pearl of
Great Price:

"Thus the heaven and the earth were finished, and all the host of them." That is, the physical
creation was completed. Then the account explains, by way of interpolation, that nevertheless
all things had been created spiritually before this physical creation. But hosts of the earth
were not on the earth, although they had been created preparatory to coming to the earth.
Where were they, then, created preparatory to their sojourn on the earth? The record informs
us:

And on the seventh day I, God, ended my work, and all things which I had made; and I rested
on the seventh day from all my work, and all things which I had made were finished, and I,
God, saw that they were good;

And I, God,blessed the seventh day, and sanctified it; because that in it I had rested from all
my work which I, God, had created and made.

And now, behold, I say unto you, that these are the generations of the heaven and of earth,
when they were created, in the day that I, the Lord God, made the heaven and the earth;

And every plant of the field before it was in the earth, and every herb of the field before it
grew. For I, the Lord God, created all things, of which I have spoken, \textit{spiritually}, before they
were naturally upon the face of the earth. For I, the Lord God, had not caused it to rain upon
the face of the earth. And I, the Lord God, had created all the children of men; and not yet a
man to till the ground; \textit{for in heaven created I them}; and there was not yet flesh upon the
earth, neither in the water, neither in the air.

From this we learn that all the hosts of the heavens and the earth that were finished, (i.e.
created.) were created in the spirit and were in heaven where they remained until the earth
was prepared to receive them. From this we learn of the pre-existence, not only of plants andherbs, but of animals and mankind, but we will continue the Lord's account of this story of
the beginning:

And I, the Lord God, formed man from the dust of the ground, and breathed into his nostrils
the breath of life; and man became a living soul, the first flesh upon the earth, the first man
also; nevertheless, all things were before created; but \textit{spiritually} were they created and made
according to my word.

And I, the Lord God, planted a garden eastward in Eden, and there I put the man whom I had
formed.

And out of the ground made I, the Lord God, to grow every tree, naturally, that is pleasant to
the sight of man; and man could behold it. And it became also a living soul. For \textit{it was
spiritual} in the day that I created it; for it remaineth in the sphere in which I, God, created it,
yea, even all things which I prepared for the use of man; and man saw that it was good for
food. And I, the Lord God, planted the tree of life also in the midst of the garden, and also the
tree of knowledge of good and evil. 3

Here we learn of the creation of the spirit of man and of the beast and of the plants of the
earth in heaven preparatory to becoming \textit{souls}, for the soul, according to the definition the
Lord has given is the spirit and the body when joined together. (D. \& C. 88:15.) This account
is the original as it was given to Moses. Attention has been called to the fact that the ante-
diluvian patriarchs, from Adam to Noah, kept records. Moreover, that these records were
handed down and were in the hands of Abraham, and according to his statement it was his
intention to hand them down to his posterity. It is not beyond reason for us to believe that
those records reached the day of Moses. It is a mistaken notion that people could not read and
write in the days of Abraham and Moses, and even that the prophets before the flood did not
have this accomplishment. We have called attention to the fact that these things were well
known. We have learned through Abraham's writings that these ante-diluvians were
acquainted with the stars and planets.

The scoffer will scoff at all of this, but Latter-day Saints should have a personal
understanding that it is true. It is true that in our day the Lord made these things known to the
Prophet Joseph Smith. Do we not accept him as a Prophet? Is not that our faith? If not then,
pray tell, why is anyone who denies, or rejects it, a member of the Church? If this is your
belief, then humbly repent and go to the Lord and receive a personal testimony that these
things are true. That is your privilege. The Lord revealed to Nephi and it is recorded in the
Book of Mormon that forces would be at work to contaminate and corrupt the record of the
Jews, and that many of the plain and precious things would be taken from those records. 4
The time would come in the latter days when many of these plain and precious things would
be restored, and according to the revelation given to Joseph Smith as recorded in the Pearl of
Great Price, many of these things have been restored:

And now, Moses, my son, I will speak unto thee concerning this earth upon which thou
standest; and thou shalt write the things which I shall speak.

And in a day when the children of men shall esteem my words as naught and take many of
them from the book which thou shalt write, behold, I will raise up another like unto thee; and
they shall be had again among the children of men—\textit{among as many as shall believe.}

(These words were spoken unto Moses in the mount, the name of which shall not be known
among the children of men. And now they are spoken unto you. Show them not unto any
except them that believe. Even so. Amen.) 5

According to the promise the time has come for these things to be revealed. It is true that this
additional knowledge has been restored in the day of unbelief; when the scholars and critics
are endeavoring to tear apart and destroy the sacred writings of the prophets. It has come in a
day when the learned deny that Moses wrote the books which bear his name; the day when
the so-called "higher criticism," which is destructive criticism, has done all in its power to
discredit the prophets and destroy faith in the sacred records of the past, and the hypothesis
of organic evolution, in this regard has likewise done its part.

Not only do we have the restored writings of Moses on the spiritual creation but also the
testimony of Abraham. Here again the Church is blessed with knowledge in relation to the
origin of man and his pre-existence, which is not had by the world, notwithstanding the fact
that the Bible even as it has come down to us with many of the plain parts taken away from
the writings of the prophets, yet it contains an abundance of testimony dealing with the pre-
mortal state of man. Now let us examine the testimony Abraham:

And the Lord said unto me: These two facts do exist, that there are two spirits, one being
more intelligent than the other; there shall be another more intelligent than they; I am the
Lord thy God, I am more intelligent than they all.

The Lord thy God sent his angel to deliver thee from the hands of the priest of Elkenah.

I dwell in the midst of them all; I now, therefore, have come down unto thee to deliver unto
thee the works which my hands have made, wherein my wisdom excelleth them all, for I rule
in the heavens above, and in the earth beneath, in all wisdom and prudence, over all the
intelligences thine eyes have seen from the beginning; I came down in the beginning in the
midst of all the intelligences thou has seen.

Now the Lord had shown unto me, Abraham, the intelligences that were organized before the
world was; and among all these there were many of the noble and great ones;

And God saw these souls that they were good, and he stood in the midst of them,and he said:
These I will make my rulers; for he stood among those that were spirits, and he saw that they
were good; and he said unto me: Abraham, thou art one of them; thou wast chosen before
thou wast born.

And there stood one among them that was like unto God, and he said unto those who were
with him: We will go down, for there is space there, and we will take of these materials, and
we will make an earth whereon these may dwell;

And we will prove them herewith, to see if they will do all things whatsoever the Lord their
God shall command them;

And they who keep their first estate shall be added upon; and they who keep not their first
estate shall not have glory in the same kingdom with those who keep their first estate; and
they who keep their second estate shall have glory added upon their heads for ever and ever.
6

Those who profess to believe in the Bible, quite generally accept the fact that Jesus Christ
existed before the time he took upon himself a tabernacle of flesh and bones but for some
unaccountable reason they refuse to believe that man also had a pre-existence. It is
strange,however, that many of them believe that man is a soul composed of both spirit and
body, but that in some manner the spirit and the body became united at birth, and the spirit
did not exist before birth. They believe that the spirit leaves the body at death and that it is
eternal. The inconsistency in this they fail to understand. The spirit is not created at birth, but
is eternal. A more careful reading of the Bible would show that the statement in the Pearl of
Great Price which I have quoted must be true, for the Bible is replete with references to the
antemortal state of man. Here is some of the evidence.

Moses wrote: (Deut. 32:7-9.)

Remember the days of old, consider the years of many generations: ask thy father, and he
will shew thee; thy elders, and they will tell thee.

When the Most High divided to the nations their inheritance, when he separated the sons of
Adam, he set the bounds of the people according to the number of the children of Israel.

For the Lord's portion is his people; Jacob is the lot of his inheritance.

Here is a saying that when the inheritances of the nations were considered the Lord set their
bounds according to the number of the tribes of Israel. This, evidently, was done long before
there was an Israel, for Israel had not at the time this was written entered into his inheritance.
It must have been a decision before the people of the nations as well as Israel were on the
earth. Here are other Bible quotations confirming this doctrine:

Then the word of the Lord came unto me, saying,

Before I formed thee in the belly I knew thee; and before thou camest forth out of the womb I
sanctified thee, and I ordained thee a prophet unto the nations. 7

And as Jesus passed by, he saw a man which was blind his birth.

And his disciples asked him, saying, Master, who did sin, this man, or his parents, that he
was born blind?

Jesus answered, Neither hath this man sinned, nor his parents: but that the works of God
should be made manifest in him. 8

Furthermore we have had fathers of our flesh which corrected us, and we gave them
reverence: shall we not much rather be in subjection unto the Father of spirits, and live? 9

Other evidence of the pre-existence of spirits is found in the Bible. One reference which is so
interpreted is found in the Book of Job when the Lord placed that worthy sufferer under
further questioning wherein he said:

Gird up now thy loins like a man; for I will demand of thee, and answer thou me.

Where wast thou when I laid the foundations of the earth? declare, if thou hast
understanding.

Who hath laid the measures thereof, if thou knowest? or who hath stretched the line upon it?

Whereupon are the foundations thereof fastened? or who laid the corner stone thereof;

When the morning stars sang together, and all the sons of God shouted for joy. 10

These sons of God were the intelligences, or spirits, we read about in the Book of Abraham
and in other scriptures. These were the spirits who were waiting for the opportunity to come
to the earth and receive bodies of flesh and bones and pass through this mortal probation. It
was for this opportunity that they sang for joy. We learn more about this pre-existence in the
Book of Revelation and the writings of Peter, as well as in the Pearl of Great Price. In the
twelfth chapter of Revelation we have a description of the restoration of the Gospel and the
Priesthood in the days of Christ's ministry and how Satan, the dragon, made war upon the
Church and drove her into the wilderness and her son, the Priesthood, was taken back into
heaven unto God. This dragon drew with him a third part of the stars of heaven. The dragon
was driven out and took with him one third of the spirits who refused to accept the plan of
salvation and Jesus Christ as their Redeemer. "And there was war in heaven: Michael and his
angels fought against the dragon; and the dragon fought and his angels, And prevailed not;
neither was their place found any more in heaven. And the great dragon was cast out, that old
serpent, called the devil, and Satan, which deceiveth the whole world: he was cast out into
the earth, and his angels were cast out with him." (Rev. 12:3, 7, 8, 9.)

In a revelation to the Prophet Joseph Smith given in September 1830, at Fayette, New York,
the Lord revealed the rebellion of Lucifer and his ejection from heaven in the following
words:

Behold, I gave unto him [Adam] that he should be an agent unto himself; and I gave unto
him commandment, but no temporal commandment gave I unto him, for my commandments
are spiritual; they are not natural nor temporal, neither carnal nor sensual.

And it came to pass that Adam, being tempted of the devil—for, behold, the devil was before
Adam, for he rebelled against me, saying, Give me thine honor, which is my power; and also
a third part of the hosts of heaven turned he away from me because of their agency;

And they were thrust down, and thus came the devil and his angels;

And, behold, there is a place prepared for them from the beginning, which place is hell.

And it must needs be that the devil should tempt the children of men, or they could not be
agents unto themselves; for if they never should have bitter they could not know the sweet.
11

Father Lehi, when instructing his son Jacob, has also given the reason why the devil is here
on earth tempting man.

Wherefore, the Lord God gave unto man that he should act for himself. Wherefore, man
could not act for himself save it should be that he was enticed by the one or the other.

And I, Lehi, according to the things which I have read, must needs suppose that an angel of
God, according to that which is written, had fallen from heaven; wherefore, he became a
devil, having sought that which was evil before God.

And because he had fallen from heaven, and had become miserable forever, he sought also
the misery of all mankind. Wherefore, he said unto Eve, yea, even that old serpent, who is the
devil, who is the father of all lies, wherefore he said: Partake of the forbidden fruit, and ye
shall not die, but ye shall be as God, knowing good and evil. 12

The penalty inflicted upon these rebellious spirits is that they are denied the blessings of this
mortal life. They are known as sons of perdition because they are denied bodies and are
partakers of the second death, which is eternal banishment from the presence of the Lord.
Their mission in this world is to tempt mankind and try to get them to deny Jesus Christ and
reject the everlasting Gospel. Having great knowledge and power, for they had great
experience before their rebellion, they use all manner of cunning schemes to entice men from
the path of rectitude and away from the kingdom of God. In their cunning craftiness they
teach some truths, but never the whole truth. The devil is the author of false religions. He is
perfectly willing that men should worship something and in some manner. He makes them
think they are worshiping Jesus Christ and his Father but sees to it that many false doctrines
contrary to the plan of salvation are introduced among men. He is the author of confusion
and laughs at the divided condition existing among the religious denominations. He it was
who brought to pass the great apostasy from the religion and Church of Jesus Christ in
former days. Satan is exercising great power and has led the great majority of mankind away
from the commandments of God, even while he makes them think that they are serving him.
In these latter days he is extremely busy and dominates the thinking and the philosophies of
the world and has led many people into "strong delusion, that they should believe a lie." 13
He knows "that he hath but a short time." 14 Nephi saw our day and how Satan, knowing that
his days are numbered, would stir the people up to all manner of iniquity. He says, speaking
of the last days:

For the kingdom of the devil must shake, and they which belong to it must needs be stirred
up unto repentance, or the devil will grasp them with his everlasting chains, and they be
stirred up to anger, and perish;

For, behold, at that day shall he rage in the hearts of the children of men, and stir them up to
anger against that which is good.

And others will he pacify, and lull them away into carnal security, that they will say: All is
well in Zion; yea, Zion prospereth, all is well—and thus the devil cheateth their souls, and
leadeth them away carefully down to hell.

And behold, others he flattereth away, and telleth them there is no hell; and he saith unto
them: I am no devil, for there is none—and thus he whispereth in their ears, until he grasps
them with his awful chains, from whence there is no deliverance.

Yea, they are grasped with death, and hell; and death, and hell, and the devil, and all that
have been seized therewith must stand before the throne of God, and be judged according to
their works, from whence they must go into the place prepared for them, even a lake of fire
and brimstone, which is endless torment.

Therefore, wo be unto him that is at ease in Zion!

Wo unto him that crieth: All is well:

Yea, wo be unto him that hearkeneth unto the precepts of men, and denieth the power of
God, and the gift of the Holy Ghost!

Yea, wo be unto him that saith: We have received, and we need no more!

And in fine, wo unto all those who tremble, and are angry because of the truth of God! For
behold, he that is built upon the rock receiveth it with gladness; and he that is built upon a
sandy foundation trembleth lest he shall fall. 15

These evil spirits have great power to tempt, persuade and entice men to deny the correct
origin of man. We do not see them, but we do feel their presence, and unconsciously we
hearken to their whisperings. Having been denied bodies they, at times, steal them. It is a
common error, especially in scientific circles to scoff at such a thing as the temptation by the
devil and more especially so to ridicule the idea that these wicked spirits have power to
possess living bodies and subdue the spirit within them. But all the scoffing and ridicule does
not change the fact. The stories of possession as recorded in the New Testament are true. The
scoffer cannot explain away successfully the casting out of devils by Jesus Christ, when they
called him by name and he commanded them to hold their peace; the story of the devils
asking to enter the bodies of swine; the story of the seven sons of Sceva, and numerous
others listed in the scriptures. There are scores of such incidents that have occurred in this
dispensation. Our missionaries can give the evidence in such cases. No, it is not always a
diseased mind that disturbs the normal thinking, the possession by devils is a positive fact.

The following excerpts are taken from an editorial appearing first in the \textit{Times and Seasons}
and written by the Prophet Joseph Smith, having to do with spirits and their powers.

TRY THE SPIRITS

Recent occurrences that have transpired amongst us render it an imperative duty devolving
upon me to say something in relation to the spirits by which men are actuated.

It is evident from the Apostles' writings, that many false spirits existed in their day, and had
"gone forth into the world," 16 and that it needed intelligence which God alone could impart
to detect false spirits, and to prove what spirits were of God. The world in general have been
grossly ignorant in regard to this one thing, and why should they be otherwise? for "the
things of God knoweth no man, but the Spirit of God."

The Egyptians were not able to discover the difference between the miracles of Moses and
those of the magicians until they came to be tested together; and if Moses had not appeared
in their midst, they would unquestionably have thought that the miracles of the magicians
were performed through the mighty power of God, for they were great miracles that were
performed by them—a supernatural agency was developed and great power manifested. . . .

It would have been equally as difficult for us to tell by what spirit the Apostles prophesied, or
by what power the Apostles spoke and worked miracles. Who could have told whether the
power of Simon, the sorcerer, was of God or of the devil?

There always did, in every age, seem to be a lack of intelligence pertaining to this subject.
Spirits of all kinds have been manifested, in every age, and almost amongst all people. If we
go among the pagans, they have their spirits; the Mohammedans, the Jews, the Christians, the
Indians—all have their spirits, all have a supernatural agency, and all contend that their
spirits are of God. Who shall solve the mystery? "Try the spirits," says John, but who is to do
it? The learned, the eloquent, the philosopher, the sage, the divine—all are ignorant. The
heathens will boast of their gods, and of the great things that have been unfolded by their
oracles. The mussulman will boast of his Koran, and of the divine communications that his
progenitors have received. The Jews have had numerous instances, both ancient and modern,
among them of men who have professed to be inspired, and sent to bring about great events,
and the Christian world has not been slow in making up the number.

"Try the spirits," but what by? Are we to try them by the creeds of men? What preposterous
folly—what sheer ignorance—what madness! Try the motions and actions of an eternal being
(for I contend that all spirits are such) by a thing that was conceived in ignorance, and
brought forth in folly—a cobweb of yesterday! Angels would hide their faces, and devils
would be ashamed and insulted, and would say, "Paul we know, and Jesus we know, but who
are ye?" Let each man of society make a creed and try evil spirits by it, and the devil would
shake his sides; it is all that he would ask—all that he would desire. Yet many of them do
this, and hence "many spirits are abroad in the world."

One great evil is, that men are ignorant of the nature of spirits; their power, laws,
government, intelligence, etc., and imagine that when there is anything like power,
revelation, or vision manifested, that it must be of God. Hence the Methodists, Presbyterians,
and others frequently possess a spirit that will cause them to lie down, and during its
operation, animation is frequently entirely suspended; they consider it to be the power of
God, and a glorious manifestation from God—a manifestation of what? Is there any
intelligence communicated? Are the curtains of heaven withdrawn, or the purposes of God
developed? Have they seen and conversed with an angel—or have the glories of futurity
burst upon their view? No! but their body has been inanimate, the operation of their spirit
suspended, and all the intelligence that can be obtained from them when they arise, is a shout
of "glory," or "hallelujah," or some incoherent expression; but they have had "the power."

The Shaker will whirl around on his heel, impelled by a supernatural agency or spirit, and
think that he is governed by the Spirit of God; and the Jumper will jump and enter into all
kinds of extravagances. A Primitive Methodist will shout under the influence of that spirit,
until he will rend the heavens with his cries; while the Quakers (or Friends) moved as they
think, by the Spirit of God, will sit still and say nothing. Is God the author of all this? If not
of all of it, which does He recognize? Surely, such a heterogeneous mass of confusion never
can enter into the kingdom of heaven.

Every one of these professes to be competent to try his neighbor's spirit, but no one can try
his own, and what is the reason? Because they have not a key to unlock, no rule wherewith to
measure, and no criterion whereby they can test it. Could any one tell the length, breadth or
height of a building without a rule? Test the quality of metals without a criterion, or point out
the movements of the planetary systems, without a knowledge of astronomy? Certainly not;
and if such ignorance as this is manifested about a spirit of this kind, who can describe an
angel of light? If Satan should appear as one in glory, who can tell his color, his signs, his
appearance, his glory?—or what is the manner of his manifestation? Who can detect the
spirit of the French prophets with their revelations and their visions, and power of
manifestations? Or who can point out the spirit of the Irvingites, with their apostles and
prophets, and visions and tongues, and interpretations, etc. Or who can drag into daylight and
develop the hidden mysteries of the false spirits that so frequently are made manifest among
the Latter-day Saints? We answer that no man can do this without the Priesthood, and having
a knowledge of the laws by which spirits are governed; for as "no man knows the things of
God, but by the Spirit of God," so no man knows the spirit of the devil, and his power and
influence; but by possessing intelligence which is more than human, and having unfolded
through the medium of the Priesthood the mysterious operations of his devices. Without
knowing the angelic form of an angel the sanctified look and gesture, and the zeal that is
frequently manifested by him for the glory of God, together with the prophetic spirit, the
gracious influence, the godly appearance, and the holy garb, which are so characteristic of his
proceedings, his mysterious windings are not known.

A man must have the discerning of spirits before he can drag into daylight this hellish
influence and unfold it unto the world in all its soul-destroying, diabolical, and horrid colors;
for nothing is a greater injury to the children of men than to be under the influence of a false
spirit when they think they have the Spirit of God. Thousands have felt the influence of its
terrible power and baneful effects. Long pilgrimages have been undertaken, penances
endured, and pain, misery and ruin have followed in their train; nations have been convulsed,
kingdoms overthrown, provinces laid waste, and blood, carnage and desolation are
habiliments in which it has been clothed. . . .

As we have noticed before, the great difficulty lies in the ignorance of the nature of spirits, of
the laws by which they are governed, and the signs by which they may be known; if it
requires the Spirit of God to know the things of God; and the spirit of the devil can only be
unmasked through that medium, then it follows as a natural consequence that unless some
person or persons have a communication, or a revelation from God, unfolding to them the
operation of the spirit, they must eternally remain ignorant of these principles; for I contend
that if one man cannot understand these things but by the Spirit of God, ten thousand men
cannot; it is alike out of reach of the wisdom of the learned, the tongue of the eloquent, the
power of the mighty. And we shall at last have to come to this conclusion, whatever we may
think of revelation, that without it we can neither know nor understand anything of God, or
the devil; and however unwilling the world may be to acknowledge this principle, it is
evident from the multifarious creeds and notions concerning this matter that they understand
nothing of this principle, and it is equally as plain that without a divine communication they
must remain in ignorance. The world always mistook false prophets for true ones, and those
that were sent of God, they considered to be false prophets, and hence they stoned, punished,
imprisoned and killed the true prophets, and these had to hide themselves "in deserts and
dens, and caves of the earth," and though the most honorable men of the earth, they banished
them from their society as vagabonds, whilst they cherished, honored and supported knaves,
vagabonds, hypocrites, impostors, and the basest of men. 17

No one was ever better qualified to speak about spirits, and the discerning of spirits, than the
Prophet Joseph Smith. He had reason to understand the power of the devil and the influence
he exerts over the souls of men and also the sweet influence of the Spirit of the Lord. He does
not speak idly or without knowledge. Of course it is natural for the man with evolutionary
tendencies and trained in the modern schools of learning to ridicule the possession by devils,
or any manifestations from this evil source, for that is the philosophy which prevails today.
For instance, Andrew D. White, a man of great renown, in Chapter XI of the first volume of
his \textit{A History of the Warfare of Science with Theology in Christendom}, and in his second
volume Chapter XV, bitterly assails the doctrine of the influence of the devil and the
possession of devils as recorded in the scriptures. In this thinking he sets forth the
modernistic views which are accepted quite universally in the scientific world. But these
articles written in ridicule and bitterness, do not change the facts. They merely call our
attention more forcibly to the words of Paul:

But as it is written, Eye hath not seen, nor ear heard, neither have entered into the heart of
man, the things which God hath prepared for them that love him.

But God hath revealed them unto us by his Spirit: for the Spirit searcheth all things, yea, the
deep things of God.

For what man knoweth the things of a man, save the spirit of man which is in him? even so
the things of God knoweth no man, but the Spirit of God.

Now we have received, not the spirit of the world, but the spirit which is of God; that we
might know the things that are freely given to us of God.

Which things also we speak, not in the words which man's wisdom teacheth, but which the
Holy Ghost teacheth; comparing spiritual things with spiritual.

But the natural man receiveth not the things of the Spirit of God: for they are foolishness
unto him: neither can he know them, because they are spiritually discerned. 18

Here I wish to present in conclusion the incident of Satan's endeavor to stop the work in
England, when the Gospel in this dispensation was first proclaimed there. You may laugh at
it, all you Whites, Millikans, Drapers. Osborns and others who ridicule the scriptures and the
power of God, but all the ridicule and contempt displayed in the world cannot destroy this
testimony and the thousands of others which have been reported, but which it is not
necessary to present in abundance here.

Sunday, July 30th (1837), about daybreak. Elder Isaac Russell (who had been appointed to
preach on the obelisk in Preston Square, that day), who slept with Elder Richards in Wilfred
Street, came up to the third story, where Elder Hyde and myself [Heber C. Kimball] were
sleeping, and called out, "Brother Kimball, I want you should get up and pray for me that I
may be delivered from the evil spirits that are tormenting me to such a degree that I feel I
cannot live long, unless I obtain relief."

I had been sleeping on the back of the bed. I immediately arose, slipped off at the foot of the
bed, and passed around to where he was. Elder Hyde threw his feet out, and sat up in the bed,
and we laid hands on him, I being mouth, and prayed that the Lord would have mercy on
him, and rebuked the devil.

While thus engaged, I was struck with great force by some invisible power, and fell senseless
on the floor. The first thing I recollected was being supported by Elders Hyde and Richards,
who were praying for me; Elder Richards having followed Russell up to my room. Elders
Hyde and Richards then assisted me to get on the bed, but my agony was so great I could not
endure it, and I arose, bowed my knees and prayed. I then arose and sat up on the bed, when
a vision was opened to our minds, and we could distinctly see the evil spirits, who foamed
and gnashed their teeth at us. We gazed upon them an hour and a half (by Willard's watch).
We were not looking towards the window, but towards the wall. Space appeared before us,
and we saw the devils coming in legions, with their leaders, who came within a few feet of
us. They came towards us like armies rushing to battle. They appeared to be men of full
stature, possessing every form and feature of men in the flesh, who were angry and
desperate; and I shall never forget the vindictive malignity depicted on the countenances as
they looked me in the eye; and any attempt to paint the scene which then presented itself, or
portray their malice and enmity, would be vain. I perspired exceedingly, my clothes
becoming as wet as if I had been taken out of the river. I felt excessive pain, and was in the
greatest distress for some time. I cannot even look back on the scene without feelings of
horror; yet by it I learned the power of the adversary, his enmity against the servants of God,
and got some understanding of the invisible world. We distinctly heard those spirits talk and
express their wrath and hellish designs against us. However, the Lord delivered us from
them, and blessed us exceedingly that day.

Elder Orson Hyde's supplemental description of the fearful scene is as follows, taken from a
letter addressed to President Kimball:

Every circumstance that occurred at that scene of devils is just as fresh in my recollection at
this moment as it was at the moment of its occurrence, and will ever remain so. After you
were overcome by them and had fallen, their awful rush upon me with knives, threats,
imprecations and hellish grins, amply convinced me that they were no friends of mine. While
you were apparently senseless and lifeless on the floor and upon the bed (after we had laid
you there), I stood between you and the devils and fought them and contended with them
face to face, until they began to diminish in number and to retreat from the room. The last
imp that left turned round to me as he was going out and said, as if to apologize, and appease
my determined opposition to them, "I never said anything against you!" I replied to him thus:
"It matters not to me whether you have or have not; you are a liar from the beginning! In the
name of Jesus Christ, depart!" He immediately left, and the room was clear. That closed the
scene of devils for that time. 19

Elder Orson F. Whitney, who wrote the \textit{Life of Heber C. Kimball}, states that some time later
when this incident was called to the attention of the Prophet Joseph Smith by Elder Kimball,
he said: "Brother Heber, at that time you were nigh unto the Lord; there was only a veil
between you and him, but you could not see him." The Prophet then related some of his own
experiences in contests he had gone through with the evil power. How similar was the contest
that Moses had with Lucifer as recorded in the Book of Moses, 20 and as recorded in Jude.
21

\newpage
REFERENCES—CHAPTER THIRTEEN

Footnotes

1. D. \& C. 88:50.

2. Genesis 2:1-9.

3. Moses 3:1-9.

4. 1 Nephi 13:20-42.

5. Moses 1:40-42.

6. Abraham 3:19-26.

7. Jeremiah 1:4-5.

8. John 9:1-3.

9. Heb. 12:9.

10. Job 38:3-7.

11. D. \& C. 29:35-38.

12. 2 Nephi 2:16-18.

13. 2 Thess. 2:12.

14. Rev. 12:12.

15. 2 Nephi 28:19-28.

16. \textit{D. H. C.}, Vol. 4, pp. 571-574

17. 1 Cor. 2:9-14

18. Whitney, O. F., \textit{Life of Heber C. Kimball}, pp. 129-131.

19. Moses 1:12-26.

20. Jude 9.

21. 1 John 4:1; D. \& C. 46:7-8.

