\chapter{THE HYPOTHESIS OF ORGANIC EVOLUTION—2}

SIR AMBROSE FLEMING in the Foreword to his excellent work \textit{Evolution or Creation?}
speaking of the theory of organic evolution has said:

This theory, though an unverified speculation, is still so widely accepted as truth that anyone
venturing to question it is regarded either as ignorant or else bigoted.

Nevertheless, its teachings are in such opposition, especially as regards the origin of the
human race, with those of the Inspired Scriptures, that it is an unquestionable duty on the part
of those who still maintain faith in the Divine origin of the scriptures to examine carefully
how far this evolutionary theory is sufficiently or scientifically demonstrated. Though the
widest publicity is generally given in the Press to any utterances of eminent men on the
evolutionary side, yet little or no notice is taken of any arguments against Evolution, and
hence for the most part students hear only one side of the case.

He further says, that the theory "cannot explain the source of Matter, Life, or Mind," and
opposes all teleological arguments for their origin in a Supreme Intelligence and Creative
Power. Moreover, the theory "dries up the fountain from which alone any remedies can be
drawn to ease the struggles, sorrows, and sins of the present evil world, it offers nothing in
exchange on which mankind can build a 'lively hope,' for the future."

It insists strongly on man's kinship with the beasts, but it is powerless to give man the
smallest hope of any life beyond the grave, or any satisfaction for the deepest yearnings of
his soul. It appears to offer a solution of certain intellectual problems, but it does so at the
cost of destroying faculties in man more necessary for the continuance of his life. It invites
man apparently to partake of the fruit of the Tree of Knowledge, but in effect it debars him
from coming to the Tree of Life. 1

It was James who said, "Do not err, my beloved brethren. Every good gift and every perfect
gift is from above, and cometh down from the Father of lights, with whom is no variableness,
neither shadow of turning." 2 Sir Ambrose Fleming is right. The powers advocating
evolution are now sitting in our schools, colleges, scientific circles and controlling the press
to a great extent. Let no one think, however, that all the scientists and all the professors
teaching in colleges and editors of the press, have been contaminated by these foolish
doctrines. There are thousands of professors and highly educated men who denounce these
theories most vehemently, among them many eminent scholars who have made a thorough
study of anatomy and other branches of science. It is not by any means a universal doctrine
and has worthy opponents among those who have faith in God. Nor should we forget that the
doctrine of organic evolution is only an hypothesis—a guess. It can never be more than a
guess, and a bad guess at that, for the beginning of life on the earth has always come from
antecedent life as far back as the history of man goes. Many evolutionists acknowledge the
weakness of their cause and frankly admit that they are under the necessity of postulating,
(guessing,) without any evidence to support their claim, that there \textit{must} have been a time
when life came upon the earth in some minute form originally, for life is \textit{now} on the earth in
its multitudinous and varied forms. In their stubbornness and determination and varied forms.

In their stubbornness and determination to destroy faith in God as the Creator and in the
scriptures which definitely declare in simple understandable language that God is the Creator
of all things, they are forced to take this untenable position so they guess that once many
millions of years ago, life \textit{must} have come on the earth spontaneously. They have no proof,
they can discover no proof, and before any court where justice is dispensed and evidence is
required, their case would have to be thrown out of court. In all seriousness, their case has
been thrown out of court before the Just Judge who rules both on earth and in the heavens;
and the day is not far distant when the advocates of this pernicious doctrine will have to
answer for the countless souls they have blinded by their craftiness and turned away from
worshiping the Living God! There is, I repeat, no greater crime than the turning away of our
fellow men by the teachings of corrupt and malicious doctrines from the true worship of their
heavenly and Eternal Father, the maker of all things.

Dr. Douglas Marsland, in his \textit{Principles of Modern Biology} (Henry Holt and Co., N. Y.,
1945) states the case of evolutionists in the following paragraph, page 16:

Because of the incompleteness of our present knowledge, no one has yet been able to
produce protoplasm "artificially." Non-living matter is continually converted into living
matter, by the processes of nutrition in living organisms; and living matter is continually
reconverted into lifeless matter by metabolism and by the death of organisms. But so far as
we actually know, non-living matter is converted into living matter only in intimate
association with existing protoplasm—that is, in living organisms. In this present geological
age, living organisms, are known to originate only from other living organisms, by the
process of reproduction. \textit{It seems logically inevitable that living matter must have originated,
in the remote past, from non-living matter, without the intervention of previously living
matter, but we have no adequate evidence as to how, or under what conditions, this
occurred.} (My italics.)

To the man who has faith in God and believes in divine revelation, there is a perfect
knowledge how life began, but to those who reject the Divine Creator of necessity there \textit{must}
be a postulate of that kind, for the fact that life exists has to be accounted for even by the
ungodly.

William Hayes Ward, formerly editor of the New York Independent, in the issue of that
magazine May 11, 1914, dealing with the evolutionary theory of the origin of spontaneous
life, had this to say:

THE MYSTERY OF LIFE

In previous articles I have aimed to make it clear that our physical universe, whether looked
at in its minutest atoms or in its total starry system, gives clear evidence that it is not self-
existent, but had an external source. Nothing exists by its own necessity, and nothing by
chance. Some superior power is the source of physical matter and of physical laws. I now
turn to that other and higher world of life, and ask what evidence it has to offer as to its
origin. Do the familiar laws of chemistry and physics account for the first beginnings of life
and for its development in the vegetable and animal worlds? In this discussion simply vital
activities will be considered; the mental activities embraced in reason, instinct and will are
reserved for later treatment.

Living matter differs from inorganic matter in that it has a more complex structure, and in
that it grows under new laws. It is made out of a few of the same chemical atoms, but chiefly
of four of them, oxygen, hydrogen, carbon and nitrogen; but these appear in much more
intricate combinations than those dealt with in inorganic chemistry. . . .

***

Living matter also has the power of growth, not possest by inorganic matter. It is not growth
when a crystal of alum is enlarged by depositing layer on layer on the outside of it; but the
plant or the animal grows by taking food within itself, and then changing it into vitalized
matter. This requires new laws, while at the same time the physical laws continue in full
force. . . .

I am compelled to believe that there is something more in life than the mere forces of
chemistry and physics. Those forces can explain a star, but not a rose. The chemist and
physicist can follow and explain everything—how the sap rises under osmotic law, the
oxygenation of the blood, its traverse to and from the heart—everything except just one
thing, namely, what is the initial impulse that sets their familiar laws at work in a way so
different, so superior to anything that those laws can do apart from life. Life stops, and those
laws no longer in subjection act in their own free way, and the matter organized under life
disorganizes in decay. It is the guidance, the direction, so palpable to create a plant, a bird, a
man, which physics cannot explain.

LIFE IS PURPOSIVE

It is of the very essence of life that it gives guidance, is purposive. This separates it from
mere physical forces, such as the attraction of chemism. It has a previsioned end to achieve.
It aims to create a tree, a man, then to keep them repairing themselves or growing to an ideal
perfection. Out of the common say the atoms distribute themselves after a preconceived
scheme to organize into bark, wood, leaves, petals, stamens, pistils, seeds, just as we knew
they would when we planted the peach-stone. That is very purposeful life. Life chooses,
sorts, selects, directs, sees and reaches a distant aim. Whence comes this out-reaching,
selective, directive power.

PHYSICS CANNOT ACCOUNT FOR LIFE

The mere biologist does not try to answer this question. He is content to see it, to state its
laws and give names to the usual processes of life, and then he too often thinks that the
naming of the law is an explanation of its force. An apple falls to the ground. We ask why?
and we are told that the attraction of the earth draws it. ATTRACTION is a Latin word that
means drawing; and so we are told that \textit{drawing} draws it; and so we have got nowhere. We
have simply given a general name to a familiar fact; but the reason why the apple falls to the
earth we have not learned. So vitalism, or vital force, is but a name we give to an observed
order of processes, and, put into English, it means nothing more than life. It explains nothing.
Its marked character is its foresight. This prevision is everywhere, in the egg, in the chick, in
the bird, and no biologist can explain, he can only describe the process. The latest biologists
are coming to see that physics cannot account for life, which is a new and added directive
principle. Says the distinguished Dr. Anton Kerner in his \textit{Natural History of Plants}, as
quoted by A. R. Wallace:

"I do not hesitate to designate as 'vital force' this natural agency, not to be identified with any
other, whose immediate instruments is protoplasm, and whose peculiar effects we call life.
The atoms and molecules of protoplasm perform the functions which we call life only so
long as they are swayed by this vital principle. If its dominion ceases they yield to the
operation of other forces. The recognition of a special natural force of this kind is not
inconsistent with the fact that living bodies may at the same time be subject to other natural
forces."

Again he says, speaking of the wonderful processes connected with chlorophyll:

"What is altogether puzzling is, how the active forces work, how the sun's rays are able to
bring it about that the atoms of the raw material abandon their previous grouping, become
displaced, intermix one with another, and shortly reappear in stable combinations under a
wholly different arrangement. It is the more difficult to gain a clear idea of these processes
because it is not a question of that displacement of atoms called decomposition, but as to that
process which is known as combination, or synthesis."

IS VOLITION LODGED IN MATTER?

This directive and selective force which we call life appears to be outside of and above the
laws of inorganic nature. Physical nature has no such power. We know molecules drawn
together into geometrical forms under mechanical forces which we do in a measure
understand. But in those forms there is no such synthesis. We cannot imagine such blind and
purposeless forces performing such purposeful combinations as are necessary to restore the
lost leg of a lizard, or to create buds and send out suckers from the spot where the bark of the
tree is bruised. \textit{Haeckel} saw the difficulty and tried to explain it in a meaningless way. He
postulated will in the form of an unconscious directive force lodged in every atom, its
unconscious soul. But that is so utterly void of evidence and so utterly contradicts the
universal sense of the race that we must dismiss it. It is easier, instead of distributing an
imaginary rudimentary mind to all the atoms of the earth and of all worlds, it is far easier to
conceive of a really intelligent Mind that guides and directs the purposeful forces and
selective movements in all the forms of growth and life.

Then in a second article June 8, 1914, he continued the subject and wrote the following :

WHERE THE LAW OF CHANCE FAILS

Not only does the order of the appearance on the earth of the successive forms of life suggest
a forward anticipatory look and purpose, but we seem to observe the same thing when we
consider the production of the parts and organs of the living body. The old argument for
creationism drawn from the eye treated it simply as a mechanism, a wonderfully complicated
and accurate mechanism, something far beyond what human intelligence could have planned,
and it asked whether it must not have had an omniscient Contriver. But evolution replied that
sensitiveness to light began in the formless \textit{amoeba}, which has no differentiated nervous
system whatever, that in the course of division and reproduction a certain portion of the
structure became somewhat sensitive to light, and that there was produced in the infusorian a
pigment spot which was more sensitive than other parts. Then by slow degrees, through
accidental favorable modifications of many generations, one improvement after another
happened to be added, until at last we have the eye of the vertebrates, with all its marvelouslyaccurate complexity of adaptation for the purpose of vision. But does not this put too much
on the unpurposed action of evolution? The eye is an instrument composed of parts
coordinated to each other. No one is of any advantage without all the others. The retina needs
a crystalline lens to focus a picture upon it. The appearance of an imperfect lump of stiffer
transparent fluid, the beginning of a crystalline lens, may be conceived to be of some
advantage; but not unless at the same time, and in the same individual, there were a
corresponding improvement in the constitution of the retina with its rods and cones fitted to
receive and define the very imperfect image cast by the gelatinous lump not yet a crystalline
lens. Every improvement in the lens requires in the same individual a parallel improvement
in the retina. The two must coincide to be of any added advantage and be transmitted. But
there is no likelihood that they will coincide by any \textit{happy} accident. Just so with the other
parts of the eye, the aqueous humor, the cornea, the iris; the evolution must be progressive,
representing coordinate changes in all the parts, each following the other, for any one change
in a single part must be met by changes in all the other parts; otherwise there will be
confusion rather than improved vision. This coordination is not to be expected in a single
individual. Under the law of chances that is too much to ask. If the changes do occur
simultaneously by a succession of those leaps which is called mutation, that makes it all the
more evident that some guiding hand has directed it. The appearance is of design, a
prearranged evolution of the eye.

The dividing chasms separating the inert inorganic elements, the animal world and the human
family, have never been bridged and no evidence worthy of consideration has ever been
produced to uphold contrary theories in all the toilsome fruitless research that has been made.
Since the beginning of history, and knowledge does not go farther back than that, no
inorganic substance has ever taken upon itself animation and developed into life. Every clod
of earth, every stone or other substance belonging to the inorganic elements has remained the
same and without some animated influence acting upon it would remain the same ever
devoid of life and animated being.

\textit{The Lord placed a line of demarcation between the animal creation and the human family in
the very beginning, before the foundations of this earth were laid. In fact there is an eternal
decree that animals of different families, or species, shall remain separate from other
species, and there are bounds they cannot pass.}

And God said, Let the earth bring forth the living creature \textit{after his kind,} cattle, and creeping
thing, and beast of the earth \textit{after his kind:} and it was so.

And God made the beast of the earth \textit{after his kind,} and cattle \tetxit{after their kind,} and everything
that creepeth upon the earth \textit{after his kind:} And God saw that it was good.

And God said, Let us make man in \textit{our image, after our likeness:} and let them have dominion
over the fish of the sea, and over the fowl of the air, and over the cattle, and over all the
earth, and over every creeping thing that creepeth upon the earth.

And God created man in his own image, In the image of God created he him; \textit{male and
female} created he them. 3 (Italics mine.)

The experiences on the part of man who has violated this commandment and crossbred
animals of different families, has met with universal failure. Whenever animals have broughtforth their offspring and these offspring have mated successfully with other animals, it has
been invariably with those from the same family tree. The decree of the Lord has been
violated by man, as in the case of the ass and the horse, but the result in such cases is for the
posterity to come to an end. A divine law has been violated. This is one of the stumbling
blocks that stands like a mountain that cannot be crossed in the way of the progress of the
Darwinian theory.

Commenting on this decree of the Almighty, that all creatures, including man, should be
confined to their own order of beings, the Prophet Joseph Smith has said:

\textit{God has made certain decrees which are fixed and immovable; for instance—God set the
sun, the moon and the stars in the heavens, and gave them their laws conditions and bounds,
which they cannot pass, except by his commandments; they all move in perfect harmony in
their sphere and order, and are as lights, wonders, and signs unto us. The sea also has its
bounds which it cannot pass. God has set many signs on the earth, as well as in the heavens;
for instance, the oak of the forest, the fruit of the tree, the herb of the field—all bear a sign
that seed hath been planted there; for it is a decree of the Lord that every tree, plant, and
herb bearing seed should bring forth of its kind, and cannot come forth after any other law or
principle.} 4 (My italics.)

With this the Book of Abraham fully agrees and states this fact in very positive terms as
follows:

And the Gods said: Let us prepare the earth to bring forth grass; the herb yielding seed; the
fruit tree yielding fruit, after his kind, whose seed in itself yieldeth its own likeness upon the
earth; and it was so, even as they ordered.

And the Gods organized the earth to bring forth grass from its own seed, and the herb to
bring forth herb from its own seed, yielding seed after his kind; and the earth to bring forth
the tree from its own seed, yielding fruit, \textit{whose seed could only bring forth the same in itself,
after his kind;} and the Gods saw that they were obeyed. 5

In Section 88 of the Doctrine and Covenants is found this self-evident and indisputable truth
which refutes all contradiction:

All kingdoms have a law given:

And there are many kingdoms; for there is no space in the which there is no kingdom; and
there is no kingdom in which there is no space, either a greater or a lesser kingdom.

And unto every kingdom is given a law; and unto every law there are certain bounds also and
conditions.

All beings who abide not in those conditions are not justified. 6

This revelation was given long before the scientific discovery of the complicated atom with
its electrons, protons and neutrons. It was far in advance of scientific discovery, yet how true
this has proved to be! All of the laws of nature are the laws of God whether they apply to the
universe as a whole or in any of its parts. It is true of the electron as well as the atom or thecombination of atoms in any of their structures. They are all subject to and controlled by law.
These laws are eternal. No man can change a law of nature. If he tries he is not justified.
Every plant from the lowly fungus to the mighty sequoia is subject to divine law. The course
of animals as well as plants is fixed. Nature bears ample evidence that the law decreed to all
living things, that they shall produce after their own kind is a divine law and wherein at any
time this law has been violated the violators have not been justified.

Commenting on this eternal law, Byron Nelson, in his excellent work, \textit{After Its Kind,} has
commented as follows after quoting Genesis 1:24-25.

The Bible is not a textbook of science. In the first chapter of Genesis, however, because it is
a matter of the greatest religious importance, the Bible speaks clearly and finally on a matter
of biology. \textit{After its kind} is the statement of a biological principle that no human observation
has ever known to fail. The most ancient human records engraved on stone or painted on the
walls of caves bear wtiness to the fact that horses have ever been horses, dogs have ever been
dogs, pigeons have ever been pigeons, elephants have ever been elephants. The most
desperate and subtle efforts of man in modern times have been unable to alter this divine
decree.

The Bible teaches that from the beginning there have been a number of types of living things,
man included, which were so created as to remain true to their peculiar type throughout all
generations. These types or kinds may be fittingly described as \textit{species.} But here a word of
caution is necessary. Among biologists there has never been any agreement as to what a
species is. It has been generally considered that any particular form of plant or animal that
possesses marked characteristics of its own and breeds true to form is a species. For example,
the fox-terrier is called a species, because it is able to produce offspring like itself. The
dachshund, the collie, the police-dog are called species, because they are able to produce
their own particular forms. In this way the human race has been divided into several species
according to the shape of the head, the color of the skin, the slant of the eye. But such species
are not what the Bible means by the word, "Kind." The Bible does not mean to say that every
distinct form of plant or animal men see about them came from the hand of the Creator in just
the form in which it is beheld. It is not the several types of dogs; fox-terrier, dachshund, that
were created to remain the same forever, but the one natural species, \textit{dog.} The "kinds" of
Genesis refer not to the "systematic" species identified by men, but to those natural species of
which the world is full, which have power to vary within themselves in such a way that the
members of the species are not all exactly alike, but which, nevertheless, cannot go out of the
bounds that the Creator set. 7

Professor William Bateson, president of the British Association for the Advancement of
Science 1914-1927, has given a definition stating that a species is a group of organisms with
marked characteristics in common and freely interbreeding. It is not denied by anti-
evolutionists, that within a family, or species of animals or plants according to this definition
we will find numerous varieties. For instance there are many varieties of roses which have
been cultivated. Some are white, some pink, some yellow and others of varied hues. Some
have many petals, some but a few, but the fact remains they are \textit{all} roses. There is no
question that the common cabbage which we obtain in grocery stores, is akin to brussel
sprouts and other similar plants, but they are of the same plant family. Field corn, sweet corn,
popcorn, are varieties that have been cultivated, but nevertheless they are still corn and will
"interbreed." Corn and carrots do not mix, neither do squash and turnips. There are variousbreeds of dogs, but they do not breed with cats. The cat family, composed of the domestic
animal and the wild varieties, may mix. The horse and the ass are not of the same family and
while man has been able to obtain from them the mule, the mule is rudely and humorously
spoken of as being "without pride of ancestry and hope of posterity." The Lord decreed that
they should not mix.

This determining factor is a sufficient answer to organic evolution. Similarity of skeletal
structure has absolutely no significance as a determining factor in relationship of animals and
man. In the American Museum of Natural History, is to be seen a skeleton of a man
compared with the skeleton of a horse. The horse is standing on its hind feet, in an artificial
position. Dr. Osborn, the great champion of evolution and enemy of divine creation, in the
guide leaflet No. 36, published April, 1921, on page 41, calling attention to the similarity of
structure says, "A careful study will reveal a most striking similarity between horse and man
in general structure, the difference being simply modification of a common plan." So the
work goes on to deceive. Why did not this eminent professor point out the fact that the horse
has hoofs, that it does not stand erect and walk on two feet? Why did he not point out the fact
that the head of the man and the head of the horse resemble each other only in the fact that
each is composed of bones? The leaflet states (p. 9), "The horse may be said to walk on its
middle finger nail, all the other fingers have disappeared."

Paul in his discourse on Mars Hill to the unbelieving Greeks informed them that "God that
made the world and all things therein, seeing that he is Lord of heaven and earth, dwelleth
not in temples made with hands; . . . And hath made of one blood all nations of men for to
dwell on all the face of the earth, and hath determined the times before appointed, and the
bounds of their habitation." 8 In the beginning the Lord did not make man black, yellow and
brown. This condition, like many characteristics found among the nations, the Lord added
later. We know how part of Adam's race became black, this we Latter-day Saints have
learned through revelation. We likewise, know why the Indian (Lamanite) is reddish-brown,
for the Book of Mormon reveals it, and for similar reasons other races have acquired their
color and national characteristics. But this we know that in the beginning they were of one
family, white and delightsome. The man who depends on his science alone cannot
understand this, for these are things discerned through the revelations of the Lord. So white,
black or brown, tall or short, mankind now divided into nations, had a common ancestry. So
it is with the animal kingdom, the inhabitants of the mighty deep and the fowls of the air.

Alfred Watterson McCann commenting on this absurd comparison, which has but one
purpose and that is to deceive the ignorant and please those who aim to destroy belief in the
scriptures, after quoting the first two sentences from the leaflet, made this pertinent remark:

The horse began 3,000,000 years ago in the Eocene. There are twelve main links connecting
him with that remote epoch. The writer has examined the "proof" only to come away, not as
so much submissive mutton but as an unbeliever and a scoffer. With an experience covering
207 prosecutions in criminal courts, supreme courts, federal district courts and all sorts of
courts the writer has never seen such proof enter the records. Not only does it (bald opinion)
never reach cross-examination, but the judges strike it out instantly upon its appearance in
direct examination.

They want no opinions of any kind. They want evidence, facts, proof, not "testimony," and
they want corroboration. 9

According to this ridiculous story the horse has developed from a small animal no larger than
a rat through the many geological ages; from a small five-toed creature to the magnificent
animal which it is today. There is to be found in the publications coming from this wonderful
museum of deception a diagram showing the geological ages and how this tiny creature
developed and gradually lost its toes which were replaced by the four horse's hoofs. The
diagram is purely artificial and the story is purely fictitious, but it serves its purpose well in
deceiving innocent children and gullible people and makes them believe in the doctrines of
organic evolution. The diagram begins with the poor little \textit{Hyracotherium} the size of a rat or
a chipmunk. A skeleton of this maligned little animal in its entirety has never been found. All
they have found of the Hyracotherium is a skull, then how do they know that it had five toes?
Well here is the manuactured story of the horse in its several stages of development and how
through the great ages it lost its toes and had them replaced by a solid hoof on each of the
four feet.

Link No. One.—Hyracotherium, a harmless little animal. No one knows just how it looked or
whether it had four feet or two, as they have never been seen. Yet the experts give it four feet
with twenty toes.

Link No. Two.—This is called Eohippus. The manufactured instruction says: "The Eohippus
is much better known and is very like the hyracotherium. The forefoot has four complete
toes. The hind foot three," and it is declared that these two animals are "in the direct line of
descent of the modern horse." (p. 15.) This deduction is made from a few scattered bones of
an animal smaller than a squirrel.

Link No. Three.—This is called Protorohippus and was supposed to have lived one million
years later. By this time it had increased in size and was larger than a tomcat, but he still had
his three and four toes.

Link No. Four.—This is Mesohippus, and nearly a million and a half years have passed, and
the unfortunate, or fortunate thing, as you may view it, is that the poor unstable fellow had
evened up his toes.

Link No. Five.—Protohippus was still holding on to his three toes, but they were becoming
useless and the middle one had grown longer than those on the sides.

Link No. Six.—By this time he had reached the age of man, the Pliotocene in geological
time, and by now the poor forgotten Hyracotherium had grown in size to that of a real horse
and was able to walk on his middle toe which had turned into a hoof.

Believe it or not! This is the rubbish palmed off on the innocent and promulgated by the
cunning, crafty, designing men bent on destroying the Bible order of creation. The fact,
however, remains that this is pure unadulterated speculation, manufactured out of a few
scattered bones of various animals which may have lived and evidently did
contemporaneously upon the earth when the real horse roamed the prairies.

It should be remembered when examining such a diagram that the substance presented is
simply an artificial table. The geological formations on the earth are not found in this
uniformity. The age of the rocks have been dogmatically arranged in certain orders of age,
according to the fossils which they contain. The whole scheme is man-made and is
speculative. Dr. Henry H. Howorth, in the preface to his great work, \textit{The Mammoth and the
Flood,} makes this rather caustic remark in relation to the hypotheses of geology:

The human mind cannot rest without hypotheses of some kind. The collection of facts is
always followed by the formulating of laws, and to reach a rational theory about things is in
fact the aim of science. In early days it was the fashion, a fashion singularly illustrated by the
methods of Greek inquiry, to first formulate a law and then to bring the facts within it. In
modern times it has been found more fruitful to arrange our facts first and then to see what
law underlies them. The former process is considered obsolete in most sciences, not so in
Geology.

With many geologists, wide-reaching a \textit{priori} theory is first accepted on the metaphysical
grounds of its simplicity and plausibility. This is made a shibboleth to test every man's
scientific capacity by, and then all the facts are expected to fit into it, and if they will not fit
by a natural process they are forced to do so by paradoxes and subtleties which if presented
by one of Hume's "Common rabble" outside the pale of scientific orthodoxy, would be
denounced as trivial and puerile. Such a theory is the scholastic one of uniformity. That
nature works uniformly is as much a dogmatic creed with a large proportion of English
geologists as that it "abhors a vacuum" was with the Schoolmen, and because it was
propounded by a very ingenious, laborious, and eloquent writer, the Gamaliel at whose feet
many of us have learned much of what we know, Sir Charles Lyall, it has acquired the same
sanctity and authority among geologists as it would have done among Roman theologians if
it had been inserted in the Syllabus. This, in the sense in which it is urged, I believe to be a
complete mistake, and one which has done infinite harm to geological reasoning for many
years past, and it is not unnatural that a revolt against it is spreading among those whose
judgment cannot be questioned. 10

The deductions made by organic evolutionists are based on the dogmatic claims of geologists
of uniformity. These claims are not maintained as universally today as they were in the days
of William ("Strata") Smith, Georges Cuvier and Sir Charles Lyall. Uniformity, be it
remembered, is still but a theory! Yet evolutionists hold to it tenaciously. Let us now
continue the discussion of the origin of the horse and other animals which play such an
important part in the drama of organic evolution. I now refer to a scientific treatise by
Douglas Dewar, one of Britain's outstanding zoologists. The following excerpts are taken
from a lecture given by this gentleman before the \textit{Victoria Institute,} or \textit{Philosophical Society
of Great Britain,} April 17, 1950, the subject being \textit{Genetics and Evolution.}

Hundreds of geneticists are now at work, and the reports of their experiments and those of
their predecessors fill many volumes. Thus the question arises: are the results of these
experiments favorable or unfavorable to the theory of evolution? In my view they are most
unfavorable. As many biologists disagree with me, let me set forth briefly the grounds upon
which my opinion is based:

1. The experimental work of geneticists and of practical breeders shows that the species are
very stable and resistant to attempts to transform them, despite the phenomenon of variation.

Practical breeders have been handling our domestic animals for centuries, but no
fundamental change has been made in any kind of animal. In the case of some of thedomestic animals,
notably horses, cats and dogs, we have pictorial evidence that during the
past six millennia they have undergone very little modification.

In the case of horses no less an authority than Lady Wentworth declares:

"The present species has walked on single hoofs and shown the same structure as far back as
history can trace him." (\textit{Horses of Britain}, 1944.) "Further, early cave and rock pictures show
that in the neolithic period both the heavy-boned northern type of horse and the lighter
southern type existed in Europe. The former exhibited one large and two small varieties,
while the southern type is depicted in European rock pictures as a speckled pony: the pure
Arabian appears only in the rock paintings of Arabia (where it is often depicted as galloping
with a rider carrying a spear) and of Egypt (1800 B.C.), where it is shown both ridden and
driven." (\textit{The Ancient Arab Horse and His Descendants} (1945.)—\textit{Transactions,} p. 154.

Similarly ancient pictorial representations show that in ancient Egypt, fully 6000 years ago,
several breeds of domestic dog existed, one of which, of greyhound type, was used for
hunting deer, and another breed had short legs like a dachshund, a third had pendant ears.

We know from the pictures that the oldest domestic animals were asses, oxen, sheep, goats,
pigs, dogs, cats, geese and ducks.

The earliest known pictures of domestic animals show that none of them has changed much,
each domesticated species was 4,000 years ago as sharply marked off from all other kinds of
animals as it is today.

It is true that in the case of animals bred for amusement rather than utility many freaks have
been produced by man. Darwin made much of this, asserting that if some of the pigeons bred
by fanciers had been found in the wild they would have been deemed new species or even
genera, and he argued that if man in a few centuries can produce by selection such forms
natural selection working during millions of years could have effected vastly greater changes.

Darwin, however, knew nothing of the effect on the body of the secretions of the ductless
glands, and shut his eyes to the fact that these freaks are quite incapable of maintaining
themselves in nature, monstrosities resulting from gland unbalance; yet, despite their
abnormality, they clearly bear the stamp of the wild ancestor. Thus fantails, pouters, jacobins,
barbs, tumblers, swallows, trumpeters, etc., all bear the hallmark "pigeon." Moreover, all
these breeds, when crossed or when mated with the parent form, yield fertile offspring.

The work of geneticists confirms that of the practical breeders. In the animals on which the
former have experimented they have produced many freaks and monstrosities (some of these
will be noticed later), but, as in the case of the domestic animal freaks, these are all clearly
members of the wild species from which they have been bred. It is true that geneticists have
been at work for less than fifty years, but in most cases they have experimented on animals
which in the laboratory produce a number of successive broods in a year. Thus in the case of
the little fruit fly \textit{Drosophila Melanogaster,} on which the majority of geneticists have
worked, 25 successive generations can be reared in a year, so that some 1,000 generations
have been bred in the laboratory. Assuming that the generation time for man is twenty-five
years on an average, or 40 generations in 1,000 years, it would require 25,000 years to
perform this experiment on man. Nor is this all, Muller discovered in 1927, that byirradiating this fly with X-rays the rate at which mutations occur is increased fifteen thousand
per cent. Needless to say, these flies and other creatures experimented on have been freely X-
rayed during the past 20 years. In consequence the number of mutations which have been
produced in the laboratory has been vastly increased. \textit{But the mutations so produced are all
of the same kind as those which occur in untreated individuals.} This is true of the mutations
induced by other rays and mutation-inducing chemicals, such as mustard gas. \textit{Transactions}
April 17, 1950, pp. 154-155. (My italics.)

As a little reflection should render it clear that if neither the effects of use and disuse nor
acquired characters are inherited, the theory of evolution is impossible, it is surprising that
the majority of geneticists in English-speaking countries seem still to accept the evolutionary
theory. The attitude of these is most illogical. The French seem to realize this and in
consequence few of their biologists are geneticists. It is significant that the article on genetics
in the French \textit{Encyclopedia} is by E. Guyenot, a professor of the University of Geneva. Dr. A.
Labbe, a professor at the School of Medicine at Nantes, and an ardent transformist, writes:

"Genetics, which is consecrated to the study of heredity, has become a kind of religion,
dogmatic, mystical, intolerant, which has its temples, its priests, its believers, its councils,
and which aims at converting all the biologists in the world. For it transformism may still
exist in theory, but in practice the very fact of transformism is incomprehensible. However,
the geneticists still call themselves transformists; just as in politics where the left and the
right parties each claim exclusively the epithet republican. Without being deliberately
opposed to these genetical ideas, nevertheless I cannot accept them without many
reservations, and in common with most French biologists, I cannot admit even the foundation
of genetics other than as a possible, but unproved entity. Genetics ends inevitably in a more
or less complete negation of evolution: at the most it can conceive of fortuitous variations. . .
. We do not want this genetics which hampers us. . . . It is only when the laws of the
transformation of species will be better known that we can attack the problem of heredity.
Let us then set aside genetics which leads us either to the strict fixity of species or a relative
variation which is not evolutionary." (\textit{Le Conflict Transformiste} (1929), p. 140.)

To the logical biologist, there are only two alternatives, either to reject evolution, or to fly in
the face of genetical evidence and believe that acquired characteristics are inherited.

Not many biologists accept the first alternative. One of the few who do is Heribert Nilsson,
of Lund University, who is a botanical geneticist. He writes: "It is obvious that the
investigations of the last three decades into the problem of the origin of species have not been
able to show that a variational material capable of competition in the struggle for existence is
formed by mutation. Further, as it has also been impossible to demonstrate a progressive
adaptation by means of the transmission of acquired characters (all the numerous
experiments made have yielded negative results), we are forced to this conclusion that \textit{the
theory of evolution has not been verified by experimental investigations of the origin of
species.}"

He continues: "Is then biology without evolution conceivable?" He replies: "Just as affinity
in Chemistry or Mineralogy need not be based on the assumption that the elements evolved
from one another, from Hydrogen to Uranium, there is no more need of our basing the related
series of biology on an evolution from amoeba to \textit{Homo} and so on." (\textit{Hereditas,} Vol. XX
(1935), page 236.) 11

Dr. Richard Goldschmidt, Professor of Zoology at the University of California in his book,
\textit{The Material Basis of Evolution,} speaking of the so-called gene mutations caused by artificial
means cannot account for the origin of new species. Concluding the first part of his book he
says: "Micro-evolution by accumulation of micromutations—is a process which leads to
diversification strictly within the species. . . . \textit{Sub-species are actually, therefore, neither
incipient species nor models for the origin of species.} [His italics.] They are more or less
diversified blind alleys within the species. The decisive step in evolution, the first step
towards microevolution, the step from one species to another, requires another evolutionary
method than that of sheer accumulation of micromutations." 12

After all this experimentation we discover that the poor little fruit fly \textit{Drosophila}
Melanogaster, with all its mutations is still a fruit fly, notwithstanding it has gone through
what would be equal to 25,000 years of experimentation if performed on man. 13

A few more words about the horse. One of the criticisms used against the authenticity of the
Book of Mormon one hundred years ago was its declaration that the Nephites found on this
continent many useful animals. The passage is as follows:

And it came to pass that we did find upon the land of promise, as we journeyed in the
wilderness, that there were beasts in the forests of every kind, both the cow and the ox, and
the ass and the horse, and the goat and the wild goat, and all manner of wild animals, which
were for the use of man. And we did find all manner of ore, both of gold, and of silver and of
copper. 14

Again:

And it came to pass that the people of Nephi did till the land, and raise all manner of grain,
and of fruit, and flocks of herds, and flocks of all manner of cattle of every kind, and goats,
and wild goats, and also many horses. 15

This argument, like all other criticisms against the Book of Mormon was soon discarded of
necessity, because the scientists in their researches have found the skeletal remains of horses;
not only of horses but of other animals spoken of in the Book of Mormon which, formerly,
were thought to have been confined to the eastern hemisphere. Of some of these we will
speak later, but now we will consider the horse. It was the firm conviction of President
Anthony W. Ivins that the horse was not extinct when the Europeans first came to this
country. I have heard him give his reason based on what was, to him, sufficient evidence. Be
that as it may, the fact remains that the horse was here in primitive times. This truth is now
revealed in the skeletons which have been found. In the La Brea tar pools at Los Angeles
were buried the bodies of human beings and many kinds of birds and animals. Among these
have been found the bones of horses, saber-tooth tigers, giant sloths, wolves, lions, elephants,
and many other animals that, according to former notions, had no business in this western
world. But they were here to the amazement of many scientists who have never yet fathomed
the problem as to how their presence came about. Some of the nations set forth the idea that
the theory of evolution was working in a similar way on both hemispheres. Otherwise just
how did they get across the oceans? Then again, it has been stated that perhaps they started
here, but the same query presents itself, if so how did they get to the other side?

It has been my privilege to visit the La Brea Museum, and I remember how the guide pointed
out the difference existing between the fossil horse taken from the pit and the modern horse.
Their skeletal structures stood side by side. I recall that the chief difference was in relation to
the jaw and teeth, and the guide spent some time discoursing on this difference. One day later
in conversation with Antoine R. Ivins, who is a great lover of horses, he said: "If the guide
will come to the Ivins ranch at Enterprise, Utah, I will show him horses with identical
markings as the fossil horse possessed." The difference was extremely slight, no more than
we see in the modern horses of different breeds.

One day when I was attending a stake conference in Reno, I was invited to visit the Nevada
State Penitentiary to see the markings of ancient animals which are imbedded in the hard
surface within the prison walls. Here I saw the prints of the feet of the elephant, the giant
sloth, birds and the horse. The imprint of the horse made in geological times, was just as
perfect as would be the print of the hoof of a modern horse. The same was true of the other
animals and birds.

The Book of Ether gives us the clue to the reason why these animals were in the western
world. The Lord commanded the Jaredites to gather all manner of useful seeds and animals,
fish and fowl, and deseret, or the honey bee, and carry them from the Mesopotamia, from
whence they started their journey, to the land of promise "choice above all other lands, where
they were to be established and to multiply into a mighty people." 16 We are informed that
among the animals that these colonists carried with them were:

All manner of cattle, of oxen, and cows, and of sheep, and of swine, and of goats, and also
many other kinds of animals which were useful for the food of man.

And they also had horses, and asses, and there were elephants and cureloms and cumoms; all
of which were useful unto man, and more especially the elephants and cureloms and
cumoms. 17

The transporting of flocks and herds and all useful animals from the Eastern to the Western
Hemisphere was not all done by the Jaredites in their barges. Later we will discuss other
means by which the fauna of the Eastern Hemisphere reached the Western Hemisphere. At
this point it is sufficient to give the Book of Mormon history of the coming of these useful
animals.

One of the strange things about the arguments and deductions for descent of man from the
lowest forms of life is the scatter-brain way in which the "evidence" is obtained. To illustrate
the point: The Piltdown skull and tooth were found in or near, Piltdown, Sussex, England,
and out of these the vivid imaginations of certain so-called scientists create a whole race of
men; the Trinil Ape-Man, was manufactured from fragments of skull and found at Trinil,
Java, clear around on the other side of the world; then one stage higher, the Neanderthal Man
was manufactured from a portion of skull and a few other fragments of bones, found at the
Neanderthal gorge near Westphalia, Germany; and as we approach further towards the Homo
perfection the discovery is made from bones found at Cro Magnon, Dordogne, France. These
poor fellows must have wandered about a good bit, from England to Java, to Germany to
France, and if we carry it further to deserts of China and even to parts of the Western
Hemisphere. It is too bad that these poor fellows did not keep all their bones in one place so
that the evolutionists would not have to be put to the inconveniences of manufacturing themissing parts. However, any man who can manufacture a Man from an Amoeba through
countless stages covering millions of years, is capable of doing most any thing. The
numerous imaginary pictures of this process published by the gentlemen in the \textit{American
Museum of Natural History} and by writers of numerous textbooks circulated and used in our
public schools and colleges reveal the startling story that we have reached an age when good
clear reasoning and logical deductions are entirely out of place.

It has been truthfully said that organic evolution is Satan's chief weapon in this dispensation
in his attempt to destroy the divine mission of Jesus Christ. It is a contemptible plot against
faith in God and to destroy the effective belief in the divine atonement of our Redeemer
through which men may be saved from their sins and find place in the Kingdom of God.
There is not and cannot be, any compromise between the Gospel of Jesus Christ and the
theories of evolution. Were evolution true, there could be no remission of sin. In fact there
could be no sin. Alma painted the picture many years before the birth of Jesus Christ when he
said:

Therefore according to justice, the plan of redemption could not be brought about, only on
conditions of repentance of men in this probationary state, yea, this preparatory state; for
except it were for these conditions, mercy could not take effect except it should destroy the
work of justice. Now the work of justice could not be destroyed; if so, God would cease to be
God.

And thus we see that all mankind were fallen, and they were in the grasp of justice; yea, the
justice of God, which consigned them forever to be cut off from his presence.

And now, the plan of mercy could not be brought about except an atonement should be
made; therefore God himself atoneth for the sins of the world, to bring about the plan of
mercy, to appease the demands of justice, that God might be a perfect, just God, and a
merciful God also.

Now, repentance could not come unto men except there were a punishment, which also was
eternal as the life of the soul should be, affixed opposite to the plan of happiness, which was
as eternal also as the life of the soul.

Now, how could a man repent except he should sin? How could he sin if there was no law?
How could there be a law save there was a punishment?

Now, there was a punishment affixed, and a just law given, which brought remorse of
conscience unto man.

Now, if there was no law given—if a man murdered he should die—would he be afraid he
would die if he should murder?

And also, if there was no law given against sin men would not be afraid to sin.

And if there was no law given, if men sinned what could justice do, or mercy either, for they
would have no claim upon the creature?

But there is a law given, and a punishment affixed, and a repentance granted; which
repentance, mercy claimeth; otherwise, justice claimeth the creature and executeth the law,
and the law inflicteth the punishment; if not so, the works of justice would be destroyed, and
God would cease to be God.

But God ceaseth not to be God, and mercy claimeth the penitent, and mercy cometh because
of the atonement; and the atonement bringeth to pass the resurrection of the dead; and the
resurrection of the dead bringeth back men into the presence of God; and thus they are
restored into his presence, to be judged according to their works, according to the law and
justice.

For behold, justice exerciseth all his demands, and also mercy claimeth all which is her own;
and thus, none but the truly penitent are saved.

What, do ye suppose that mercy can rob justice? I say unto you, Nay; not one whit, if so, God
would cease to be God. 18

Organic evolution tends to rob God of his mercy, his justice and his saving grace. It denies
the resurrection of the dead and the gift of Jesus Christ to all men that they will live again. It
denies the spiritual creation and places the earth and all of its inhabitants beyond the power
of redemption. It teaches that in some unknown way and at some unknown time, life
commenced in some spontaneous way in a speck of protoplasm. It cannot explain how this
speck of protoplasm, or cell, happened to be. It is merely a postulate, a guess that such a
thing really happened. Therefore man is beholden to no one for his existence. He is not,
according to this theory, the offspring of God. He had no divine origin, no spirit in his body
that is eternal. When he dies he shall return to the dust and death is the end of all. There is no
other conclusion; no doctrine more hopeful than total extinction of the individual. These are
the rewards offered to you and to me and to every creature through this wicked doctrine
which today prevails so nearly universal, making atheists of mankind. Is there any wonder
that men of renown like John Fisk can say: "Theology has much to say about original sin.
This original sin is neither more nor less than the brute-inheritance which every man carried
with him." 19 Or, Sir Oliver Lodge: "As a matter of fact, the higher man of today is not
worrying about his sins at all, still less about their punishment. His mission, if he is good for
anything, is to be up and doing; and insofar as he acts wrongly or unwisely he expects to
suffer. He may consciously plead for mitigation on the ground of good intentions, but never
either consciously or unconsciously will any one but a cur ask for the punishment to fall on
someone else, nor rejoice if told that it already has so fallen." 20

This is what comes naturally out of the doctrine of organic evolution. It ridicules religion. It
denies the Fatherhood of God and the Sonship of Jesus Christ. It places man as the natural
kin of the animal, a descendant of a rat, a worm and an amoeba. Those who like it may have
this doctrine, but they have no right to attempt to drag their fellow men, who are "begotten
sons and daughters unto God," down to their level.

And now, after the many testimonies which have been given of him, this is the testimony,
last of all, which we give of him: That he lives!

For we saw him, even on the right hand of God; and we heard the voice bearing record that
he is the Only Begotten of the Father—

That by him, and through him, and of him, the worlds are and were created, and the
inhabitants thereof are begotten sons and daughters unto God. 21

This is the testimony of two men who were privileged to see the glory of God and his Son
Jesus Christ, not two thousand years ago, but in the dispensation of The Fulness of Times and
their witness confirms the witness of the many prophets from the beginning. In the hardness
of their hearts, men reject all of this evidence and blindly pursue their way to eternal
damnation dragging as many as they can, through their pernicious teachings, with them. The
author of it all is Satan, the deceiver and enemy of all righteousness, who swore to destroy
the world and has been true to his determination through all the ages. Today, as prophetically
predicted by Nephi over five hundred years before the birth of Jesus Christ, Satan rages in
the hearts of the people and "cheateth their souls, and leadeth them away carefully down to
hell." 22

In the spirit of fairplay we may ask WHY are these advocates privileged to teach their soul-
destroying doctrines in our public schools and colleges, when the doctrines of Jesus Christ
are barred? Why are so many of the textbooks adopted in these schools steeped in these
theories and by these means the minds of our children poisoned and their faith destroyed? In
this land of America our fathers fought for religious and political freedom and our nation is
now confronted with a far more deadly sin—the destruction of their faith in the Living God!
These perpetrators of this doctrine either know that they are deceivers, or else they have been
completely blinded by the arch-enemy of divine truth. Why should those who believe in the
creation of man, the fall and the atonement of Jesus Christ, have to submit to these dogmatic
theories being promulgated and taught in the textbooks of our schools? We should take a
stand for our religious freedom from this contaminating influence which dominates so much
of our education.

This same practice prevails in the schools of Great Britain, but there THE VICTORIA
INSTITUTE, or PHILOSOPHICAL SOCIETY OF GREAT BRITAIN, which has in its
membership many noted scientists, educators, lawyers and statesmen, entered a protest. By
their Secretary, they appealed to the President of the Board of Education in the following
communication:

To the President of the Board of Education,

Whitehall, London, S. W. 1.

The Victoria Institute, or Philosophical Society of Great Britain, 1, Central Building,
London, S. W., 1.

October 6th, 1941

Sir,

I have the honour, by request of the Council of the Victoria Institute, or Philosophical
Society of Great Britain, to bring to your notice a complaint that is being made in respect of
the manner of teaching biology in schools.

Representations have been made to the Council that correct teaching of biology is prejudiced
by the use, \textit{with official sanction,} of textbooks in which the theory of organic evolution is
assumed to be a fact of science and, while facts which appear to favour the theory are
stressed, others which tell heavily against it are left unmentioned; are, in effect, suppressed; a
method of teaching which, if employed, is unscientific and unphilosophical.

In consequence of these representations the Council have felt it desirable to appoint, from
among the Fellows of the Society, a small committee of gentlemen possessing necessary
scientific qualification, to examine into the matter and report to them on the allegation of
unfair teaching.

A copy of this Committee's report, substantiating the original representations, is submitted
herewith, for favour of perusal by the Board of Education. It embodies a few quotations only,
by way of examples, but many more can be furnished, if desired.

The Council venture to hope that the Report may be given the serious consideration it merits,
since, in their opinion, ill consequences are bound to result; not only from suppression of
truth, but worse, from the harnessing of immature minds in advance to a theory, adoption of
which in certain quarters as the basis of philosophy of life has already, they have good reason
to know, served to undermine belief in God and in man's accountability to Him, with
disastrous results in the moral realm.

They desire me to lay special emphasis on the point at issue in this submission, which is
solely that of \textit{unfair presentation.} They raise no objection whatever to employment of the
theory, or any modification thereof, as a \textit{mode} of teaching, a working hypothesis, \textit{provided
the facts standing against it are given equal prominence with those in its favour,} and that
discussion of the subject be reserved entirely for older scholars; who should also be warned
that, though at present popular with many biologists, the theory is not accredited by all, has
certainly not been proved true, and to which, in point of fact, the fossils at present known are
very unfavourable.

I have the honour to be, Sir

Your obedient Servant,

T. C. Skinner, \textit{Lt.-Col.}

\textit{Honorary Secretary.}

Accompanying this communication was the report of the Committee of Fellows showing
their findings and calling attention to, careless and inaccurate phraseology, dogmatic
statements of theories which have not been proved true, the one-sided presentation of
theories and "assertions demonstrably false." It would be a timely thing if such action could
be taken in the United States in fairness to all those who believe in the mission of Jesus
Christ and the divine origin of man.

\newpage
REFERENCES—CHAPTER EIGHT

Footnotes

1. Fleming, Sir Ambrose, \textit{Introduction to Evolution or Creation,} VIII, IX.

2. James 1:17.

3. Genesis 1:24-27.

4. \textit{Documentary History of the Church,} Vol. 4, pp. 554-555.

5. Abraham 4:11-12.

6. D. \& C. 88:36-39.

7. Nelson, Byron C., \textit{After Its Kind,} pp. 19-20.

8. Acts 17:24, 26.

9. McCann, Alfred W., \textit{God or Gorilla,} pp. 167-168.

10. Howorth, Sir Henry, \textit{The Mammoth and the Flood,} Preface XI-XII.

11. \textit{Journal of Transactions,} Victoria Institute, Vol. 82, pp. 152-173 (1950).

12. \textit{Ibid}., p. 165.

13. \textit{Ibid}., p. 155.

14. 1 Nephi 18:25.

15. Enos 21.

16. Ether 2:1-3.

17. \textit{Ibid}., 9:18-19.

18. Alma 42:13-25.

19. Fisk, John, \textit{The Destiny of Man}, p. 103.

20. \textit{Man and the Universe}, p. 204.

21. D. \& C. 76:22-24.

22. 2 Nephi 28:20-21.

