\chapter{THE FLOOD}

THE story of the flood is one of the events recorded in the Bible that has met with contempt
and ridicule by the ``wise'' and ``prudent'' of Isaiah's prediction, in these modern days of
worldly learning. So vigorous and positive have been the attacks by these learned gentlemen,
that many professed believers in the Bible have either accepted the modernistic view, or have
compromised, and reached the conclusion that there was a flood, but it was only a local flood
somewhere in or near the Mesopotamia valley. One writer accepting this compromise while
admitting the dimensions of the ark were as great as those of a modern ocean liner, endeavors
to show that Noah in the ark was carried on the crest of an exceptional flood that swept over
the boundaries of Lake Van, while the people in the surrounding territory perished. I regret to
say that there are some members of the Church, lacking in faith, who have taken a similar
view. One local professor of geology states that torrential rain for forty days and forty nights
would be required to equal twenty-five feet of rain every hour, or about six hundred fifty feet
every day.'' 1 This writer states that at one time—according to geological time however,
antedating the time of Noah—the northern country was covered with glaciers, and while he
has no ``fixed'' opinion in the matter, ``it would be interesting to know whether the period of
flooding'' from the melting ice, and the Biblical flood were ``identical.''

He further says:

The writer can also picture the manner in which God may have brought about his purposes.
In order to supply an illustration, one needs to go only to an ordinary valley and witness the
conditions due to stream flooding. Almost monthly our newspapers print accounts of towns
and villages having been invaded by high waters incident to so-called cloudbursts or to
rapidly melting snows.

Following the flood, the writer can think of the ark as drifting down the swollen stream and
perhaps out into the ocean. He can also think of the long days of waiting and the final
landing. (Bible students will recall that the narrative is indefinite as to the precise country in
which Noah and his associates lived in.)

The writer can readily understand how that under the simple conditions the Bible account of
the flood could have been written. The interpretation suggested above is, of course
hypothetical, but it does not detract from the dignity of the account; rather it heightens it,
since it shows how simply and how literally God could bring about his purposes. 2

Of course, this learned gentleman had a very vague idea as to the extent and purpose of the
flood. There are others in the Church, limited, we hope, in number, who reject the story of
the flood entirely on the ground that it could not be an act of God; for he is a ``kind Father''
and ``a father, even a normal, earthly father, could not command such wholesale slaughter,'' 3
and such doctrines as this and others regarding great destructions portrayed in the Bible, are
``the doctrine of devils,'' which a merciful God could not do, and therefore these
humanitarians ``refuse to bring the God they worship into any of these bloody conflicts.'' 4
Therefore these stories cannot be sanctioned by them.

It seems to be extremely absurd to think of Noah building an ark as big as the \textit{Queen Mary},
to sail down an unusual flood in a valley to the ocean. Moreover, why would the Lord go to
the trouble of having all the animals and birds come to the ark for protection rather than to
have them driven out of the afflicted valley? And why make Noah and his sons go to the
trouble of building such an ark when he and they too, might move to another more peaceful
valley? Evolutionary writers invariably attack the story of the flood as they do the account of
the divine purpose of the creation. It interferes with their theories and therefore it must not
stand! Dr. Andrew D. White has been most bitter in his denunciations of the divine creation
and the fall. They have received his sharpest criticism. Likewise his writings are without
mercy for believers in the story of the flood; but he is only one of the multitude of such
writers. The arguments used by Dr. White and others in their attempt to refute the story of the
flood are varied and unique. Among the most outstanding of these criticisms are the
following: Out of the legends of the Chaldeans these stories come and have been
incorporated in the Hebrew scriptures; there are too many animals on the earth to get in the
ark; there are too many animals and races of men in America, Australia, and other parts of
the earth that never could have reached the ark and if they could there was no hope of their
ever getting back to those distant lands; the mountains are too high and no flood could cover
them.

Dr. White, on these points, has this to say:

Even more and more difficult, too, became the question of the geographical distribution of
animals. As new explorations were made in various parts of the world, this danger to the
theological view went on increasing. The sloths of South America suggested painful
questions: How could animals so sluggish have got away from the neighborhood of Mount
Ararat so completely and have travelled so far?

The explorations in Australia and neighboring islands made matters still worse, for there was
found in those regions a whole realm of animals differing widely from those of other parts of
the earth. 5

``By the middle of the nineteenth century,'' says this noted gentleman, ``the whole theological
theory of creation—though still preached everywhere as a matter of form—was clearly seen
by all thinking men to be hopelessly lost.'' 6

The honored Doctor is sadly mistaken; but to the many who are ignorant of the restoration of
the Gospel these criticisms carry considerable weight, and may appear conclusive arguments
against the belief in the universal flood, causing many defenders of the flood to admit their
inability to cope with the situation. I shall endeavor, therefore, to answer these criticisms,
most of which have been successfully met by others. The most difficult one to the defenders
of the universal flood is this one presented by Dr. White in relation to the distribution of
animals and races to all parts of the earth. There is in the Book of Mormon, Jacob, chapter
five, a parable given by the revelation of the Lord that every member of the Church should
read carefully. It is the story of the scattering of Israel and other peoples, to all parts of the
earth. Not only did the Lord send colonies to America after the flood, but to many other
lands. The Jaredites who left at the confounding of tongues, brought to this western
hemisphere the elephant, ox, goat, sheep and all kinds of animals. They could have brought
even the sloth. Moreover we have learned from the Book of Mormon that the Lord
distributed the inhabitants of the earth and planted colonies in the islands of the sea, more
than likely in Australia, New Zealand and in all parts. To Australia colonies could have taken
the kangaroo. The evidence is furnished us that colonies went forth to the islands of the
Pacific from America after America had been colonized. Reason tells us that they may have
taken animals with them. So the peopling of the earth, and the distributing of animals as the
Lord thought good, was carried on under his direction. The Jaredites, Nephites, Mulekites, all
were directed to the western hemisphere. It should also be remembered that the Lord said at
the time of Babel, ``So the Lord scattered them abroad from thence upon the face of all the
earth.'' (Gen. 11:8.) And in verse 9: ``and from thence did the Lord scatter them abroad upon
the face of all the earth.'' According to his promise and decree, he had to scatter them to
every land.

Again, in the days before the flood, the land surface of the earth was \textit{all in one place}. It was
not until shortly before the confusion of tongues that the land surface of the earth was
divided, and following that time many colonies could have been sent just as the Jaredites
were, to inhabit various parts of the earth. The time must come in the restoration of all things,
for all the land to come back again to its original place, and the waters to be driven back from
whence they came. (D. \& C. 133:19-35.) So the \textit{difficult} problem is easily solved when we
have the truth with which to solve it. And, it is not out of order at this point, to call attention
to the fact that these highly cultured gentlemen are blaming the Bible and the Church of Jesus
Christ as it existed in the beginning with all the errors in theology, doctrine and practices,
that crept into the apostate church and which are not believed in, and never were accepted, by
the people of God in any age of the world's history. The misinterpretations of the scriptures,
the mistakes that crept into them, or were deliberately placed there by scribes and priests who
did not understand the truth, cannot be blamed upon the original writers—the prophets who
wrote and spoke who had the inspiration of the Holy Ghost.

It has been computed by reliable persons that the ark according to the sacred cubit giving its
dimensions, would be as large as one of the largest steamships ever built. Now no one, I
suppose, has ever tried to put on one of these great \textit{Leviathans} or \textit{Queen Elizabeths} two of
every species of animal and seven of the clean kind.

Paul said in truth on Mars Hill:

God that made the world and all things therein, seeing that he is Lord of heaven and earth,
dwelleth not in temples made with hands;

Neither is worshiped with men's hands, as though he needed any thing, seeing he giveth to all
life, and breath, and all things;

And hath made of one blood all nations of men for to dwell on all the face of the earth, and
hath determined the time before appointed, and the bounds of their habitation;

That they should seek the Lord, if haply they might feel after him, and find him, though he be
not far from every one of us:

For in him we live, and move, and have our being; as certain also of your own poets have
said, For we are also his offspring. 7

All of the human inhabitants of the earth are descendants of Noah. It is written in the word of
the Lord:

And now, after the many testimonies which have been given of him, this is the testimony,
last of all, which we give of him: That he lives!

For we saw him, even on the right hand of God; and we heard the voice bearing record that
he is the Only Begotten of the Father—

That by him, and through him, and of him, the worlds are and were created, and the
inhabitants thereof are begotten sons and daughters unto God. 8

The world does not know why there are white, black, brown and reddish colored people on
the earth—all descendants of Noah, but the Latter-day Saints do. The Lord placed the black
skin on the descendants of Cain, and they came through the flood in the family of Ham. We
know why the Lamanite is reddish in appearance, and for like reasons other races are so
marked, but they had one ancestry. So in the ark was a pair of dogs, a pair of lions, a pair of
elephants, and so on through all the animal and bird kingdoms. From these original pairs
have come the varieties of dogs, lions, bears, horses, elephants, etc. But they are all brought
into the world \textit{after their kind}. A dog never begets a sheep, nor a horse a cow, an elephant or
anything else but its own kind, and so on throughout the whole creation. Now no man has the
knowledge, much less the authority, to say that all the original pairs of animals and birds
could not be placed on the ark! The word of the Lord says they \textit{were} placed there, and
through the providence of God they were distributed over all the face of the earth after the
flood.

We read in the word of the Lord to Moses:

Remember the days of old, consider the years of many generations: ask thy father, and he
will shew thee; thy elders, and they will tell thee.

When the Most High divided to the nations their inheritance, when he separated the sons of
Adam, \textit{he set the bounds of the people according to the number of the children of Israel.}

For the Lord's portion is his people: Jacob is the lot of his inheritance. 9

Let it be remembered that the Lord set these bounds for the nations according to the number
of the children of Israel, and the vast majority of the children of Israel were not yet born!
From this we learn that the Lord had a controlling hand in where the nations were to locate,
and it was not merely a matter of chance. We know the Jaredites were led and appointed, so
were the Nephites and the Mulekites, to the western hemisphere. So likewise the ``Lost Ten
Tribes,'' and many others according to the revealed word of the Lord. So we discover that
there is no great unsolved, and unsolvable mystery about the kangaroo being in Australia, the
sloth in South America with the anteater and the buffalo in the United States. The Lord who
gathered them to preserve them also distributed them and this by natural means. The poor
sloth was not forced to make the tedious journey across continents and seas to his home in
America, neither the kangaroo in Australia.

Another criticism of the universal flood is supposed to be indisputable. This is that there is
not enough water in the oceans, rivers and lakes, plus what may be stored in the atmosphere
and in the earth to rise and cover the mountains. The account of the flood and from whence
the water came is but briefly stated. We have no extensive details. The story is as follows:

In the sixth hundredth year of Noah's life, in the second month, the seventeenth day of the
month, the same day were all the fountains of the great deep broken up, and the windows of
heaven were opened. 10

The rain was upon the earth forty days and forty nights. No man is able to explain just what
the Lord meant by the opening of the ``windows of heaven.'' Was water brought for the
occasion from some other place above the earth? We do not know; but the true believer in the
scriptures having faith in the word of the Lord, has no doubt that sufficient water was
supplied. It is an interesting study nevertheless, considering just the condition that exists with
the amount of water in the oceans, lakes and in the atmosphere. We receive the information
from government sources, gathered by scientific investigation of the following interesting
facts:

The land area of the earth is 57,510,000 square miles.

The water area is 139,440,000 square miles.

Circumference at the equator is 24,902 miles.

The figure given for the height of Mt. Everest is 29,002 feet. To make the problem simple we
will say Mt. Everest is six miles high and that the circumference of the earth is 25,000 miles
at the equator. This would make Mt. Everest 6/25,000 of the earth's surface rising into the air.
Charles Babage, the English scientist, said, ``The highest ranges of mountains we have are
relative to the circumference of the earth's crust infinitely smaller than the puckers on an
orange-skin.'' 11 With nearly two and one half times more water than land, I am sure the
Lord with his infinite power could manipulate this immense amount of water to cover the
earth. The ocean has an average depth of 13,000 feet and the average height of the land is
only about 2,300 feet. The ocean varies in depth from 300 to 400 to about 31,000 feet, so it is
about 5.6 times as deep as the average land is high and the ocean area is far more than two
times that of the land. Then again, we have learned that the mountains were \textit{not} as high in the
days of Noah as they now are. There were great changes that came to the surface of the earth
during the flood. We know from the dynamic force of water a flood of such proportions
could not occur without making great changes in the surface of the earth. This the
evolutionist does not take into account. Moreover, other great changes came in the day of
Peleg when the earth was divided. (Genesis 10:25.) This was the surface of the earth, not the
division in relation to the habitations of the people. One fact easily overlooked in the Lord's
instruction to Noah, (see Genesis 6:13,) is as follows:

And God said unto Noah, The end of all flesh is come before me; for the earth is filled with
violence through them; and, behold, I will destroy them \textit{with the earth}.

A deluge such as that described in the Bible, it must be admitted, would make great changes
in the earth. A small flood can create enormous damage, such as we occasionally have in
parts of Utah. Then to think of the terrific force of the flood that would cover the earth we
must conclude that the entire face of the earth was changed, so that many of the great
geological conditions ascribed by geologists to a time millions of years ago could have taken
place suddenly, and evidently did. We know from the record in the Book of Mormon, that at
the crucifixion of our Lord, the whole face of the land on the western hemisphere was
altered. Mountains arose, others sank, and they were broken into ``faults'' and crags by the
great earthquakes, wherein they were smooth before, and this was not millions of years ago.

Another things should be considered. There is definite evidence that at one time, and I am
convinced it was in antediluvian days since the time of Adam, the climate of the earth was
just as Elder Parley P. Pratt and President John Taylor have described it. Evidence is found in
the Arctic that tropical, or semi-tropical plants, grew there in abundance. The frozen animals
spoken of by Sir Henry H. Howorth were frozen at the time of the flood when the climatic
conditions of the earth were changed. The time is drawing near when this primitive condition
will be restored again. In the restitution of all things it has to be.

In this dispensation we are promised by revelation coming from the prophets ``since the
world began,'' that there is to be a restoration of the earth to its primitive condition, when the
promised millennium shall come. We are led to believe, then, that in the beginning the
mountains were not as high as they are now, for the prophecies declare that in that day
``Every valley shall be exalted, and every mountain and hill shall be made low: and the
crooked shall be made straight, and the rough places plain: And the glory of the Lord shall be
revealed, and all flesh shall see it together: for the mouth of the Lord hath spoken it.'' 12
When the valleys are exalted it will be the natural result of the mountains being debased or
made low. In this way the earth will be restored to its primitive condition.

We are taught in the Doctrine and Covenants (Section 133:24), that the ``great deep'' in this
day of restoration will be driven \textit{back} to the north, ``and the land of Jerusalem and the land of
Zion shall be \textit{turned back into their own place}, and the earth shall be like it was in the days
before it was divided.'' There are several passages of prophecy in the Bible where the Lord
speaking through his prophets, and having reference to the restoration of all things, declared
that the mountains will be lowered and the valleys raised when the Lord comes. Here are
some of the references:

In a little wrath I hid my face from thee for a moment; but with everlasting kindness will I
have mercy on thee, saith the Lord thy Redeemer.

For this is as the waters of Noah unto me: for as I have sworn that the waters of Noah should
no more go over the earth; so have I sworn that I would not be wroth with thee, nor rebuke
thee.

For the mountains shall depart, and the hills be removed; but my kindness shall not depart
from thee, neither shall the covenant of my peace be removed, saith the Lord that hath mercy
on thee. 13

For in my jealousy and in the fire of my wrath have I spoken, Surely in that day there shall be
a great shaking in the land of Israel;

So that the fishes of the sea, and the fowls of the heaven, and the beasts of the field, and all
creeping things that creep upon the earth, and all the men that are upon the face of the earth,
shall shake at my presence, and the mountains shall be thrown down, and the steep places
shall fall, and every wall shall fall to the ground. 14

And every island fled away, and the mountains were not found. 15

All of this will come to pass as part of the restoration. This will happen when our Savior,
Jesus Christ, comes in his power, and ``he shall utter his voice out of Zion, and he shall speak
from Jerusalem, and his voice shall be heard among all people; And it shall be a voice as the
voice of many waters, and as the voice of a great thunder, which shall break down the
mountains, and the valleys shall not be found.'' 16

Sir Henry H. Howorth has written a number of wonderful books in which he endeavors to
correct the mistakes of present-day scientists in relation to the geology of the earth. These
books are, \textit{The Mammoth and the Flood}, 464 closely written pages, in which he definitely
proves that there is evidence in all parts of the earth of a universal flood. He disagrees
emphatically with the prevailing notion that there was an ``Ice Age,'' and on this question has
published two large volumes which have the title, \textit{The Glacial Nightmare and the Flood}; and
two other large volumes, \textit{Ice or Water}. These books are masterful presentations of the
subjects under discussion; but they are too voluminous for any more than a reference to a few
salient items, and I recommend their study to all who are troubled with these inconsistent
theories of evolution, and if they will do so sincerely I am sure they will be greatly benefited.

One important study of these volumes having a bearing upon the flood, is the evidence that,
not only the mammoth of Siberia were buried suddenly having met with violent death, but in
all parts of the earth similar evidences are found in relation to other animals. This could not
have happened from the freezing of ice, but it came through a sudden and violent flood of
water. I quote a few paragraphs:

Not only does the frozen ground preserve the flesh deposited in it, but it is quite clear that no
flesh could remain intact in this way unless it were permanently frozen, and it follows
inevitably that the bodies of the mammoths, etc., which are now found intact in the Siberian
tundras must have been frozen immediately after death, and have remained frozen since they
were first entombed. If they had been subject to alternating congelation and melting with the
intermittent seasons, they would assuredly have long since decayed. An exposure to one
summer's sun, to one season's melting, would have induced putrefaction and dissipation. We
are not dealing here with animal substances deposited in bogs, and changed into such organic
compounds as adipocere, but of flesh so unchanged that it has all the character of that of
animals which have recently died, when examined under the microscope, while it is readily
eaten by the wild animals that live on the tundra. The flesh is as fresh as if recently taken out
of an Esquimaux cache, or a Yakut subterranean meatsafe. There cannot be a moment's doubt
that this condition was secured by one cause only, namely, that ever since the bodies were
entombed they have been in a state of continuous congelation without a break.

This is assuredly the only possible conclusion. It is one which I have urged at different times
before the Geological Society and the British Association, in the pages of \textit{Nature} and the
\textit{Geological Magazine}, and personally to several of my most distinguished scientific friends
with a much wider experience of such problems than my own, and on all occasions there has
been a consensus of opinion that what is here urged is inevitable, nor is the opinion, so far as
I know, contested. . . .:

This view seems incontrovertible, but it involves as a simple necessary corollary a conclusion
from which Lyall and his followers have continually shrunk, namely, that this change of
climate must have been sudden and must also have been continual. To avoid this conclusion
some desperate efforts have been made. 17

Again, as I have said, the instance of the soft parts of the great pachyderms being preserved
are not mere local and sporadic ones, but they form a chain of examples along the whole of
Siberia, from the Urals to the land of the Chukchis, so that we have to do here with a
condition of things which prevails, and with meteorological conditions that extend over a
continent.

When we find such a series ranging so widely preserved in the same perfect way, and all
evidencing a sudden change of climate from a comparatively temperate one to one of great
rigor, we cannot help concluding that they all bear witness to a common event. We cannot
postulate a separate climatic cataclysm for each individual case and each individual locality,
but we are forced to the conclusion that the now permanently frozen zone in Asia became
frozen at the same time from the same causes. 18

At the risk of being burdensome I feel it necessary to present other quotations in regard to the
sudden and violent deaths seen everywhere on the earth, where myriads of animals and fish
perished together. All the evidence points to the fact that death came upon them suddenly.

Again, in the pleistocene beds we are arrested by the large number of young animals which
occur. When Nature puts a term to an animal's life in her normal way, it is exceedingly
seldom she does so when the animal is young. Animals do not die naturally in crowds when
young, and yet we find remains of quite young animals abounding in all classes from
mammoths to mice. How are we to account for this fact, save by summoning an abnormal
cause? How, again, can we account for the fact that the mummied animals found in Siberia
seem to have been in robust health, stout and strong? Is this, again, consistent with a natural
death? Again, if the death was natural, and in an area where we know hyaenas and other
carniverous animals abound, would the corpses be left to the useless duties of decay, as they
must have been, since the bones are ungnawed, and (where the flesh is preserved) the flesh is
uneaten? One cause, no doubt, of the scarcity of remains of animals which are dying at
present where animal life abounds, is the diligence of the scavengers. What were they doing
in pleistocene times to pass by those myriads of corpses, and in so many cases not to leave a
tooth-mark anywhere, and in fact to leave their own bones with the rest? Surely the point is
clearly and unmistakably to the fact that the animals, or the greater part of them died
together. If the remains were the silent chronicles of centuries of time and generations of life,
we would assuredly have found that some or a large portion of the bones would have been
broken and gnawed, but this is not the case, and it points strongly to their death having been
more or less simultaneous.

The most obvious cause we can appeal to as occasionally producing mortality in a wide scale
among animals is a murrain or pestilence, but what murrain or pestilence is so completely
unbiased in its action as to sweep away all forms of terrestrial life, including man as we shall
see presently, the fowls of the air and the beasts of the field, elephants and mice, rhinoceroses
and frogs, bisons and snakes, tigers and land-snails, and this is not one corner only, but, so
far as we know, over the whole length of two continents, irrespective of latitude or
longitude? The problem has only to be stated thus to make it obvious that a murrain or a
pestilence is quite incompetent to meet our difficulties. Such a pestilence, again, would not
collect herds of incongruous animals in the same places, and kill them all together, and then
bury them; and if it did so, we should assuredly have some evidence of its work in the
remains themselves, where we find none, but rather that the animals died in full health, with
their bodies strong and hearty.

Another cause of wide-spread death sometimes, is the occurrence of unusual drought. A
fierce summer may dry up the streams and destroy the means of life, but how can we invoke
such a cause operating not only in Siberia, but in the Mediterranean borders, and, as we shall
presently see, from one end of America to the other? The same reason precludes our
appealing to a wide-spread burning of forests or a sudden succession of cold, etc. In Siberia,
as we have seen, the death of the mammoth and its companions was immediately followed by
a sudden declension in temperature, which if itself and in the absence of other causes made
that area incompatible with the life conditions of the mammoth, but this was a correlative
occurence, and not the immediate cause of the great destruction, since it only operates over a
very small portion of the area from which the mammoth and its companions disappeared.
Nor, again, is it likely that the cold would have killed the bear, and glutton, the musk-sheep
and reindeer, the snowy owl and ptarmigan, which we know were in many instances
overwhelmed by the same cause, whatever it was that destroyed the mammoth.

The fact that great bones occurring in great caches or deposits, in which various species are
mixed pell-mell, is very important. If animals die occasionally from natural causes when they
become toothless and old, different species do not come together to do so, nor does the lion
come to take his last sleep with the lamb. The fact of finding masses of animal remains of
mixed species, all showing the same state of preservation, not only points to a more or less
contemporary death, but is quite fatal to the theory that they ended their days peaceably by
purely normal causes.

Bones, again wither and decay very fast, if exposed to the air. And consequently when we
find bones, with their fine and delicate angles and muscular attachments preserved intact, in
many cases lying together as when articulated, over the wide area of a whole continent, and
for the most part, so far as we can judge, in the same mineral condition and state of decay, we
must conclude that these facts are only consistent with the animals having died together, and
been together protected from decay. . . .

We must next inquire what the nature of this catastrophe was. Let us, then, focus the
necessary conditions. We want a cause that should kill the animals, and yet not break to
pieces their bodies, or even mutilate them, a cause which would in some cases disintegrate
the skeletons without weathering the bones. We want a cause that would not merely do this
as a wide-spread murrain or plague might, but one which would bury the bodies as well as
kill the animals, which could take up gravel and clay and lay them down again, and which
could sweep together animals of different sizes and species, and mix them with trees and
other debris of vegetation. What cause competent to do this is known to us, except rushing
water on a great scale? Water would drown the animals, and yet would not mutilate the
bodies. It would kill them all with complete impartiality, irrespective of their length, age, or
size. It would take up clay and earth, and cover the bodies with it. This is the very work it is
doing daily on a small scale. Not only could it do this, but it is the only cause known to me
capable of doing the work on a scale commensurate with the effects we see in Siberia. What
direct evidence, then, have we that it was in fact a great flood of water?

The first piece of evidence I would quote is of a singularly direct kind, and we owe it to the
experienced skill of Professor Brandt. Speaking of the famous rhinoceros found on the Wilui
by Pallas, he says, ``On a careful examination of the head of the \textit{Rhinoceros Tichorinus} from
the Wilui, it was further remarkable that the blood-vessels and even the fine capillaries were
seen to be filled with brown coagulated blood, which, in many places still preserved its red
color.'' This is exactly the kind of evidence we look for when we want to know whether an
animal has been drowned or suffocated. Asphyxia is always accompanied by the gorging of
the capillaries with blood, and the facts justify at all events a probable inference that this
particular rhinoceros was the victim of drowning.

Brandt goes on to tell us how, in conjunction with Hedenstrom, he made a careful
microscopic examination of the earth which was attached to these rhinoceros remains, and
found it to consist of two kinds, the most important being mould containing vegetable
fragments, and which he took for remains of fresh-water plants, and the soil from a fresh-
water deposit. . . .

To continue: The occurrence of immense caches in which the remains of many species of
wild animals are incongruously mixed together pell-mell, often on high ground, seems
unaccountable, save on the theory that they were driven to take shelter together on some
point of vantage, in view of an advancing flood of water, a position which is paralleled by the
great floods which occur occasionally in the tropics, where we find the tiger and its victims
all collecting together on some dry place, and reduced to a common condition of timidity and
helplessness by a flood which has overwhelmed the flat country. . . . In the present case all
were overtaken by the water, tossed and tumbled together in a common destruction, and then
covered quickly with a mantle of clay or gravel,—a mantle, be it remembered, spread over
immense areas, without a break external or internal, and in which we can find no traces of
local disturbance, such as would be caused by any process of subsequent burying, and
showing the bones and covering were laid down together. 19

Sir Henry H. Howorth continues his story and gives indisputable evidence that these
evidences of the universal flood are in many parts of the world. It was not confined to
Siberia, but evidence shows it prevailed in America and in the caves all over the world. He
also continues his study and shows by positive evidence, ``That man was a contemporary of
the extinct animals.'' 20 Not Neanderthal or any other kind of manufactured man, but real
human beings. Near the close of the book, \textit{The Mammoth and the Flood}, the author gives
evidence that the tradition of the flood is found in countries in all parts of the globe, and of
this he says (pages 412-413), ``But a few words remain to be said about another class of
evidence than that which I have hitherto adduced, and which supports the same conclusion;
namely, the evidence of human tradition. It is in itself not improbable that if such a
catastrophe really occurred it would leave such an impression on the survivors that it would
be recorded in their traditions, and if such tradition occurred in various isolated localities it
would no doubt strengthen its force. In many places such a reminiscence would die out as the
traditions of the direst disasters are apt to die out, but the fact of its existence in several
localities would, as I have said, be an element of singular cogency in the general argument.''

He then records the traditions of people in many lands from Egypt, Israel, Chaldea, Syria,
Phrygia, India, Persia, Greece, China, Mexico, Peru, Brazil, Tahiti, and Polynesian islands,
Fuegian, Philippine and other lands. The fact that such traditions are found in all parts of the
world cannot be lightly ignored.

After presenting all of this by way of argument, the thought is still impressed on my mind,
why should time be spent in argument to prove true what the Lord has so definitely declared
by the words of his own mouth? There are, however, so many possessed with skeptical minds
that it seems that more than the direct word from the Lord is required to convince them. The
real reason for the covering of the earth by a flood thus far has not been mentioned. All the
Lord had to say to Noah about the corruption of all things on the face of the earth is true, but
there was another reason for such a flood. We have called attention to the fact that the earth
itself had become corrupt: ``And God looked upon the earth, and, behold, \textit{it was corrupt}; for
all flesh had corrupted his way upon the earth.'' 21 We learn from the revelations given to the
Prophet Joseph Smith, that the earth is a living being. It is obedient to every command,
therefore it is entitled to be sanctified, just as human beings who keep the commandments of
the Lord are to be sanctified. ``And again,'' said the Lord, ``the earth abideth the law of a
celestial kingdom, for it filleth the measure of its creation, and transgresseth not the law—
Wherefore, it shall be sanctified; yea, notwithstanding it shall die, it shall be quickened again,
and shall abide the power by which it is quickened, and the righteous shall inherit it.'' 22
Therefore, in order to sanctify the earth and cleanse it, it had to pass through the water of
baptism, and by water be cleansed. Then, at a later day, it will be baptized with the fire and
the Holy Ghost, just the same for the earth as for humanity who will possess it forever. Now
we begin to see why there had to be a flood. It had to be a complete immersion, just the same
as any other baptism, so the Lord did two things, first, he cleansed the earth of all of its
corruption brought upon it by the wickedness of ``all flesh,'' and second, the earth having
reached the age for such an ordinance, was cleansed from all the sin upon its face.

Moreover, we, the members of the Church, do not have to rely on the record in the Bible
alone. We have also the word of the Lord and his servants in the Book of Mormon and the
Pearl of Great Price, and in these records we find the confirmation of the story of the flood as
recorded by Moses, and also by the antediluvians in the record which was handed down to
Abraham. We will conclude our evidence by quoting the words of our brethren in relation to
the baptism of the earth.

TESTIMONY OF PRESIDENT BRIGHAM YOUNG

Here are the earth and the inhabitants upon its face, organized for the express purpose of a
glorious resurrection. The terra firma on which we walk, is looking forth for the morning of
the resurrection, and will get a resurrection, and be cleansed from the filthiness that has gone
forth out of her. This is Bible doctrine. What filthiness has gone forth out of her? You and I,
and all the inhabitants of the earth; the human body, and all earthly bodies, both animal and
vegetable; are composed of the native element that we breathe, that we drink, and that we
walk upon, we till the earth for our bread, which is one of the materials of which your body is
composed, it comes forth from the native elements into an organized state; what for? To be
exalted, to get a glorious resurrection. We are of the earth, earthy, and not only will the
portion of mother earth which composes these bodies get a resurrection but the earth itself. \textit{It
has already been baptized}. [My italics.] You who have read the Bible must know that that is
Bible doctrine. What does it matter if it is not stated in the same words that I use, it is none
the less true that it was baptized for the remission of sins. The Lord said, ``I will deluge (or
immerse) the earth in water for the remission of the sins of the people''; or if you will allow
me to express myself in a familiar style, to kill the vermin that were nitting, and breeding,
and polluting its body; it was cleansed of its filthiness; and soaked in the water, as long as
some of our people ought to soak. The Lord baptized the earth for the remission of sins, and
it has been once cleansed from the filthiness that has gone out of it, which was in the
inhabitants who dwelt upon its face. 23

The following is from a sermon by President Brigham Young spoken in Ogden, June 12,
1860:

Brethren and sisters, I wish you to continue in your ways of well doing; I desire that your
minds may be opened more and more to see and understand things as they are. This earth, in
its present condition and situation, is not a fit habitation for the sanctified; but it abides the
law of its creation, has been baptized with water, will be baptized by fire and the Holy Ghost,
and by-and-by will be prepared for the faithful to dwell upon. 24

This is from a discourse by Elder Orson Pratt:

The first ordinance instituted for the cleansing of the earth, was that of immersion in water; it
was buried in the liquid element, and all things sinful upon the face of it were washed away.
As it came forth from the ocean flood, like the new-born child, it was innocent, it arose to
newness of life; it was its second birth from the womb of mighty waters—a new world
issuing from the ruins of the old, clothed with all the innocency of this first creation. As men
cannot be born again of water, without an administrator, so the earth required an agency
independent of itself, to administer this grand cleansing ordinance, and restore it to its infant
purity. That administrator was the Redeemer himself. 25

Following is an excerpt from an article entitled, ``Leaves From the Tree of Life,'' by President
Charles W. Penrose:

Thus the inhabitants of earth with the few exceptions that are beyond the power of
redemption will eventually be saved. And the globe on which they passed their probation,
having kept the law of its being, will come into remembrance before its Maker. It will die
like its products. But it will be quickened again and resurrected in the celestial glory. It has
been born of the water, it will also be born of the Spirit, purified by fire from the corruption
that once defiled it, developed into its perfections as one of the family of worlds fitted for the
Creator's presence, all its latent light awakened into scintillating action, it will move up into
its place among the orbs governed by celestial time, and shining, ``like a sea of glass mingled
with fire,'' every tint and color of the heavenly bow radiating from its surface, the ransomed
of the Lord will dwell upon it; the highest beings of the ancient orbs will visit it; the garden
of God will again adorn it; the heavenly government will prevail in every part; Jesus will
reign as its King; the river of life will flow out from the regal throne; the tree of life, whose
leaves were for the healing of the nations, will flourish upon the banks of the heavenly
stream, and its golden fruit will be free for the white-robed throng, that they may eat and live
forever. This perfected Earth and its saved inhabitants will then be presented to the Eternal
Father as the finished work of Christ, and all things will be subject unto the Great Patriarch,
Architect, Creator, Ruler, the Almighty, to whom be obedience and reverence and praise in
all the countless worlds that shine as jewels in His universal Crown. 26

This is the testimony of President John Taylor:

The earth, as part of the creation of God, has fulfilled and will fulfil the measure of its
creation. It has been baptized by water, it will be baptized by fire; it will be purified and
become celestial, and be a fit place for celestial bodies to inhabit. It will become the
residence of those who have abode a celestial law, and of none other, after it has become
purified, and made celestial. 27

Testimony of Elder Orson F. Whitney:

Of Noah's day, it is written: ``As the days of Noe were, so shall also the coming of the Son of
Man be.'' And it was the Son of Man—though I prefer to call him the Son of God—who said
it. Baptized with water in the days of Noah, the earth will yet be baptized with fire and with
the Holy Ghost. 28

\newpage
REFERENCES—CHAPTER TWENTY-ONE

Footnotes

1. Pack, Dr. F. J., \textit{Science and Belief in God}, p. 213.

2. \textit{Ibid.}, pp. 123-4.

3. Smith, J. F., \textit{Signs of Times}, pp. 78-79.

4. \textit{Ibid.}, p. 80.

5. White, Dr. A. D., \textit{History of Warfare of Science with Theology}, Vol. 1. p. 48.

6. \textit{Ibid.}, Vol. 1, p. 49.

7. Acts 17:24-28.

8. D. \& C. 76:22-24.

9. Deut. 32:7-9.

10. Gen. 7:11.

11. Howorth, Sir Henry H., \textit{Glacial Nightmare and the Flood}, Preface XVI.

12. Isaiah 40:4.

13. \textit{Ibid.}, 54:8-10.

14. Ezekiel 38:19-20.

15. Rev. 16:20.

16. D. \& C. 133:21.

17. Howorth, Sir Henry H., \textit{The Mammoth and the Flood}, pp. 93-94.

18. \textit{Ibid.}, p. 96

19. \textit{Ibid.}, p. 229.

20. \textit{Ibid.}, p. 229.

21. Genesis 6:12; Moses 8:28.

22. D. \& C. 88:25-26.

23. \textit{Journal of Discourses}, Vol. 1, p. 274.24. Ibid., Vol. 8, p. 83.

25. \textit{Ibid.}, Vol. 12, p. 364.

26. \textit{Contributor}, Vol. 2, p. 364.

27. \textit{Journal of Discourses}, Vol. 5, pp. 408-9.

28. \textit{Conferences Reports}, April, 1927.

