\chapter{THE HYPOTHESIS OF ORGANIC EVOLUTION—5}

ONE of the greatest evidences of creation is the human body. Whether it is the body of a
man or a beast the evidence is overwhelming, but this discussion will be confined to man.
The bones of the body may be compared to the walls and partitions of a house. Without these
walls and partitions there could be no house. Without the bones there could be no human
body. Muscles, tendons, cartilages, could not exist without the bony structure with which
they are tied. Organic evolution says the bones came because of ``expediency.'' They had to
come because they were needed and it took several millions of years to accomplish it. In the
meanwhile animals were gradually growing bones. According to this guess the first creatures
had no bones, but when they became necessary they ``emerged.'' What could be more absurd!
Let us commence the study of parts of the human body by a brief consideration of the nature
and purpose of the bones.

We will begin with the skull and the bones of the head. The skull is composed of a number of
bones knit together which makes the bony covering for the brain. There are orifices in certain
places that evidently were not placed there by any chance. These are openings for the use of
the eyes and nose and mouth in the front and for ears on each side. These holes are placed
there symmetrically for the protection and beneficial use of these parts. There are also other
openings for nerves, each placed in its proper position according to the need. The use of the
great cavity of the skull is largely for the home of the brain and for its protection against
injury. The brain is the power house and central station for the entire bodily structure. The
upper jaw is composed of two bones, the maxilla, situated right and left which meet in the
middle. The lower jaw, or mandible, is a single bone. In these bones are sockets arranged for
the teeth prepared by divine appointment. This is in general the nature of the bones of the
head.

The backbone, or spine, is connected to the head by two bones, significantly named the atlas
and axis enabling the individual to turn his head forward, backward and from side to side.
The spine is not one bone, but a series of bones joined by cartilages. If it were one bone it
would be rigid so that the individual could not bend and would have to stand or lie—if there
could be any way possible for him to lie, for he could not bend. There are seven vertebrae in
the neck called cervical; twelve from which the ribs extend, called dorsal, and five not
attached to the ribs called lumbar, or generally the ``floating ribs.'' The backbone is the
foundation of the bony structure of the body. It extends from the base of the brain to the
sacrum, a triangular-shaped bone which forms a basis of support for the body. The coccyx
concludes the end of the spinal column. The fact that this spine is a series of bones joined by
cartilages, or gristle, binding the bones together each in its place and making it possible for
free movement of the body, definitely is by design, not by chance. Wherever a bone is joined
to another bone, they are covered with a smooth portion of articular, or joined segments of
cartilage. The ribs are in this manner attached to the backbone and likewise the twelve dorsal
ribs to the breast bone. Thus they form a cavity in which the vital organs of the body are
placed, and by this array of ribs and spine is formed a protection to heart, lungs and other
vital organs.

Then the shoulder blades, or scapula bones, are plates of bone somewhat triangular on the
upper part of each side of the back of the chest. This bone has an inferior angle and a superior
angle, and at the site of the third or external angle there is a thickened mass of bone, on
which in a shallow oval cavity, the glenoid cavity, this is for the head or ball of the humerus.
1 In other words, this cavity is the socket in which the humerus, or upper arm bone fits. The
upper end of the humerus is shaped like half a ball which fits into this cavity, thus making it
possible for the arm to swing right or left, up or down, with great freedom. And they have the
audacity to tell us that all of this came by chance! The scapula is a very heavy bone which
greatly strengthens the shoulders and the back. The collar bone, scientifically called clavical,
is the curved bone connected to the breast bone and to the shoulder. This bone also fits in a
sort of shallow cavity at the shoulder and can be dislocated or easily broken. The upper arm,
or humerus, is a single bone, the lower part of the arm is composed of two bones, the ulna
and the radius, they are joined at the elbow. One may wonder why the upper arm, or
humerus, is a single bone and the lower arm is composed of two bones. This lends to
flexibility and adds strength to the wrist and hand. These two bones of the lower arm are
joined to the bones of the wrist. The wrist is composed of a number of small bones which
give flexibility to the wrist, and in turn these bones are connected with the bones of the hand
and those of the hand to those of the fingers. There is no guesswork in the formation of the
bones of the arm, wrist and hand, nor for that matter, in any other part of the skeletal
structure. If the wrist were composed of one bone it would be useless. If the hand and fingers
were composed of one bone each, they likewise would be useless. If the arm were one bone
from the shoulder it would be useless, and so we might go on. The workmanship of the body
shows definitely and beyond logical contradiction, that it was designed by the Creator.

Briefly we will consider the lower part of the bone formation. The skeleton of the lower part
of the body is attached to the pelvis, a basin-shaped structure of very strong bones. These
bones help to balance the body and like the ribs form a protection for certain other vital parts
of the body. To the pelvis the femur, or upper bone of the leg, is attached. This bone like the
bone of the upper arm is formed with a ball-shaped end that fits into a socket in the pelvis
bone. This gives multiple movement of the leg as the shoulder socket does to the arm, where
multiple movement is most needed, and we may be sure that this was no accident and not due
to ``emergent evolution,'' a meaningless term manufactured to bolster up a faulty theory—but
the intelligent work of the Divine Creator. We might imagine—yet that would be a great
stretch of the imagination—that nature might have made one joint connecting two bones but
the relation between bone and bone, each part being built in a definite way to serve the
definite purpose, and this occurring scores of times, we are forced to the conclusion that all
of this came by design and not by ``emergency,'' or chance. This is just too much to believe,
and there is no real evidence sustaining it. We might as well say—to use a familiar
comparison—that the wheels, joints and springs in a watch came by means of chance, or that
the lense in a telescope just happened to grow there. It would be just as consistent! The
femur, or upper leg, is joined to the fibula, or lower leg bone at the knees, and like the hand,
wrist and arm, it is attached at the knee by design thus giving free movement to the leg, and
this was no accident. Moreover, the lower or fibula bone is attached to the bones of the ankle
and the ankle to the foot and the foot to the toes thus composing a flexible system to make
walking easy. That the bones of the arm, wrist, hand and fingers correspond in large measure
in their unity to the leg, ankle, foot and toes, is evidence that should be good in any court of
justice that they did not reach that condition by any law of chance.

To a person who has faith in God, and believes in the mission of Jesus Christ, such child's-
play is worthy of ridicule. It is astonishing how men can get so far away from the truth in
their stubborn, wilful, blindness to believe that these wonderful parts of the skeleton so
skilfully and artistically placed in the body, each bone having a definite part to play peculiar
to itself to give to the body power and efficient service, came to that place by chance, and by
a continuous change from a jelly-like creature millions of years ago. This doctrine of descent
from lower forms is a trick of the devil whose self-appointed mission is to destroy the work
of God and the divine mission of Jesus Christ. It is unfortunate that the devil finds so many
followers.

Let us now briefly consider the functions of some of the organs of the body, taking as a start
the working of the human heart. The heart starts on its mission beating regularly before the
birth of the infant. It continues so to beat about 72 times a minute throughout life even if it is
one hundred years or more. It is not necessary to go into much detail about the construction
of the heart more than to make clear its main functions. It is normally about five inches in
length, three and one-half inches in its greatest width and two and one-half inches in its
greatest thickness. It is not a very large organ, but it performs a tremendous amount of work
in each single day essential to the continuation of life. When it decides to stop the light of life
goes out; hence it is a most vital part of the human body. It is divided by a longitudinal
muscular septum into two lateral halves, those on the right side being named the right auricle
and right ventricle, and those on the left, the left auricle and the left ventricle. The right side
is the venus (vain) side of the heart. Blood returning from all parts of the body through the
veins enters the right auricle and from this auricle it passes to the right ventricle. From here it
passes through the pulmonary artery to the lungs where it is purified by its passage through
the lungs and then returns to the heart on the left side through the pulmonary veins, into the
left auricle then to the left ventricle and out again on its mission through the arteries and
capillaries to all parts of the body and back again through the veins to the heart—one
constant round. This flow of blood goes on day and night, the heart and the circulatory
system being involuntary in their action over which the person has no control. In this manner
the heart controls two streams of blood, one going from the heart and the other returning to
the heart in separate channels. The blood goes out on its journey charged with oxygen and
returns with carbon-dioxide and impurities gathered on its journey to be cleansed by the
lungs. The blood is prevented from flowing in a contrary direction by valves in the heart
between each auricle and ventricle. The blood is propelled from each chamber by the
contracting of its muscle walls and when the muscle relaxes blood flows into the chamber
which opens to receive it and then closes so that it cannot return. The capacity of each
chamber of the heart in adults is about three to three and one-half ounces. If you are a good
mathematician you might figure the weight of the blood pumped through the heart to all parts
of the body each twenty-four hours, at 72 beats per minute. It is stated that each chamber of
the heart has a brief rest period between the time it is empty and the time it fills again, which
is a providential provision that must have been, as the entire action of the heart must also
have been, by divine appointment. This is a very brief and incomplete treatment of the
working of the heart, but it is sufficient for the purpose and reveals the truth that it had to
come by divine decree through the power of the Creator. It is too marvelous and wonderful
for any reasonable mind to seriously believe that the action of the heart and the circulation of
the blood by which the body is built up, and torn down, and kept free from disease, could
possibly have come by chance or any other cause except the will of the Divine Creator. It is
just another marvelous evidence of the creation of man. Why should anyone be fooled or
turned away from faith in God by such nonsensical theories as these which prevail in the
world in this supposedly enlightened age?

Wo unto the deaf that will not hear; for they shall perish.

Wo unto the blind that will not see; for they shall perish also. 2

Now a few words about the marvelous and untiring duties of the blood. The knowledge of
most of us regarding the blood is that when we get cut we bleed, and that it is a red liquid
pumped by the heart to all parts of the body. As to why the blood should thus flow through
arteries and veins in a constant circulation we have given little thought. We have been
satisfied to leave the technical knowledge to the physician, without bothering to know the
cause. However, the circulation of the blood is one of the most vital actions of the body. So
many are the duties performed that when we learn of them we stand overwhelmed with awe.
So it is with the humble man when considering the duties and structure of every other part of
the body. Should the blood decide not to flow and become stagnant we would instantly die. If
for some cause it should cease to pass through the heart to the lungs and be cleansed,
likewise we would die.

The carrying of oxygen from the lungs to the various tissues of the body is not the only
function of the blood; nor is the gathering of the carbon dioxide in the body and delivering it
to the lungs to be thrown off into the air, the only other function. Yet if the oxygen taken in
by the lungs should cease to be delivered, within a few minutes we would die. So also if the
poison gases gathered by the blood were not delivered to the lungs to be exhaled into the air,
in a remarkably short time we would die. Nor is this all. The blood gathers other waste
particles and delivers them to other organs and glands to be eliminated, and if any of these
fail in their purpose, serious results are bound to follow. Now to say that all of this and the
many other actions of the blood come without intelligent direction and appointment, but by
chance, is just too much to ask any man to believe. If it is not by the directing hand of God, it
never could have happened.

We are told that the blood is a ``complex mixture.'' It consists of a fluid called plasma which
contains many important substances in solution. Floating in the plasma are millions of red
cells, or corpuscles (erythrocytes) and a smaller, but very important number of white blood
cells (leucocytes), these are the soldiers and scavangers whose duty in large part is to fight
our battles within the body and destroy infectuous germs. The red cells average about 5
million per cubic millimeters in men and 4.5 million per cubic millimeter in women. So we
see that they are very small but they perform a wonderful work in sustaining the health of the
body. These cells are formed largely in the marrow of the bones, the bones being hollow and
the home of the marrow and a place for the protection of nerves. We are informed that we
receive a new supply of red corpuscles about every three or four weeks. The old being
delivered to the liver to be prepared for elimination. The corpuscles evidently work so hard
they wear out and become old in that length of time and have to be replaced by new ones
with about an equal number. Therefore a healthy condition of the blood and marrow is
extremely important. The red color of the blood is due to the hemoglobin. It is this which
gathers the oxygen in the lungs when we breathe and carries it to the waiting tissues to
replace the wornout parts. The blood, as previously stated, gathers the carbon dioxide, the
waste product of the tissues which is eliminated by the lungs. The blood is also the vehicle
for food for every part of the body. Our food is digested in the stomach and intestines and the
needed juices are carried by the blood to the various parts where they are needed. Here the
blood acts as if it were intelligent. It carries to the various parts and distributes the very
particles required by each part. For instance, if a bone is fractured, then the required
phosphates and carbonate of lime etc., needed for the mending of the fracture will be carried
to the afflicted part. If it is an injury to a part as delicate as the eye, then the special
nourishment will be delivered to that afflicted part. To a wound in the flesh, like a muscle,
the blood as a master builder will supply the elements to make the repair. It does not make a
mistake. So the blood both feeds the body, builds it up, carries off the wornout particles and
deposits them where they can be eliminated from the body.

The white blood cells (leucocytes) in the blood play a very important part both in the blood
and by passing through the walls of the veins to carry on their assigned duties of repair and in
battling infected parts. There are in the blood about 1 white corpuscle to 500 of the red in the
healthy body, or there are about 8,000 per cubic millimeter of the white. When the blood
count rises considerably beyond this the physician seeks for some serious infections and the
soldiers have been reinforced.

This is a very interesting study and this is a very brief and insufficient treatment, but it is
enough to show us, one and all, a few of the responsibilities of the blood stream which is
sufficient evidence to convince the humble of the fact that the whole plan and purpose of the
blood must have come by the providence of the Lord, our Creator. It could not have been by
chance.

The brain is the seat of power for the nervous system. It is divided into two sections, the
larger section being divided into two halves in some respects resembling the half of an
English walnut. This is called the cerebrum, and is divided into two hemispheres, right and
left. Behind and almost beneath this part of the brain is the ``little brain,'' called the
cerebellum which is also separated into two halves. Connected with these divisions of the
brain and running through all parts are tiny nerve fibers, nerve cells, and cavities called
ventricles. These ventricles contain a fluid called the cerebro-spinal fluid which also bathes
the outside of the brain, and is continuous with the central cavity of the spinal cord. The
entire brain is filled with nerve fibers and nerve cells. These nerve cells, with their fibers, are
the units of the nervous system. Some of the nerves carry impulses towards the nerve centers,
these are called afferent nerves; in others the impulses flow in the opposite direction and
these are called efferent nerves. There are other nerves which carry messages both ways.

Many of the nerves in the brain are joined in a cable of nerves called the spinal cord which
extends from the lowest part of the brain downward as far as the lower border of the first
lumbar vertebrae, where it ends in a blunt point. Its average length in a man is about 18
inches and its thickness about the size of the little finger. From there extending downward to
the ends of the first coccygeal vertebrae is a cord known as the filum terminate. These cords
serve as cables, between the brain and all parts of the body, the nerve fibers being grouped in
bundles according to their special functions. The brain is divided into departments and the
nerves extending from there have connection with the different parts of the body. For
instance, there are twelve pairs of nerves coming from the brain with the following functions:

(1) Oilafactor, subserving smell; (2) optic, nerve of vision; (3) oculimotor, supplying most of
the muscles which move the eyeball and the muscle which contracts the pupil; (4) nerve
supplying the muscle which turns the eyeball downwards and outwards; (5) trigeminal, nerve
supplying sensation to the face, etc., and to the muscles of mastication; (6) nerve supplying
the muscle which turns the eyeballs outwards; (7) facial, nerve supplying the muscles of the
face; (8) auditory, subserving hearing; (9) glossopharyngeal, a nerve of taste, also supplying
sensation to the inside of the throat and activating some muscles there; (10) vagus, or
wandering nerve, supplying the heart, lungs, stomach, and other viscera, etc.; (11) spinal
accessory, supplying muscles in the neck; (12) hypoglossal, supplying the muscles moving
the tongue. 3

In addition there are thirty-one pairs of spinal nerves, eight cervical, twelve dorsal or
thoracic, five lumbar, five sacral and one coccygeal. 4 The first cervical passes out above the
topmost vertebrae, the axis, the others pass between adjoining vertebrae, delivering their
messages and receiving them from every part of the body. The nerves of the body, its organs
and, tendons, fibers and muscles and the skin are all connected by these cables to some part
of the brain. When a nerve in any part is destroyed or loses its power to function, the
corresponding part of the body fails to function. For instance, a serious injury to the spine
may sever the connection with the lower limbs, or from the arms, or one side of the body.
When that happens paralysis sets in and the afflicted parts having no connection with the
brain, cease to function. A message from a burn on the hand is carried to a definite part of the
brain and back again almost instantly and there comes the sensation of pain. This sensation is
not in the brain but cabled to the afflicted part. Doctors are able, knowing the nature of the
nerves to temporarily paralize certain parts of the body in order to perform delicate
operations. This is but a brief statement of some of the functions of the brain and the nerves
which center in it. It is marvelous, and so complicated that none but a fool will think that it
came by chance and not by the directing hand of the Almighty.

We have considered but a few of the parts and functions of the parts of the body. Perhaps we
should mention another remarkable fact that indicates divine arrangement, and that is the fact
that some of these nerves serve our bodies independently of our will. Such is the case of the
heart, in a very large degree the lungs and all or most all of the vital organs of the body.
These parts are said to act involuntarily. For example, we breathe unconsciously even in our
sleep, the heart beats regularly in the healthy body independent of our command or will. The
blood flows on its mission without our giving it a thought, our food is taken care of and
digested in the stomach and intestines without our conscious thinking. We move the muscles
of our arm or leg by our own free will, we may take a deep breath on purpose, but we breathe
when we are not conscious of it either awake or asleep. To say that this is the result of chance
is also out of reason.

Let us consider another function of the brain. The brain is the vehicle, or medium of thinking
and there are those who, because of their opposition to God and their foolish theories in
opposition to the revelations from the Lord, try to make us think that the brain of itself has
this wonderful power. This is not true. The brain is merely the medium through which some
other power acts. The mere combination of carbon, calcium, oxygen or any other physical
elements, no matter how wonderfully they may be merged, can not have the power to make
us think. They have no power to make us love or hate or perform any other function of the
will. These elements are agencies, or conductors, through which a more subtle influence acts.
The mind is not the brain. No one has, by the aid of science ever been able to discover the
mind, to circumscribe it, to declare its length or breadth, or to weigh it. It has been to all most
elusive and yet they know it exists, and the brain is the seat of its operations.

It is a very strange thing that men can believe that the combination of certain elements can
exercise the power of thinking; that intelligence is a material thing, or the result of the
physical combination of earthly elements. That the mind came by chance and ``emerged''
because the creature, whatever it was supposed to be, as an ancestor of man needed it, is
foolish. I have read many of Anderson's and Grimm's fairy tales and enjoyed them knowing
that they were fairy tales and written for amusement, but they are far more reasonable with
all of their inconsistencies than these ridiculous teachings that have been accepted by
worldly-minded men, as one brother stated it, ``educated beyond their intelligence.'' The
mind, or intellect, with which we do our thinking is not composed of physical matter. What is
a thought? Does the scientist know from whence it comes? Can they tell us the source of our
emotions, our love, one for another, or for the things that are beautiful? Can they tell us why
we remember? Is it due to imprints on the brain, a sort of photographic gallery and library
wherein all such things are stored? The fact remains that scientists do not know. There is no
earthly way they can know, for all such things are intangible to the human faculties or any
instruments yet invented that can weigh or photograph them and reveal their shape or
dimensions. All they know, like all they teach about organic evolution is but a guess and
nothing more. They prate about the ``descent of man'' but they cannot fathom the mind. They
cannot tell why the heart ticks, a question that is frequently asked. They cannot tell us why
the involuntary action takes place in the body—the lungs, the heart, the blood, the digestive
organs, etc., keep going even when we sleep or a man may be unconscious. They may say it
is due to life; but they do not know the source of that life! This is the mystery to them for
which they have searched down through the ages and for which they still search and hope to
find.

Well, are these questions unanswerable? Not so, if we are willing to accept the word of him
who knows. Our Eternal Father has revealed them unto us, but when men reject him and treat
with contempt the knowledge he has revealed they will remain in darkness. We have
previously called attention to the fact that the Lord has made it known that \textit{all} men are given
the Spirit of Christ and that it is this Spirit that quickens their intellects, for the sake of clarity
I repeat:

For the word of the Lord is truth, and whatsoever is truth is light, and whatever is light is
Spirit, even the Spirit of Jesus Christ.

And the Spirit giveth light to every man that cometh into the world; and the Spirit
enlighteneth every man through the world, that hearkeneth to the voice of the Spirit.

And every one that hearkeneth to the voice of the Spirit cometh unto God, even the Father. 5

And the light which shineth, which giveth you light, is through him who enlighteneth your
eyes, which is the same light that quickeneth your understandings. 6

In him was life; and the life was the light of men.

And the light shineth in darkness; and the darkness comprehendeth it not. 7

It is from this Spirit that we obtain our understanding; it is this Spirit that quickens our mind,
that helps us to think, to gain knowledge, and even with those who pervert the ways of the
Lord and rebel against him, it is the light of truth or Spirit of Christ which they attempt to
prostitute and turn from its rightful channel. Where, then, do we search for the mind? It is not
a part of the mortal body although the mortal body—brain—is the agency through which it
works. The Mind is in the Spirit which inhabits the body, or the source from which comes
our thinking, our love of knowledge, our power to know and understand. It is the \textit{life}, for the
spirit is life and is eternal, and when the spirit leaves the body, the body is dead. It may
appear the same, it may weigh the same, to all appearances it has lost nothing of material
substance; but from that moment the various materials of which it is composed have a
tendency to separate, to decompose, the real \textit{substance}, the intelligent part, the part that made
it think and exercise its intelligence, has departed.

Why do men love darkness rather than light? The Lord has given us the answer.

\newpage
REFERENCES—CHAPTER ELEVEN

Footnotes

1. \textit{The New Modern Physician}, p. 706.

2. 2 Nephi 9:31-32.

3. \textit{The New Modern Physician}, p. 737.

4. \textit{Ibid}., p. 737.

5. D. \& C. 84:45-47.

6. \textit{Ibid}., 88:11.

7. John 1:4-5.

