\chapter{THE SCATTERING OF THE PEOPLE}

"SO the Lord scattered them abroad from thence upon the face of all the earth: and they left
off to build the city." 1

The organic evolutionists assume the position that man has descended through millions of
years from the very lowest forms of life through countless "missing links," the fossils of
which for a century, they have been seeking. In scattered places on the earth parts of skeletal
remains have been discovered which these advocates maintain were parts of bodies of
human, or semi-human beings. They have not, however, maintained a solid front. Many of
these bones are believed by some of the experts to be bones of apes or other animals. The
manner in which they were found also throws doubt upon their origin. Evidence has
previously been given in relation to these bones dug up in divers places, yet brought together
with the assumption that they were all of the same bodily structure. Considering the great
ages which they maintain these structural changes have been going on, it is astonishing that
so much difficulty is encountered in finding the evidences which they seek. Nothing so far
has been discovered that has not carried with it grave doubt even among the experts. The
discoveries are too few and too widely scattered, yet a few bones of a skull, a jaw, a tooth, or
a part of an arm or leg, seem to be sufficient on which to build an imaginary individual, and
to these imaginary beings such names are given as \textit{Pitheconthropus, Trinil Man, Neanderthal
Man, Dawn Man} and so forth. There are many eminent scientists who maintain that these
bones, if human at all, do not present evidence that they constitute a criterion by which a
whole race of intermediary creatures between the ape and \textit{Homo Sapiens}, or modern man,
can be shown.

If there were any truth in this hypothesis, then we should find all forms of life in the
multitudinous stages from the amoeba, or worm, through the "higher apes" to man. Not only
would they be found in the fossils but occupying places on the face of the earth which would
be teeming with them. It would be impossible \textit{not} to find them in animate existence. We
would then find these "higher apes," just short of emerging into men, burying their dead,
making fire, and possessing some of the simple knowledge which these theorists ascribe to
"primitive man." No such creatures can be found on the earth. Our friends frequently refer to
the Fuegians of South America, the Botocudos of Brazil, the American Indian and other
peoples including the South African, or Negro 2 as having developed far enough to be in the
"stone age." Every member of the Church knows, or should know, that the Fuegians, the
Botocudos and the American Indians (Lamanites) are descendants of a very superior race of
people. A people who had the Gospel of Jesus Christ and who were enlightened by its
principles so that they were intelligent enough to live together in peace and harmony, without
"envyings, nor strifes, nor tumults, nor whoredoms, nor lyings, nor murders, nor any manner
of laciviousness," for nearly two hundred years. 3 Then they ceased to serve the Lord and
turned to wickedness. This resulted in a very short time in their spiritual light going out.
From their exalted state they fell to the pitiable condition in which Mr. Darwin and others
found them. Bible students know that the African is descended from an enlightened ancestry,
men who were skilled in the arts, in mechanics and music. 4 It was sin that brought them to
the sad condition of degradation. The natives of the Congo, "who pounds his tomtom to scare
away the god that he fears," and the American Indian, "with the incantations, the totem poles,
and the medicine man," are not "primitive" in the sense in which Dr. Millikan pictures them.
5 The original man was not a savage. He did not evolve out of lower forms of life. He was a
son of God! This has been previously discussed. He and his children in the beginning were
intelligent and worshipers of the living God, and were made acquainted with the mission of
Jesus Christ, who was to come. They were not savages, they were not "primitive" except in
the sense of being the earths first inhabitants. To confound the term "primitive" with
savagery and ignorance, is a display of ignorance, and is not justified.

If the story of evolution were true, then, I repeat, the earth today would be filled with living
creatures in all these stages of evolutionary development. We would not be able to travel a
mile without discovering some of them, for if the factor of increase was as great as these
scientists tell us it was, in the millions of years the earth would have been over-populated.
Under such circumstances, were this hypothesis true, scientists could save their time
searching from one end of the earth to the other to find in fossils the evidence which they
seek. It would be before them constantly in living forms inhabiting the earth in great
abundance.

Sir Ambrose Fleming, British scientist and mathematician, has presented some very
interesting figures on this question. In his very excellent work, \textit{The Origin of Man}, pages 84-
86, he has given us the following important information:

If the population at any place always increases by the same percentage of its amount at the
time, then it grows according to the same law by which money increases when invested at
compound interest—that is, when the interest as it accrues is added to the capital. Under this
law the money doubles itself in a certain time. Thus if a certain sum, say (1)100, is deposited
at 5 per cent interest per annum in a bank and left to accumulate at compound interest, it
would become (1)200 in about fourteen years. Approximately some such law governs the
growth of population, only the annual rate of increase in percentage is not constant. Thus
from the article "Population" in \textit{The Encyclopaedia Britannica}, 14th edition, we learn that in
A.D. 1928 the world population was estimated at 1879 millions. In A.D. 1920, it was 1811
millions, and in A.D. 1845 it was 1009 millions. It is now, A.D. 1945, probably near to 1940
millions. It has therefore nearly doubled in ninety years. We have no exact data for the world
population 1900 years ago, but certain not improbable reckonings would make it about 70 to
80 millions. It has therefore doubled several times over in the last 1900 years. But the span of
time in which it has doubled in that past 1900 years has been much larger than at present. If
we take it as likely that the world population at the present time is doubling in 100 years,
then it can be shown that on the average the time of doubling has been four or five times as
long in the period of time between, say, A.D. 45 and A.D. 1845.

In order to give time for the evolution of the human race from animal progenitors, the
evolutionists are obliged to suppose an immense time period, a million years or more, for the
process. The animal brain or the monkey-like paw cannot be changed by natural selection
into the human brain or hand in a few generations. It must take vast ages, and hence each
fragment of fossil "man" that is found is declared to be of great age, reckoned in hundreds of
thousands of years. But there is an insuperable difficulty connected with such suppositions.

Supposing that the human race has been on this earth even for a period of a hundred thousand
years and multiplying only at such a slow rate that the human population took 1000 years to
double, then it can easily be shown that the progeny of even one single pair of human beings
in the course of the hundred thousand years would have amounted to a number only
expressed by 30 digits. It would be a million million billion at least. There would not be
standing room for them on the earth at present. To produce the present population of 2000
million from one couple in 100,000 years it would be necessary for the rate to increase to be
so slow that it took on the average 3,330 years to double in number. Still more slowly would
it have to increase if we take the space of human existence to be a million years. But if this
very slow average rate of increase is to take place it invariably implies that the birthrate must
be very small or the death-rate very large. But both these conditions are antagonistic to
evolutionary progress.

The Darwinian natural selection demands a copious birth-rate, because it is not every germ or
embryo which possesses by chance some advantageous improvement fitting it more perfectly
to live in its environment. Moreover, if it has such evolutionary improvement it must have a
large chance to live and propagate its advance, because it cannot procreate in infancy. But a
high death-rate is not favorable to such survival of the fittest. Accordingly the evolutionists
are in this position:

The conditions they demand in order that some form of anthropoid ape or other mamal may
be transformed into a man by Darwinian natural selection are quite inconsistent with the state
of the world at present as regards human population. We have no justification for assuming
some vague process called Evolution if the definite conditions required for Darwinian natural
selection do not hold good. The outcome therefore of the inquiry as made in this chapter
must be that the evolution of man from an ape-like or other animal ancestor is not proven. An
increasing number of quite competent naturalists are adopting this opinion.

The lay reader need not, therefore, have his or her faith in the Creative origin of mankind
disturbed by over-confident assertions that all scientific men now accept the evolution of
mankind from an animal stock as definitely proved.

This argument and evidence is simple and complete and a thorough refutation of the organic
evolutionary theories. Moreover, it is an historical fact that every historian knows, life is
traced back to the Mesopotamia as the starting point. The story in the Book of Genesis states
that the "ark rested in the seventh month, on the seventeenth day of the month, upon the
mountain of Ararat." From there Noah and his sons made their way towards the fertile valley
to the south where their descendants began to build cities and cultivate the land. The oldest
cities known were located in the Mesopotamia and are mentioned in Genesis as Babel, Erech,
Accad, Calneh, Nineveh, Hur, or Ur, the home of Abraham's father. It was at, or near, Babel
that the confusion of the language of the people took place and from that place the Lord
scattered them to all parts of the earth. One of the earliest countries beyond these borders
where settlers immigrated was Egypt and this land was first settled by the daughter of Ham,
so the time of its settlement goes back close to the days when Noah came forth from the ark.
We read in the Book of Abraham:

Now this king of Egypt was a descendant from the loins of Ham, and was a partaker of the
blood of the Canaanites by birth.

From this descent sprang all the Egyptians, and thus the blood of the Canaanites was
preserved in the land.

The land of Egypt being first discovered by a woman, who was the daughter of Ham, and the
daughter of Egyptus, which in the Chaldean signifies Egypt, which signifies that which is
forbidden.

When this woman discovered the land it was under water, who afterwards settled her sons in
it; and thus, from Ham, sprang that race which preserved the curse in the land.

Now the first government of Egypt was established by Pharaoh, the oldest son of Egyptus,
the daughter of Ham, and it was after the manner of the government of Ham, which was
patriarchal.

Pharaoh, being a righteous man, established his kingdom and judged his people wisely and
justly all his days, seeking earnestly to imitate that order established by the fathers in the first
generations, in the days of the first patriarchal reign, even in the reign of Adam, and also of
Noah, his father, who blessed him with the blessings of the earth, and with the blessings of
wisdom, but cursed him as pertaining to the Priesthood.

Now, Pharaoh being of that lineage by which he could not have the right of Priesthood,
notwithstanding the Pharoahs would fain claim it from Noah, through Ham, therefore my
father was led away by their idolatry. 6

This information which we have received from the writings of Abraham, throws a great deal
of light on ancient history. Nearly all ancient histories begin with Egypt as being, perhaps,
the oldest of the nations. This comes to pass through the discoveries which in the early part
of the nineteenth century, made it possible for the savants to learn the Egyptian writings and
thus interpret much of their history and traditions. In the days of Abraham, communications
between Chaldea, the former home of Abraham, and Egypt was common, in fact they were
under the same governing power at the time. It is also a conceded fact that civilization in
ancient times spread from the valleys of the Mesopotamia, which by interpretation means,
the land between the two rivers. These two regions, the Mesopotamia and Egypt, were two of
the most fertile spots of ancient times. And here two civilizations were established. All
known history centers in the Mesopotamia and it was from here the Lord scattered the
nations. Professor Willis Mason West, in his \textit{Ancient History}, says: "The first homes of
civilization were in the lower valley of the Nile and the Euphrates. In each of these regions a
cheap food supply made possible at an early date a dense population, with a leisure and
military class supported by the agricultural masses. In both districts, too, at a still earlier
time, the marvelously fertile soil attracted enterprising tribes from different sources, and so
brought about a mixture of races—apparently a condition favorable to progress."

George Rawlinson wrote:

That the great alluvian plain at the mouth of the Euphrates and Tigris was among the
countries first occupied by man after the Deluge, is affirmed by Scripture, and generally
allowed by writers upon ancient history. Scripture places the original occupation at a time
when language had not yet broken up into its different forms, and when, consequently, races,
as we now understand the term, can scarcely have existed. It is not, however, into the
character of these primeval inhabitants that we have here to inquire, but into the ethnic
affinities and characteristics of that race, whatever it was, which first established an
important kingdom in the lower part of the plain—a kingdom which eventually became an
empire. 7

Both history and revelation confirm the fact that it was from this point that mankind began to
spread over the face of the earth following the days of Noah. Contrary to the general view in
the world, Noah built his ark and commenced his voyage in the flood from that part of the
earth which is now known as the United States of America. All the land surface of the earth
at that time was in one place and it was not until some time after the flood that the land
surface of the earth was divided. However, the journey from the western part of the world as
we know it today to Armenia was a very great journey. The first and oldest civilization was
destroyed and it was very largely occupying this western hemisphere. The peopling of the
earth commenced anew through the family of Noah about 1600 years from the fall of Adam.
The history of nations reveals the fact that 3000 years ago the earth was sparsely inhabited.
At the time of the Revolutionary War, there were, perhaps, less inhabitants in all of Europe
and western Asia than are within the borders of the United States today.

\newpage
REFERENCES—CHAPTER TWENTY-THREE

Footnotes

1. Gen. 11:8.

2. Darwin, Charles, \textit{Descent of Man}, pp. 143-144.

3. 4 Nephi 16.

4. Gen. 4:21-22.

5. Millikan, Dr. R. A., \textit{Evolution in Science and Religion}, p. 66.

6. Abraham 1:21-27.

7. Rawlinson, George, \textit{The Seven Great Monarchies}, Vol. 1, p. 28.

