\chapter{CONFLICT BETWEEN SCIENCE AND RELIGION—1}

IS there a conflict between religion and science? Between the revealed word of the Lord to
his prophets and the truth revealed through scientific investigation, there is no conflict. There
can be no disagreement wherein truth is arrayed against truth. Could such a thing be possible
all things would perish in confusion. There is a conflict existing between revealed truth
coming from the Lord to his chosen servants and the false doctrines advocated by men of
science. There is also a conflict between false religion and truth revealed through scientific
investigation. The time will come when nothing will remain except truth. The Lord has made
this known through revelation to the Prophet Joseph Smith as follows:

Yea, verily I say unto you, in that day when the Lord shall come he shall reveal all things—

Things which have passed, and hidden things which no man knew, things of the earth, by
which it was made, and the purpose and the end thereof—

Things most precious, things that are above, and things that are beneath, things that are in the
earth, and upon the earth, and in heaven. 1

Again:

I am the Lord thy God; and I give unto you this commandment—that no man shall come unto
the Father but by me or by my word, which is my law, saith the Lord.

And everything that is in the world, whether it is ordained of men, by thrones, or
principalities, or powers, or things of name, whatsoever they may be, that are not by me or by
my word, saith the Lord, shall be thrown down, and shall not remain after men are dead,
neither in nor after the resurrection, saith the Lord your God.

For whatsoever things remain are by me; and whatsoever things are not by me shall be
shaken and destroyed. 2

As long as men of science ignore the light of truth and have no faith in the Divine Creator,
they will search the hard way to find out the works of the Lord and will formulate false
theories which may prove both harmful and pernicious in that they will guide their fellows
who accept them away from the revealed plan leading to eternal life.

Unfortunately most scientists depend entirely upon their own intelligence and wisdom
without a thought of divine aid. It is said repeatedly that scientists do not take God into their
reckoning in the search of truth. This is an unfortunate condition, for if they were men of
prayer and faith, seeking divine help they would come to the truth more readily and would
avoid the many pitfalls of false hypotheses.

The Church, nevertheless, is very grateful to the wonderful men and women who have been
benefactors of humanity in a thousand ways, perhaps unconscious of the fact that divine aid
has been given them, even if they did not seek it. Such as Harvey, Jenner, Lister, Pasteur, the
Curies, Morse, Tesla, Whitney, Farnsworth and thousands of others. It is foolish for one to
think, however, that these discoveries and inventions have come to light during the past few
centuries because men are more intelligent today than they were in former times. The poetry
of the Psalms, the wisdom of Proverbs, the admonition and teachings of the Old Testament
prophets, have never been surpassed. The painting and sculpture of the ancients equal or
surpass anything produced today. The Damascus steel has been the marvel of the world and
thousands of other things known and practiced by the ancients indicate an intelligence of the
highest quality. Dr. Robert A. Millikan, who has been deceived by some modern theories,
testifies that the ancients were as intelligent as men are today and that there have been
billions of people who have lived and died with "as high intelligence" as people manifest
today. 3

"If this is true," it has been said, "why, then, is it that the ancients did not have modern means
of travel, of communication, numerous inventions for their convenience, such as we have
today?" The answer to this question is, because the time had not come for these inventions
and discoveries to be revealed. Let us not forget that the Spirit of the Lord has played a great
part in all of the developments that have come to benefit mankind in these latter centuries.
Can there be an honest doubt that the ancient seers saw these things in vision? They wrote
about them and have predicted their coming; but the Almighty withheld from them these
benefits for his own purpose. It is inconsistent to think that the Son of God, who by
command of his Father organized the earth on which we stand, who could turn water into
wine, raise the dead, give eyesight to the blind, cleanse the lepers, command the elements
and have them obey him, and then could not have revealed, had it pleased him, these modern
inventions. Nahum without question saw our present means of transportation. 4 John on the
Isle of Patmos without doubt saw our airplanes and described them as they are now being
used in war. It is revealed in the Book of Mormon that the Brother of Jared saw our day; 5
and the Lord revealed to Enoch the history of this earth and its inhabitants down to the end of
time. 6 Isaiah also must have seen these modern conditions when he speaks of the gathering
of Israel and that those being gathered would "fly as a cloud, and as the doves to their
windows." 7 Joel declared that in the last days the Lord would pour out his spirit upon all
flesh; and "your sons and your daughters shall prophesy, your old men shall dream dreams,
your young men shall see visions: and also upon the servants and upon the handmaids in
those days will I pour out my spirit." 8 We are living in that day.

The reason for all of this great activity which was denied the ancient peoples is that we are
living in the day of restitution; a day in which preparation shall be made to bring the earth
back again to its primitive glory as it was when the Lord pronounced it good. It is the day
spoken of by Peter as a time of "refreshing," and restitution of all things spoken of by all the
prophets. 9 The same that was prophesied of by Paul in his epistle to the Ephesians, that "in
the dispensation of the fulness of times he might gather together in one all things in Christ,
both which are in heaven, and which are on earth; even in him." 10

Henry George Bohn, printer and historian, writing of the \textit{Origin and Progress of Printing},
wrote as follows:

"Nature does not advance by leaps," says an old proverb; neither does her offspring, Art. All
the great boons vouchsafed to man by a munificent providence are of gradual development;
and though some may appear to have come upon us suddenly, reflection and inquiry will
always show that they have had their previous stages.

Indeed, nothing in this great world which concerns the well-being of man takes place by
accident, but is brought forward by divine will, precisely at the moment most suitable to our
condition. So it was with astronomy, the mariner's compass, the steamengine, gas, the
electric telegraph, and many other of those blessings which have progressed with civilization.
The elements were there and known, but the time had not arrived for their fructification. 11

Daniel in his vision of the last days saw a "time of trouble" such as "never was since there
was a nation even to that same time." He saw the time of the redemption of the righteous, and
the day of the resurrection, when the dead should come forth, "some to everlasting life, and
some to shame and everlasting contempt." It was to come in a day when "many should run to
and fro, and knowledge shall be increased." 12 Who can deny that this day is here? Paul also
saw our time and described the conditions accurately as they now are. In writing to Timothy
he said:

This know also, that in the last days perilous times shall come.

For men shall be lovers of their own selves, covetous, boasters, proud, blasphemers,
disobedient to parents, unthankful, unholy.

Without natural affection, trucebreakers, false accusers, incontinent, fierce, despisers of those
that are good,

Traitors, heady, highminded, lovers of pleasure more than lovers of God;

Having a form of godliness, but denying the power thereof: from such turn away. 13

In writing to the members of the Church at Thessaly, Paul also predicted that because men
would turn away from the truth the Lord, "Shall send them strong delusion, that they should
believe a lie: That they all might be damned who believe not the truth, but had pleasure in
unrighteousness." 14

What the prophets predicted has come to pass. Enoch, Moses, Isaiah, Daniel, Peter, Paul, the
Jaredite and Nephite prophets, told the story of man's turning away from God, boasting in his
own wisdom and his denial of the truth as truly and clearly by prophecy thousands of years
ago, as we see it fulfilled today. We live in an age of materialism; of unbelief in the power of
God, and denial of the divine calling of Jesus Christ as the Redeemer of the world and the
Only Begotten Son of God in the flesh. This reminds us of the words of Jacob, brother of
Nephi:

O that cunning plan of the evil one! O the vainness, and the frailties, and the foolishness of
men! When they are learned they think they are wise, and they hearken not unto the counsel
of God, for they set it aside, supposing they know of themselves, wherefore, their wisdom is
foolishness and it profiteth them not. And they shall perish.

But to be learned is good if they hearken unto the counsels of God.

But wo unto the rich, who are rich as to the teachings of the world. For because they are rich
they despise the poor, and they persecute the meek, and their hearts are upon their treasures;
wherefore, their treasure is their god. And behold, their treasure shall perish with them also.

And wo unto the deaf that will not hear; for they shall perish.

Wo unto the blind that will not see; for they shall perish also.

Wo unto the uncircumcised of heart, for a knowledge of their iniquities shall smite them at
the last day.

Wo unto the liar, for he shall be thrust down to hell.

Wo unto the murderer who deliberately killeth, for he shall die.

Wo unto them who commit whoredoms, for they shall be thrust down to hell.

Yea, wo unto those who worship idols, for the devil of all devils delighteth in them.

And in fine, wo unto all those who die in their sins; for they shall return to God, and behold
his face, and remain in their sins. 15

There is great divergence between the path traversed by the scientific investigator, who
labors without faith in God, and that traveled by those who accept the divine revelations and
walk in the light of his revelations. The scientist, searching without divine aid, depending
solely on his own research and wisdom, reaches certain conclusions in relation to the origin
of things; how worlds come into being and how they are eventually to die, perhaps to
disintegrate and be recreated again in one eternal round. These theories are uncertain; they
change from time to time, but quite generally scientists all agree that all things must
eventually come to an end and then, some of them believe, the recreation will take place
again. They like to talk about the "dying sun" 16 and have written volumes about the
Universe how it came into being by chance and is subject to the temporal laws which we
mortals understand and by which we are governed, and therefore, according to their
calculations based on things which they see naturally, they conclude that all things in the
Universe are governed by the same mutable laws that bind us to this mundane sphere. The
Lord said to Isaiah:

. . . Forasmuch as this people draw near me with their mouth, and with their lips do honour
me, but have removed their heart far from me, and their fear toward me is taught by the
precept of men:

Therefore, behold, I will proceed to do a marvellous work among this people, even a
marvellous work and a wonder: for the wisdom of their wise men shall perish, and the
understanding of their prudent men shall be hid. 17

By revelation direct from our Eternal Father, we learn a very different story in relation to the
beginning of things and the end thereof. There is a Supreme Ruler in the Universe, the
Creator of worlds and all things that are in them. He does not create worlds to be destroyed.
They are created to bring to pass his divine purpose for he said his work and glory is, "To
bring to pass the immortality and eternal life of man." 18 Life is not an accident 19 and the
human race is not destined to "die and cease to be." 20 Neither is life "an unimportant by-
product . . . off the main line." 21

The story told by revelation has a very different and more hopeful ring. The Lord has not
created worlds by accident, they are not "wandering blindly through space." 22 The Universe
is well organized according to divine law and it is revealed that—

All kingdoms have a law given;

And there are many kingdoms; for there is no space in the which there is no kingdom; and
there is no kingdom in which there is no space, either a greater or a lesser kingdom.

And unto every kingdom is given a law; and unto every law there are certain bounds also and
conditions.

All beings who abide not in those conditions are not justified. 23

This applies to the great stars traveling through the Universe. It applies to our solar system
and to every other kingdom great or small, including the atom. The Lord creates worlds in a
systematic manner. He places them throughout the Universe by perfect and infinite
calculations. It is not a matter of chance but of divine order. Earths are created primarily as
habitations for living creatures. Man is superior to an earth, for he is the offspring of God,
and he will, if he is obedient to eternal divine law, become a son of God and possess the
powers by which earths are created. Our Father has said that he has created worlds "without
number," and this for his own purpose, and "as one earth shall pass away, and the heavens
thereof, even so shall another come:" and there is no end to his works, neither to his words.
24 The destiny of the earth on which we dwell, is that it shall pass through the measure for
which it was created; it will eventually die, for it is a living thing, and after it is raised in the
resurrection which will come to it, it will be a celestial body which will shine like the sun,
and become the abode of celestial beings, that is, of those who have proved faithful in
keeping all the laws of God. 25 The course prepared for our earth is the course through which
countless other earths have passed and gone on to their glory, and as countless others, like
our earth, are preparing to do.

Our heavens are filled with eternal worlds that have attained their exaltation. They are not
dying. They have passed through the stage of death and have become eternal, glorious
bodies, the habitations of sons and daughters of God who likewise have attained to glory.
This is not a fairy tale but the story as the Lord has revealed it to his servants the Prophets.
Unfortunately the wise and learned after the manner of the world, will not believe it. They
prefer to believe the uncertain and ever-changing theories so prevalent among uninspired
men, who travel blindly trying to fathom the deep mysteries of the Universe which elude
them, but which the Lord has revealed to those who believe. These glorious truths the Lord
has made known. Here are a few passages of scripture:

I said, O my God, take me not away in the midst of my days: thy years are throughout all
generations.

Of old hast thou laid the foundation of the earth: and the heavens are the work of thy hands.

They shall perish, but thou shalt endure: yea, all of them shall wax old like a garment; as a
vesture shall thou change them, and they shall be changed:

But thou art the same, and thy years shall have no end.
 
The children of thy servants shall continue, and their seed shall be established before thee.
(Psalms 102:24-28.)

Lift up your eyes to the heavens, and look upon the earth beneath: for the heavens shall
vanish away like smoke, and the earth shall wax old like a garment, and they that dwell
therein shall die in like manner: but my salvation shall be for ever and my righteousness shall
not be abolished. (Isaiah 51:6.)

And, Thou, Lord, in the beginning hast laid the foundation of the earth; and the heavens are
the works of thine hands:

They shall perish; but thou remainest; and they all shall wax old as doth a garment. (Hebrews
1:10-11.)

But the day of the Lord will come as a thief in the night; in the which the heavens shall pass
away with a great noise, and the elements shall melt with fervent heat, the earth also and the
works that are therein shall be burned up.

Seeing then that all these things shall be dissolved, what manner of persons ought ye to be in
all holy conversation and godliness,

Looking for and hasting unto the coming of the day of God, wherein the heavens being on
fire shall be dissolved and the elements shall melt with fervent heat? (2 Peter 3:10-12.)

The Lord himself called attention to these things when discussing his second coming with his
disciples, said he, "Heaven and earth shall pass away, but my words shall not pass away."
(Matthew 24:35.)

This passing away does not mean that the earth will cease to be. It is merely the cleansing
process from the temporal condition to the eternal condition where death will come to an end.
To the Prophet Joseph Smith the Lord made this matter perfectly clear in the following
words:

And again, verily, verily, I say unto you that when the thousand years are ended, and men
again begin to deny their God, then will I spare the earth but for a little season;
And the end shall come, and the heaven and the earth shall be consumed and pass away, and
there shall be a new heaven and a new earth.

For all old things shall pass away, and all things shall become new, even the heaven and the
earth, and all the fulness thereof, both men and beasts, the fowls of the air, and the fishes of
the sea;

And not one hair, neither mote, shall be lost, for it is the workmanship of mine hand. 26

Again:

And the redemption of the soul is through him that quickeneth all things, in whose bosom it
is decreed that the poor and the meek of the earth shall inherit it.

Therefore, it must needs be sanctified from all unrighteousness, that it may be prepared for
the celestial glory;

For after it has filled the measure of its creation, it shall be crowned with glory, even with the
presence of God the Father;

That bodies who are of the celestial kingdom may possess it forever and ever; for, for this
intent was it made and created, and for this intent are they sanctified. . .

And again, verily I say unto you, the earth abideth the law of a celestial kingdom, for it filleth
the measure of its creation, and transgresseth not the law—

Wherefore, it shall be sanctified; yea, notwithstanding it shall die, it shall be quickened again,
and shall abide the power by which it is quickened, and the righteous shall inherit it. 27

Latter-day Saints should be extremely grateful for this light which the Lord has revealed to
them concerning their destiny and the destiny of the earth. Let us add this comment from
President Brigham Young:

And this world, so benighted at present, and so lightly esteemed by infidels, as observed by
Brother Clements, when it becomes celestialized, it will be like the sun, and be prepared for
the habitation of the saints, and be brought back into the presence of the Father and the Son.
It will not then be an opaque body as it now is, but it will be like the stars of the firmament,
full of light and glory: It will be a body of light. John compared it, in its celestialized state to
a sea of glass. 28

This from Elder Orson Pratt:

Who, in looking upon the earth as it ascends in the scale of the Universe does not desire to
keep pace with it? that when it shall be classed in its turn, among the dazzling orbs of the
blue vault of heaven, shining forth in all the splendors of celestial glory, he may find himself
proportionably advanced in the scale of intellectual and moral excellence? Who, but the most
abandoned, does not desire to be counted worthy to associate with those higher orders of
beings who have been redeemed, exalted, and glorified together with the worlds they inhabit,
ages before the foundations of our earth were laid? O man, remember the future destiny and
glory of the earth, and secure thine everlasting inheritance upon the same, that when it shall
be glorious, thou shalt be glorious also. 29

\newpage
REFERENCES—CHAPTER ONE

Footnotes

1. D. \& C. 101:32-34.

2. D. \& C. 132:12-14.

3. Millikan, Dr. R. A., \textit{Evolution in Science and Religion}, p. 34.

4. Nahum 2:3-4.

5. Ether 3:25.

6. Moses 7:47-67.

7. Isaiah 60:8.

8. Joel 2:28-29.

9. Acts 3:18-21.

10. Ephesians 1:10.

11. \textit{Great Events by Famous Historians}, Vol. 8:3.

12. Daniel 12:1-4.

13. 2 Timothy 3:1-5.

14. 2 Thessalonians 2:10-12.

15. 2 Nephi 9:28-38.

16. Jeans, Sir James, \textit{The Mysterious Universe}, pp. 1-16.

17. Isaiah 29:13-14.

18. Moses 1:39.

19. Jeans, Sir James, \textit{The Mysterious Universe}, pp. 4, 12.

20. \textit{Ibid.}, p. 14.

21. \textit{Ibid.}, p. 6.

22. \textit{Ibid.}, p. 4

23. D. \& C. 88:36-39.24. Moses 1:38.

25. D. \& C. 88:15-29

26. D. \& C. 29:22-25.

27. D. \& C. 88:17-26.

28. \textit{Journal of Discourses}, Vol. 7:163.

29. \textit{Millennial Star} 12:72.

