\chapter{THE CONFUSION OF TONGUES}

``AND the whole earth was of one language, and of one speech.

``And it came to pass, as they journeyed from the east, that they found a plain in the land of
Shinar; and they dwelt there.

``And they said one to another, Go to, let us make brick, and burn them throughly. And they
had brick for stone, and slime had they for morter.

``And they said, Go to, let us build us a city and a tower, whose top may reach unto heaven;
and let us make us a name, lest we be scattered abroad upon the face of the whole earth.

``And the Lord came down to see the city and the tower, which the children of men builded.

``And the Lord said, Behold, the people is one, and they have all one language; and this they
begin to do: and now nothing will be restrained from them, which they have imagined to do.

``Go to, let us go down, and there confound their language, and they may not understand one
another's speech.

``So the Lord scattered them abroad from thence upon the face of all the earth: and they left
off to build the city.

``Therefore is the name of it called Babel; because the Lord did there confound the language
of all the earth: and from thence did the Lord scatter them abroad upon the face of all the
earth. 1

In the revision by the Prophet Joseph Smith it reads as follows:

And the whole earth was of the same language, and of the same speech. And it came to pass,
that many journeyed from the east, and as they journeyed from the east, they found a plain in
the land of Shinar, and dwelt there in the plain of Shinar.

And they said one to another, Come, go to, let us make brick, and burn them thoroughly. And
they had brick for stone, and they had slime for mortar.

And they said, Come, go to, let us build us a city, and a tower whose top will be high, nigh
unto heaven; and let us make us a name, lest we be scattered abroad upon the face of the
whole earth.

And the Lord came down, beholding the city and the tower which the children of men were
building;

And the Lord said, Behold, the people are the same, and they all have the same language; and
this tower they begin to build, and now, nothing will be restrained from them, which they
have imagined, except I, the Lord, confound their language, that they may not understand one
another's speech. So, I, the Lord, will scatter them abroad from thence, upon all the face of
the land, and unto every quarter of the earth.

And they were confounded, and left off to build the city, and they hearkened not unto the
Lord, therefore, is the name of it called Babel, because the Lord was displeased with their
works, and did there confound the language of all the earth; and from thence did the Lord
scatter them abroad upon the face thereof. 2

This story of the confounding of tongues and the building of the tower, is today almost
universally disbelieved. It is classed among the folklore and mythology of the ancient
peoples. Textbooks on sociology, all linguistic studies and history, ridicule, or ignore
entirely, this history of the confounding of tongues. There are many theories in the world as
to the origin of speech. Some of these have been mentioned previously and will not be
enlarged upon at this point. 3 Out of this interesting account of the confounding of speech
opponents of the Bible and its scientific critics think they have given the Holy Scriptures a
vital blow, and from this story they get great amusement. For instance, Dr. Andrew D. White
seems greatly amused that ``God came down'' to see what men were doing, for the God of the
Old Testament could not know what was going on among mankind without coming down to
see it. Moreover, he attempts to refute the story on the ground that ``the Egyptian language
had even before the invention of this story been developed in all essential particulars to the
highest point it ever attained,'' 4 and Dr. John W. Draper treats this subject in the same spirit
of unbelief. 5 It is unnecessary to go into details in relation to the criticisms and countless
comments that have been published in ridicule of the story of the confounding of tongues. As
previously stated this Bible account is almost universally rejected, and there is not to be
found today a textbook on sociology or ancient history that does not treat the origin of man
and the gradual development of his language from the evolutionary point of view; and the
story of the confusion of tongues is regarded as a fable or ignored with silent contempt. We
will therefore present the evidence as revealed to the Church of Jesus Christ of Latter-day
Saints.

In the translation, or divine revision, of the Book of Genesis by the Prophet Joseph Smith, the
story of the flood and the confusion of tongues are recorded with some corrections of the
Bible account. In other words, many of the parts that originally were in the writings of Moses
and which were eliminated in course of time, have been restored. The Lord, in these
restorations, placed the stamp of approval on the story of Babel and the scattering of the
people after the confusion of their language. We have the evidence that this story of Babel is
true and is written in the history of the Jaredites as recorded in the Book of Mormon.

On the title page of every copy of the Book of Mormon are two brief paragraphs that were
written by Moroni as he closed the abridgment on the plates of Mormon. He says:

An abridgment taken from the Book of Ether also, which is a record of the people of Jared,
who were scattered at the time the Lord confounded the language of the people, when they
were building a tower to get to heaven—Which is to show unto the remnant of the House of
Israel what great things the Lord hath done for their fathers; and that they may know the
covenants of the Lord, that they are not cast off forever—And also to the convincing of the
Jew and Gentile that JESUS is the CHRIST, the ETERNAL GOD, manifesting himself unto
all nations.

Fortunately we have the abridged record of these ancient Jaredites and the account of their
departure from the valley of Shinar at the time of the confusion of tongues. The members of
the Church are, therefore, not left to flounder, criticize, and perhaps be led astray, by the
modern philosophies and criticisms of sacred things which is so prevalent in the world today.
We are highly favored of the Lord in that he has given us many things for our eternal benefit
and to strengthen our faith in his divine purposes. All of this the people of the world may
receive if they will repent. They have hardened their hearts and with the exception of a few,
have refused to believe the pleadings and instruction which has been offered them. They,
seemingly, prefer to believe the uninspired notions of men.

Ether was the last of the historians of the Jaredite race. His position was very similar to that
of Moroni. Each was left to record the final destruction of his people. It fell to the lot of
Moroni to give us an abridgment of the writings of Ether. The finding of the record of the
Jaredites is given in the writings of Omni (Omni 20-23.) as follows:

And it came to pass in the days of Mosiah, there was a large stone brought unto him with
engravings on it; and he did interpret the engravings by the gift and power of God.

And they gave an account of one Coriantumr, and the slain of his people. And Coriantumr
was discovered by the people of Zarahemla; and he dwelt with them for the space of nine
moons.

And it also spake a few words concerning his fathers. And his first parents came out from the
tower, at the time the Lord confounded the language of the people; and the severity of the
Lord fell upon them according to his judgments, which are just; and their bones lay scattered
in the land northward.

The finding of Coriantumr is the only human contact recorded between the Jaredites and their
successors, and this was in fulfillment of the prophecy Ether made to Coriantumr who was
the Jaredite king, which is as follows:

And in the second year the word of the Lord came to Ether, that he should go and prophesy
unto Coriantumr that, if he would repent, and all his household, the Lord would give unto
him his kingdom and spare the people—

Otherwise they should be destroyed, and all his household save it were himself. And he
should only live to see the fulfilling of the prophecies which had been spoken concerning
another people receiving the land for their inheritance; and Coriantumr should receive a
burial by them; and every soul should be destroyed save it were Coriantumr. 6

Coriantumr did not repent and all that Ether predicted came upon him. During his nine
months sojourn among the people of Mulek evidently he learned to converse with them and
inform them of some of the history of the Jaredites. However, that may be, in the Book of
Mormon we have all of the necessary information to confirm the story of the confounding of
tongues as given in the Bible and the scattering of the people. There is another record, far
more important than the one written on the stone. This is the record of Ether written on gold
plates, twenty-four in number. They were discovered by a group of forty-three men sent out
by the colony of Nephites that had gone forth to live in Lamanite territory where they were
sorely oppressed. These men were sent to get help from the main body of Nephites in
Zarahemla. They lost their way and traveled many days, without doubt the Lord leading
them. They came to a place covered with bones and weapons of war that were partly
destroyed by the elements and while there they discovered the plates of Ether. They were
brought back and eventually were placed in the hands of King Mosiah, who was a prophet
and seer, who translated them. It is from these plates that Moroni made his abridgment.

The story as recorded by Moroni is an interesting one in full accord with the account of the
story of Babel in the Bible. While this interesting story is given in the Book of Ether, it will
not be out of place to state a portion of it in this writing:

And I take mine account from the twenty and four plates which were found by the people of
Limhi, which is called the Book of Ether.

And as I suppose that the first part of this record, which speaks concerning the creation of the
world, and also of Adam, and an account from that time even to the great tower, and
whatsoever things transpired among the children of men until that time, is had among the
Jews—

Therefore I do not write those things which transpired from the days of Adam until that time;
but they are had upon the plates; and whoso findeth them, the same will have power that he
may get the full account.

But behold, I give not the full account, but a part of the account I give, from the tower down
until they were destroyed.

And on this wise do I give the account. He that wrote this record was Ether, and he was a
descendant of Coriantor. 7

Then follows the genealogical record of Ether, the last prophet of the Jaredites, from
Coriantor back to Jared who with his brother, their families and friends, came from Babel at
the time of the confounding of tongues, and the record continues:

Which Jared came forth with his brother and their families, with some others and their
families, from the great tower, at the time the Lord confounded the language of the people,
and swore in his wrath that they should be scattered upon all the face of the earth; and
according to the word of the Lord the people were scattered.

And the brother of Jared being a large and mighty man, and a man highly favored of the
Lord, Jared, his brother, said unto him: Cry unto the Lord, that he will not confound us that
we may not understand our words.

And it came to pass that the brother of Jared did cry unto the Lord, and the Lord had
compassion upon Jared; therefore he did not confound the language of Jared; and Jared and
his brother were not confounded.

Then Jared said unto his brother: Cry again unto the Lord, and it may be that he will turn
away his anger from them who are our friends, that he confound not their language.

And it came to pass that the brother of Jared did cry unto the Lord, and the Lord had
compassion upon their friends and their families also, that they were not confounded.

And it came to pass that Jared spake again unto his brother, saying: Go and inquire of the
Lord whether he will drive us out of the land, and if he will drive us out of the land, cry unto
him whither we shall go. And who knoweth but the Lord will carry us forth into a land which
is choice above all the earth? And if it so be, let us be faithful unto the Lord, that we may
receive it for our inheritance.

And it came to pass that the brother of Jared did cry unto the Lord according to that which
had been spoken by the mouth of Jared.

And it came to pass that the Lord did hear the brother of Jared, and had compassion upon
him, and said unto him:

Go to and gather together thy flocks, both male and female of every kind; and also of the
seed of the earth of every kind; and thy families; and also Jared thy brother and his family;
and also thy friends and their families, and the friends of Jared and their families. 8

The record continues to say that all of this was accomplished. They also ``laid snares'' and
caught fowls of the air, and in vessels they carried fish, and they also carried swarms of bees
which they called ``deseret.'' Thus fully equipped they made their journey, crossing the ocean
in barges which they built according to the plan the Lord had given them.

These ancient records which were given to the Prophet Joseph Smith are true. We should be
very grateful to the Lord for them, for in them is given confirmation of many of the
prophecies and historical records of the far distant past. They confirm the Bible story of the
creation, the story of Adam and Eve, their fall and banishment from the garden, and the
doctrine of the atonement of Jesus Christ and his resurrection from the dead, far more clearly
than it is found in the Bible.

One of the grand purposes of the Book of Mormon is to be a witness for the Bible. The Lord
informed Nephi that it would come forth as ``one crying from the dust,'' as a witness for
Christ and for the record of the Jews—the Bible; and that it would come forth in a day when
men would be denying the prophecies of the Hebrew scriptures. We are able to see the
fulfillment of those promises and the Book of Mormon and the Doctrine and Covenants,—
the ``other books, which should come forth by the power of the Lamb, from the Gentiles unto
them, unto the convincing of the Gentiles and the remnant of the seed of my brethren, and
also the Jews who were scattered upon all the face of the earth, that the records of the
prophets and of the twelve apostles of the Lamb are true.'' 9

The fact that the traditions in relation to our first parents, the fall of man, the Garden of Eden,
the destruction of mankind in the flood and the building of the tower of Babel are found in
the traditional stories of most all nations is strong presumptive evidence that such events at a
very early time actually occurred. In June 1842, the Prophet Joseph Smith, who was editor of
the \textit{Times and Seasons}, published the following with the title:

TRAITS OF THE MOSAIC HISTORY FOUND AMONG THE AZTECA NATIONS

The tradition commences with an account of the deluge, as they had preserved it in books
made from the buffalo and deerskin, on which account there is more certainty than if it had
been preserved by mere oral tradition, handed down from father to son.

They begin by painting, or as we would say by telling us that Noah, whom they call Tezpi,
saved himself with his wife, whom they call Xochiquetzal, on a raft or canoe. Is not this the
ark? The raft or canoe rested on or at the foot of a mountain, which they call Colhuacan. Is
not this Ararat? The men born after this deluge were born dumb. Is not this the confusion of
language at Babel? A \textit{dove} from the top of a tree distributes languages to them in the form of
an olive leaf. Is not this the dove of Noah, which returned with the leaf in her mouth, as
related in Genesis? They say that on this \textit{raft}, besides Tezpi and his wife, were several
children, and animals, with grain, the preservation of which was of importance to mankind. Is
not this in almost exact accordance with what was saved in the ark with Noah, as stated in
Genesis?

When the Great Spirit, Tezcatlipoca, ordered the waters to withdraw, Tezpi sent out from his
raft a \textit{vulture}, which never returned, on account of the great quantities of dead carcasses
which it found to feed upon. Is not this the raven of Noah, which did not return when it was
sent out the second time, for the very reason here assigned by the Mexicans? Tezpi sent out
birds one of which was the humming bird; this bird alone returned, holding in its beak a
branch covered with leaves. Is not this the dove? Tezpi, seeing that fresh verdure now
clothed the earth, quitted his raft near the mountain of Colhuacan. Is not this an allusion to
Ararat of Asia? They say the tongues which the dove gave to mankind were infinitely varied;
which when received, they immediately dispersed. But among them there were 15 \textit{heads} or
\textit{chiefs} of families, which were permitted to speak the same language, and these were the
Taltecs, the Aculhucans and Azteca nations who embodied themselves together, which was
very natural, and traveled they knew not where, but at length arrived in the country of
Aztalan, or the lake country of America.

The plate or engraving presented here is a surprising representation of the deluge of Noah;
and of the confusion of the ancient languages at the building of the tower of Babel, as related
in the Book of Genesis, (see chapters vii and xi).

We have derived the subject of this plate from Baron Humboldt's volume of \textit{Researches in
Mexico}, who found it painted on a manuscript book, made of the leaves of some kind of tree,
suitable for the purpose, after the manner of the ancient nations of the sultry parts of Asia
around the Mediterranean.

Among the vast multitude of painted representations found by this author on the books of the
natives, made also frequently of prepared skins of animals, were delineated all the leading
circumstances and history of the deluge, of the fall of man, and of the seduction of the
woman by the means of the serpent, the first murder as perpetrated by Cain on the person of
his brother Abel.

The plate, however, here presented shows no more than a picture of the flood, with Noah
afloat on a raft, or as the traditions of some of the nations say on a tree, a canoe and some say
in a vessel of huge dimensions. It also shows by the group of men approaching the bird, a
somewhat obscure history of the confusion of the ancient language at the building of Babel,
by representing them as being born dumb, who receive the gift of speech from a dove, which
flutters in the branches of the tree, while she presents the languages to the mute throng, by
bestowing upon each individual a leaf of the tree, which is shown in the form of small
commas suspended from its beak.

Among the different nations, according to Humboldt, who inhabited Mexico, were found
paintings which represented the deluge, or flood of Tezpi. . . .

The painting of which the plate is the representation, shows Tezpi, or Noah, in the midst of
the waters laying on his back. The mountain, the summit of which is crowned by a tree and
rises above the waters is the peak of Colhuacan, the Ararat of the Mexicans. At the foot of
the mountain on each side appear the heads of Noah and his wife. The woman is known by
the two points extending up from her forehead, which is the universal designation of the
female sex among the Mexicans. The horn at the left hand of the tree with a human hand
pointing to it, is the character representing a mountain and the head of a bird placed above
the head of Tezpi or Noah, shows the vulture which the Mexicans say Tezpi sent out of his
acalli or boat to see if the waters had subsided.

In the figure of the bird with the leaves of a tree in his beak, is shown the circumstance of the
dove returning to the ark, when it had been sent out the second time bringing a branch of the
olive in its mouth; but in their tradition it had to become misplaced, and is made the author of
the languages. That birds have a language was believed by the nations of the old world. Some
of those nations retain a surprising traditional account of the deluge; who say that Noah
embarked in a spacious \textit{accalli} or boat, with his wife, his children, several animals, and grain,
the preservation of which was of great importance to mankind. When the Great Spirit,
Tezcatlipoca, ordered the waters to withdraw, Tezpi or Noah sent out from his boat a vulture.
But the bird's natural food was that of dead carcasses with which the earth now dried in some
places abounded.

Tezpi sent out other birds one of which was the humming bird; this bird alone returned again
to the boat, holding in its beak a branch covered with leaves. Tezpi now knowing that the
earth was dry, being clothed with fresh verdure, quitted his bark near the mountain
Colhuacan or Ararat. A tradition of the same fact, the deluge is also found among the Indians
of the Northwest. I received, says a late traveler, the following account from a chief of one of
the tribes in his own words, in the English. ``An old man live great while ago, he very good
man, he have three sons. The Great Spirit tell him go make a raft—build wigwam on top; for
he make it rain very much. When this done, Great Spirit say, put in two of all the creatures,
then take sun moon—all the stars, put them in—get in himself with his Equa, (wife) children,
shut door, all dark inside. Then it rain much, hard many days. When they stay there long
time—Great Spirit say, old man go out. So he take, diving animal, sao gy see if find the
earth; so he went, come back, not find any thing. Then he wait few days—send out
mushquash see what he find. When he come back brought some mud in his paw; old man
very glad; he tell mush-quash he wary good, long this world stand be plenty mushquash, no
man ever kill you all. Then few days more he take wary pretty bird send him out see what it
find; that bird no come back; so he sent out one white bird that come back, have grass in he
mouth. So old man know water going down. The Great Spirit say, old man, let sun, moon,
stars go out, old man too. He go out, raft on much big mountain when he see pretty bird he
sent out first, eating dead things—he say, bird you do no right, which me send you no come
back, you must be black, you no pretty bird any more—you always eat bad things. So it was
black.''

There are many things contained in the above that go to support the testimony of the Book of
Mormon, as well as that of the Mosaic history. The Mexican records agree so well with the
words of the book of Ether (found by the people of Limhi, which is contained in the Book of
Mormon) in relation to the confounding of languages, that we insert the following:

(Then is recorded from the Book of Ether the story of the departure of the Jaredites from the
building of the tower of Babel and the story continues):

Here, then, we have two records found upon this continent, that go to support the word of
eternal truth—the Bible; and whilst these records, both of them, sanction the testimony of the
scriptures in regard to the flood, the tower of Babel, and the confusion of languages; the
tradition and hyroglyphics of the Zaltees, the Colhuacans, and the Azteca nations, in regard
to the confusion of languages and their travels to this land, is so like that contained in the
Book of Mormon, that the striking analogy must be seen by every superficial observer.

In regard to the confusion of languages it is said of the above nations, that there were ``fifteen
heads, or chiefs of families, that were permitted to speak the same language.'' The Book of
Mormon, concerning the same event, says: ``And it came to pass that the brother of Jared did
cry unto the Lord; and the Lord had compassion on Jared, therefore he did not confound the
language of Jared''—and it further states that Jared's brother's language was not confounded;
and they then prayed for their families and friends also, and the Lord heard them in their
behalf; and their language was not confounded. These accounts then, precisely agree, one of
which was found in Ontario County, N. Y., and the other in Mexico.

Again, those nations, or families, embodied themselves together and traveled \textit{they knew not
where}, but at length arrived in the country of Aztalan, or the lake country of America. The
Book of Mormon says, that the brother of Jared cried unto the Lord, that he would give them
another land; the Lord heard him, and told him to go to a certain place, ``and there I will meet
thee and go before thee into a land which is choice above all the lands of the earth.'' This it
further speaks is the land of America. The coincidence is so striking that further comment is
unnecessary.

Distorted accounts of Adam and Eve, the Garden of Eden, the fall, flood and confusion of
languages, are found among people all over the earth. They must have come originally from
the same source, before the people were scattered.

\newpage
REFERENCES—CHAPTER TWENTY-TWO

Footnotes

1. Genesis 11:1-9.

2. I. S. Gen. 11:1-6.

3. White, Dr. A. D., \textit{A History of the Ware of Science with Theology}, pp. 96, 262.

4. Draper, Dr. J. W., \textit{Conflict Between Religion and Science}, pp. 185-186.

5. Ether 13:20-21.

6. \textit{Ibid.}, 1:2-6.

7. \textit{Ibid.}, 1:33-41. \textit{See also Mosiah}, Chapters 8, 21, 28.

8. 1 Nephi 13:29, 39, 40.

9. For the origin of language, see ``The Language of Adam,'' \textit{The Way to Perfection}, Chapter
10.
