\chapter{MAN THE OFFSPRING OF GOD}

THE contaminating influence of organic evolution has found its way into every branch of
education. Naturally history cannot escape because history has to deal with man and his
civilization. Historical facts, however, cannot go back of written records, therefore it is
customary for scientists to deal with ancient peoples through the study of geology, biology,
anthropology and archaeology. These studies are young and full of surmises and guesses, and
thus are based on many uncertain deductions. The first great error made by these researchers
is in relation to the age of the earth. Evolution demands great periods of time for its purpose,
and the advocates in the imagination of their hearts have chosen a time far beyond the
knowledge and the power of research to fathom. Sociology has followed in their footsteps
and these studies cannot, and rightfully should not, be considered authentic. There are many
who have accepted these theoretical conclusions and thus we find the blind leading the blind
and all have fallen into the ditch. It is not the purpose here to consider the age of the earth, or
to enter into a discussion of this so-called scientific research. At this point it is sufficient to
discuss the origin of man from the scriptural point of view. It is nevertheless necessary to
refer to these theories in relation to early man for they have been published in nearly every
historical textbook on ancient history since the middle of the nineteenth century. At the
present day in the public schools ancient history begins with ``prehistoric man,'' and places
him many thousands of years before there were any written records. The ``primitive man''
according to these texts was a savage without culture, crude in his appearance, and
worshiping his shadow, the sun, moon, stars, the forces of nature, animals that awed him and
objects or forces which he did not understand. According to the theory he became an idol
worshiper when he had advanced enough in experience to carve himself gods of wood or
stone. Thus he was polytheistic in his worship. Much of this nonsense has been refuted by
better research.

W. Schmidt, professor in the University of Vienna, an expert in his field on Comparative
Religions, in his book, \textit{The Origin and Growth of Religion—Facts and Theories}, maintains
that the farther we go back in time the more it is discovered that the ``primitive'' nations were
monotheistic in their worship. In this book, translated by H. J. Rose, we find the following in
Chapter XVI on \textit{The Primitive High God:}

That the Supreme Being of the primitive culture is really the god of a monotheism, and the
religion which includes him is genuinely monotheistic—this is the position which is most
attacked by a number of authors. To this attack we may reply that there is a sufficient number
of tribes among whom the really monotheistic character of their Supreme Being is clear even
to a cursory examination. This is true of the Supreme Being of most Pygmy tribes, so far as
we know them; also of the Tierra del Fuegians, the primitive Bushmen, the Kurnai, Kulin and
Yuin of South-East Australia, the peoples of the Arctic culture, except the Koryaks, and well-
nigh all of the primitives of North America.

Among other races, the fact of their monotheistic belief has been obscured. This is partly due
to crosses with later forms, partly to differentiation, partly to other causes, all of which can
be discovered only by exact historical analysis. (Page 262.)

In nearly every separate area of the primitive culture the First Father plays an important part,
especially in the initiation ceremonies; originally he and the First Mother were the parents of
the race. Owing to the latter influence of the matrilineal cultures, he develops a lunar
character, is brought into connection with the moon and not uncommonly obscures the
Supreme Being or blends with him. . . . The Supreme Being is everywhere represented
among them as absolutely good, having nothing to do with evil either in conduct or in the
outer world. Evil therefore must have another vehicle or originator; and he, especially in the
mythology of the North American primitives and of those of the Arctic, is opposed to the
Supreme Being; his origin, however, remains darkly mysterious. . . . (Page 263.)

This leads us to a whole series of other races, among whom the Supreme Being is described
as ``shining white'' ``or like fire''; for example, among the North-Western Semang, the
Southern Andamanese, the Wiyot and Patwin of North Central California, the Lenape, an
Algonquin tribe, and the Winnebago, a Sioux tribe influenced by the Algonquin. Among the
Maidu of North Central California we are assured that the whole form of the Supreme Being
shines like the light of the sun, but that his face is always covered and no one has ever seen it,
except the Evil Spirit, who did so once. The Kurnai and Wiradyuri teach that the Supreme
Being is surrounded by an aureole of sunrays. Among the Samoyeds a shaman saw him
blazing with so bright a light that he could not look at him. . . .

The name ``father'' is applied to the Supreme Being in every single area of the primitive
culture when he is addressed or appealed to. It seems, therefore, that we may consider it
primeval and proper to the oldest primitive culture. We find it in the form ``father'' simply,
also in the individual form (``my father'') and the collective (``our father''). So far, this name
has not been discovered among the Central African Pygmies, but it exists among the
Bushmen and the Mountain Dama. It is lacking also among the Andamanese and the
Philippine Negritos, but is found, although not commonly, among the Semang. Among the
Samoyeds we find the formula, ``My Num-father,'' i.e. sky-father. In North Central
California, the name occurs among the Pomo and the Patwin; all three forms of it are widely
distributed among the Algonquins. It is also widely current among the two oldest Tierra del
Fuegian tribes, the Yamana and the Halakwulup who use the form ``my father.'' Among all
the tribes of Southeast Australia it is in common use, in the form ``our father.'' There it is the
oldest name of all, and even the women and children know it; the oldest of the tribes, the
Kurani, have no other name for him. There is no doubt possible that the name ``father'' is
intended in this connection to denote, not physiological paternity (save in cases where the
figures of the Supreme Being and the First Father have coalesced), but an attitude of the
greatest reverence, of tender affection and steadfast trust on the part of man towards God.
(Pages 266-268.)

From all of this it would appear that the knowledge of God had been handed down by
tradition from the days of Adam and Noah. For the antediluvians knew him as shining like
the sun, and the Israelites were frightened when they were given the invitation to meet him in
the Mount.'' (Exodus 19:9-22; 33:9, 10, 11, 21, 22, 23.)

To show this modern trend here are two illustrations:

The first step towards civilization must have been uncertain and slow. No doubt these
beginnings took long periods of time, but we can know little about them, for no people leaves
records that the historian can use until it has advanced a long way from primitive savagery.

To be sure, there are tribes still in primitive stages; and, by comparing them with what can be
gleaned from traditions, customs, words, and early records of our civilization, scholars have
learned something of how our fore-fathers must have lived before Homer and before the
oldest inscriptions upon Egyptian stone. . . .

Still it is well for us to remember that our imposing and varied civilization rests upon this
unrecorded work of prehistoric man through slow, uncounted ages. The development of
language, the invention of the bow, of making fire, of pottery to stand the fire; the
domestication of the dog and cow; the learning to live together, not in droves, but in families
and tribes; the rude beginnings of agriculture; the smelting of metals to replace stone tools;—
these are steps any of which are infinitely more important than the discovery of electricity or
the growth of federal government: but all this, and much more, had become the common
property of many races before history began anywhere. 1

It has taken thousands of years for man to develop from his early state of savagery and
helplessness to the condition in which we now live. In order to understand thoroughly our
present life, it is necessary to study the slow growth of mankind through these past ages. This
story or record of the past life and development of man is the science called history. That part
of the story which is commonly called ``Ancient History'' covers over 4,000 years, extending
from the time when first we know of men through reliable records, down to about 800 years
after the birth of Christ. . . .

The prehistoric ages of man's development stretch back for unknown thousands of years.
During this time, man was slowly learning by bitter experience to light fires, to cook food,
and to tame and make use of some of the gentler animals, such as the dog and the horse.
Then came the knowledge of the value of certain kinds of grain, and the raising of crops.
This long space of time has been divided by historians into four periods, according to the
material used in making hatchets, knives, spearheads, and arrow heads.

1. The Paleolithic or Rough Stone Age.

2. The Neolithic or Polished Stone Age.

3. The Bronze Age.

4. The Iron Age.

The change from savage or barbarous ways of living to what we call ``civilized'' life, did not
take place at one time in all parts of the earth. Some tribes of the Philippine Islands, of
Australia, Africa, and South America are still using tools made of bone and stone. Even now
they are in the state of savagery characteristic of the Stone Age. Yet, at least 6,000 years ago,
the Egyptians were a cultured people, far advanced in the civilization of the Bronze Age and
to a limited extent they even used iron tools. 2

It is the purpose of this chapter to show that these conclusions are incorrectly based and are
founded in imagination and speculation, influenced by the evolutionary theory. If they be
true then the Bible is fiction. If they be true, the Lord never revealed himself to ancient man.
If so, the story of Adam and Eve falls to the ground as a fable. If there was no ``Fall''
consequently there could be no Atonement, for if there had been no Fall there had been no
need of Jesus Christ who came into the world to be its Redeemer from a fallen state, and give
back to mankind that which had been taken away by Adam's transgression. There is no
middle ground! If these writers of history and these teachers of man's gradual progress from
animal forms were correct—and that is definitely implied in all of these so-called historical
conclusions in relation to ``savage primitive man''—then the Christian faith would be false.
On the other hand if Christ is true; if he came into the world as he proclaimed himself, to
restore man to life that he might have it more abundantly, and he is verily ``The Resurrection
and the Life,'' then all of this twaddle and speculation is trash! By overwhelming evidence it
is my duty to show that this modernistic teaching is false and that God is true and his
Beloved Son Jesus Christ accomplished his mission as the Savior of men and Redeemer of
the world!

I write to members of the Church of Jesus Christ of Latter-day Saints. You have a right to
know the truth, to be endowed with an abiding testimony. You have the right to know that
God lives and that Jesus Christ is his Only Begotten Son in the flesh, who came into the
world, preordained to be a sacrifice for a fallen world and give to all those who repent and
obey his commandments eternal life. This truth you have the means of knowing, for the
promise has been made to you that—

If ye continue in my word, then are ye my disciples indeed;

And ye shall know the truth, and the truth shall make you free. 3

Also it is written:

My doctrine is not mine, but his that sent me.

If any man will do his will, he shall know of the doctrine, whether it be of God, or whether I
speak of myself. 4

I assure you that this promise never fails the sincere, humble searcher after truth.

These advocates of evolution say there are four ages of man: 1. The Paleolithic; 2. The
Neolithic; 3. The Bronze Age; 4. The Age of Iron, but these have little meaning in the Gospel
of Jesus Christ, as representing four stages of man from the lowly savage to the cultured man
of our day. All of this is in contradiction of the revelations of God to man. No doubt the first
man on the earth had to learn by experience, line upon line and precept upon precept, but he
was not left to grope in darkness from the use of a bone weapon or a stick with which to
plow. Nor was he under the necessity of finding out the hard way that there were grains and
vegetables, and flocks at his disposal. To cultivate the soil he was commanded; but he had a
heavenly instructor and by revelation many things for his advancement were made known to
him and to his children. One of the first lessons that he learned, by divine instruction, was to
clothe his naked body. He was commanded to till the soil. When sons were born to him they
followed the course of the father. The first man was not a savage, but a gentle god-fearing
man, full of faith and understanding the Gospel for he was taught by our Eternal Father, and
angels sent from his presence. He was commanded to eat the herb of the field, to cultivate it
by the industry of his hands, and the sweat of his face, and prepare the grain to make his
bread. Domestic animals were placed at his command. One of the first sons born to him was
a ``keeper of sheep'' that clothing might be provided, not skins. The other was a tiller of the
soil and raised ``the fruit of the ground.'' They were not idle roaming men, but builders of
communities. They did not have to learn to domesticate ``the dog and cow.''

These authors delight in referring to the fact that there are ``primitive'' people on the face of
the earth today, but their ancestors were not ``primitive'' in the sense in which this word is
used, nor were they in ``savagery.'' The ancient ancestors of the African savage now in the
``stone age,'' in the third generation from the first man, were musicians and made harps and
organs, and were artificers in brass and iron. 5 And mark you, this was in the morning of
man's presence on the earth. Then be it remembered that the ancestors of the savage South
American were men of superior intelligence who loved the Lord and kept his
commandments. They were a very enlightened people. They had prophets among them and
received revelation from the Lord to guide them. It was due to rebellion against the light of
truth and turning from righteousness to wickedness that brought these South Americans to
the pitiful plight in which they now find themselves. They were not polytheistic and
worshipers who had to manufacture their religion and create their gods. The Eternal Father of
us all gave them commandments and revealed to them the plan of salvation and taught them
of the coming of the Son of God, even Jesus Christ, to redeem them from their fallen state.
The knowledge of God was known by the first man placed on this earth. At one time he
dwelt in the presence of God; was taught by him, learned his language and taught his
children to worship the Father in the name of Jesus Christ his Son. The first man and his sons
and daughters were not forced to develop a language, to make fires and come up through
``thousands of years'' from a state of savagery and helplessness, to the condition in which we
now live.'' The Lord has revealed to us much of the history of the first man and his
descendants. It is recorded in the revelations of the Lord that after Adam, the first man, had
been driven from the Garden to till the ground he was commanded to worship the Father and
offer sacrifice to him.

And he gave unto them commandments, that they should worship the Lord their God, and
should offer the firstlings of their flocks, for an offering unto the Lord. And Adam was
obedient unto the commandments of the Lord.

And after many days an angel of the Lord appeared unto Adam, saying: Why dost thou offer
sacrifices unto the Lord? And Adam said unto him: I know not, save the Lord commanded
me.

And then the angel spake, saying: This thing is a similitude of the sacrifice of the Only
Begotten of the Father, which is full of grace and truth.

Wherefore, thou shalt do all that thou doest in the name of the Son, and thou shalt repent and
call upon God in the name of the Son forevermore. 6

The knowledge of God was known among the first inhabitants of this earth. Furthermore,
they were not left to struggle without divine aid in matters of language for it has been also
revealed that Adam and his immediate posterity were taught in the language of God.

And God revealed himself unto Seth, and he rebelled not, but offered an acceptable sacrifice,
like unto his brother Abel. And to him also was born a son, and he called his name Enos.

And then began these men to call upon the name of the Lord, and the Lord blessed them;

And a book of remembrance was kept, in the which was recorded, in the language of Adam,
for it was given unto as many as called upon God to write by the spirit of inspiration;

And by them their children were taught to read and write, having a language which was pure
and undefiled.

Now this same Priesthood, which was in the beginning, shall be in the end of the world also.

Now this prophecy Adam spake, as he was moved upon by the Holy Ghost, and a genealogy
was kept of the children of God. And this was the book of the generations of Adam, saying:
In the day that God created man, in the likeness of God made he him;

In the image of his own body, male and female, created he them, and blessed them, and
called their name Adam, in the day when they were created and became living souls in the
land upon the footstool of God. 7

Here we are informed that man was created in the image of God. This is repeated several
times in the Book of Genesis in speaking of the creation of man. 8 This is the answer to the
evolutionist in relation to the descent of man, and to all religionists as well as scientists who
ridicule the anthropomorphic nature of God. Man was created in the likeness of the body of
God. We call him Father, we are taught that he is literally the Father of the spirits of all men,
and in the spirit they were created, or begotten, sons and daughters unto him. 9 Paul declared
that we are his offspring and that he has ``made of one blood all nations of men for to dwell
on all the face of the earth, and hath determined the times before appointed, and the bounds
of their habitation.'' 10

Members of the Church of Jesus Christ of Latter-day Saints are under obligation to accept the
Bible as the word of God as far as it is translated correctly. We know that there are many
errors in the translations current among the nations. There are many places where different
interpretations are placed upon important passages thus causing confusion and the teaching
of false doctrine. Nevertheless in regard to the relationship of man to God in practically all of
these translations there is agreement. God is our Father. Adam and Eve were created in his
image. Modern revelation, the scriptures which have been restored in the Book of Mormon,
the Book of Moses, of Abraham, the Doctrine and Covenants, all bear witness that man is the
offspring of God, and that man was created in his image. Therefore there is a challenge to all
these theories of men who teach the descent of man through countless ages from lower forms
of life. The revelations of the Lord being true, these theories are false.

We discover in the revelations that man was intelligent in the beginning. It is made known
that Adam is none other than Michael, the Arch-angel, who was sent to this earth to stand at
the head and be a prince over his posterity forever. 11 It was revealed to the Prophet Joseph
Smith that Methuselah was acquainted with the heavenly bodies, and that he had it revealed
unto him that Kolob is the first grand governing star and this was made known to the fathers.
Moreover, it was revealed that the records of the fathers before the flood were handed down
from generation to generation so that mankind could have knowledge of the dealings of God
with the fathers before the flood. These records fell into the hands of Abraham who wrote
about them as follows:

But the records of the fathers, even the patriarchs, concerning the right of Priesthood, the
Lord my God preserved in mine own hands; therefore a knowledge of the beginning of the
creation, and also of the planets, and of the stars, as they were made known unto the fathers,
have I kept even unto this day, and I shall endeavor to write some of these things upon this
record, for the benefit of my posterity that shall come after me. 12

So we learn from this and from many other sources, that the art of writing and the keeping of
records and knowledge of astronomy are very old. In fact this knowledge has come down
from the beginning, even from Adam who was taught to read and write in a perfect language,
for it was the language of God. We are informed that Enoch wrote a record, the history of the
earth and its inhabitants from the beginning down to the end of time. This was done by vision
and revelation, for the Lord opened to his mind and ``showed Enoch all things, even unto the
end of the world; and he saw the day of the righteous, the hour of their redemption; and
received a fulness of joy.'' 13 Moses also was blessed with a similar view and this is
recorded:

And calling upon the name of God, he beheld his glory again, for it was upon him; and he
heard a voice, saying: Blessed art thou, Moses, for I, the Almighty, have chosen thee, and
thou shalt be made stronger than many waters; for they shall obey thy command as if thou
wert God.

And lo, I am with thee, even unto the end of thy days; for thou shalt deliver my people from
bondage, even Israel my chosen.

And it came to pass, as the voice was still speaking, Moses cast his eyes and beheld the earth,
yea, even all of it; and there was not a particle of it which he did not behold, discerning it by
the Spirit of God.

And he beheld also the inhabitants thereof, and there was not a soul which he beheld not; and
he discerned them by the Spirit of God; and their numbers were great, even numberless as the
sand upon the sea shore.

And he beheld many lands; and each land was called earth, and there were inhabitants on the
face thereof.

And it came to pass that Moses called upon God, saying: Tell me, I pray thee, why these
things are so, and by what thou madest them?

And behold, the glory of the Lord was upon Moses, so that Moses stood in the presence of
God, and talked with him face to face. And the Lord God said unto Moses: For mine own
purpose have I made these things. Here is wisdom and it remaineth in me.

And by the word of my power, have I created them, which is mine Only Begotten Son, who
is full of grace and truth.

And worlds without number have I created; and I also created them for mine own purpose;
and by the Son I created them, which is mine Only Begotten.

\textit{And the first man of all men have I called Adam, which is many}

But only an account of this earth, and the inhabitants thereof, give I unto you. For behold,
there are many worlds that have passed away by the word of my power. And there are many
that now stand, and innumerable are they unto man; but all things are numbered unto me, for
they are mine and I know them.

And it came to pass that Moses spake unto the Lord, saying: Be merciful unto thy servant, O
God, and tell me concerning this earth, and the inhabitants thereof, and also the heavens, and
then thy servant will be content.

And the Lord God spake unto Moses, saying: The heavens, they are many, and they cannot
be numbered unto man; but they are numbered unto me, for they are mine.

\textit{And as one earth shall pass away, and the heavens thereof even so shall another come; and
there is no end to my works neither to my words.}

\textit{For behold, this is my work and my glory—to bring to pass the immortality and eternal life of
man.} 14

Here we have the Lord's reason for the building of earths and the peopling of them with the
sons and daughters of God. Moses was favored with a vision in which he saw all of the earth
and its inhabitants. Afterwards the Lord explained to him how his works are carried on and to
which there is no end. The first man on the earth was named Adam, because he was to be the
father of many. This earth on which we stand is only one of many earths, numberless unto
man, but numbered and known unto God. We also learn that earths are formed as habitations
for his offspring, his sons and daughters. His great work and glory being to people these
earths and grant unto his offspring the blessings of immortality or eternal life. In order to
obtain immortality, the spirit and the body must be inseparably connected, the children of
God must pass through a probation such as we are passing through on this earth. They must
have a season of mortality and become familiar with all the vicissitudes of a temporal
probation. Mortal life is a probationary state 15 where we are to be tried, proved, as gold is
tried in the crucible, to see if we will keep all of the commandments of God. If we pass
through this probation successfully we will be entitled to eternal life, if we fail we will be
given immortality. Eternal life is to have the same kind of life, with its glory, that God
possesses; immortality is to have the blessing of living forever, after the resurrection of the
dead, but not with the same glory and blessings which are held in store for those who are just
and true. We learn from these teachings given to Moses that Man is the greatest of all the
creations—for he is the offspring of God. Worlds, or earths, are built as habitations for man
and they too pass through a temporal probationary state corresponding to that of man. Some
earths that have been built and some that are now being created are for habitations for those
who receive immortality; others are to become celestial earths and the eternal abodes of those
who become members of the ``Church of the Firstborn,'' in other words, are true and faithful
to every commandment, covenant and obligation required for exaltation in the Gospel of
Jesus Christ. The Lord in his mercy provides eternal earths for all mankind, each individual
going to his own place according to his works.

The doctrine taught by science, especially by astronomers, is that we live in a changing
universe, that the heavenly bodies had a beginning in which great heat was generated, but
that they are cooling and eventually will become dead cold bodies traveling their uncertain
courses eternally, unless in eons of ages they disintegrate to be made over again. Some
members of the Church, in reading the words of the Lord to Moses have interpreted these
words: ``And as one earth shall pass away, and the heavens thereof even so shall another
come'' with this same understanding. This, however, is not the meaning of the Lord. He does
not create anything to be destroyed. He has said that at no time has he given unto men ``a law
which was temporal; neither any man, nor the children of men; neither Adam, your father,
whom I created.'' 16 The key to this statement is given to the Church in relation to the destiny
of this earth, in the Doctrine and Covenants (88:15-20, 25-26.)

And the spirit and the body are the soul of man.

And the resurrection from the dead is the redemption of the soul.

And the redemption of the soul is through him that quickeneth all things, in whose bosom it
is decreed that the poor and the meek of the earth shall inherit it. [i.e. the earth when
celestialized.]

Therefore, it must needs be sanctified from all unrighteousness, that it may be prepared for
the celestial glory;

For after it hath filled the measure of its creation, it shall be crowned with glory, even with
the presence of God the Father;

That bodies who are of the celestial kingdom may possess it forever and ever; for, for this
intent was it made and created, and for this intent are they sanctified. . .

And again, verily I say unto you, the earth abideth the law of a celestial kingdom, for it filleth
the measure of its creation, and transgresseth not the law—

Wherefore, it shall be sanctified; yea, notwithstanding it shall die, it shall be quickened again,
and shall abide the power by which it is quickened, and the righteous shall inherit it.

We read in the Bible where the prophets speak of this passing away, first we quote Peter who
declares this will come:

The Lord is not slack concerning his promise, as some men count slackness; but is
longsuffering to us-ward, not willing that any should perish, but that all should come to
repentance.

This from Isaiah:

Lift up your eyes to the heavens, and look upon the earth beneath: for the heavens shall
vanish away like smoke, and the earth shall wax old like a garment, and they that dwell
therein shall die in like manner: but my salvation shall be for ever, and my righteousness
shall not be abolished. 17

From the Psalms:

Of old hast thou laid the foundation of the earth: and the heavens are the work of thy hands.

They shall perish, but thou shalt endure: yea, all of them shall wax old like a garment; as a
vesture shalt thou change them, and they shall be changed. 18

Our Lord himself said:

Heaven and earth shall pass away, but my words shall not pass away. 19

The passing away of the heavens has reference to the heavens which surround the earth, not
the sidereal heavens. So we have a key to the meaning of one earth passing away and another
coming. As our earth shall pass away and receive its resurrection, so has it been with other
earths and so will it be with earths yet to come. They will be re-created, made eternal and
find a place perpetually which the Lord has designed for them in the sidereal heavens. These
great orbs that we see in the heavens are not ``passing away.'' Most of them evidently have
attained their state of permanency. They have filled the measure of temporal probation as this
earth is now filling its probation of mortality, and when its work is finished as a temporal
earth it will be exalted. Likewise will others be exalted as countless earths have been and
have attained their state of immortality.

I suppose this is one of the lowest kingdoms that ever the Lord Almighty created, and on that
account is capable of becoming exalted to be one of the highest kingdoms that has ever had
an exaltation in all the eternities. In proportion as it has been reduced so it will be exalted,
with that portion of its inhabitants who in their humiliation have cleaved to righteousness and
acknowledged God in all things.—President Brigham Young, \textit{J. of D.}, 10:175.

This is the testimony of Elder Orson Pratt:

But there is another thing to be considered. Are the wicked to receive this earth as an
inheritance? No; for Jesus did not say, Blessed are the wicked, for they shall inherit the earth;
this promise was made only to the meek. Who are the meek: None but those who receive the
same ordinances the earth has received, and be baptized with fire and with the Holy Ghost, as
this earth will be when Jesus comes to reign upon it a thousand years; and be clothed upon
with the glory of God, as the earth will be; and after they have died as the earth will die, they
will have to be resurrected, as this earth will be resurrected, and then receive their inheritance
upon it.—Orson Pratt, \textit{J. of D.} 1:293.

This from President Charles W. Penrose:

The destiny of this globe is to be fitted as a habitation for the righteous and ``meek'' of the
earth, who will inherit it in their resurrected state. The Lord has revealed that ``the earth
abideth the law of its creation,'' and when it has fully filled the measure thereof, it shall be
crowned with glory, ``even with the presence of the Father;'' that ``although it shall die, it shall
be quickened again'' and shall be inhabited by beings clothed with the celestial glory; that
``for this intent was it made and created.'' (See D. \& C. 88:17-26.) There are many other
particulars concerning the future of this planet, formed by the Eternal as a dwelling place for
this branch of the great family of which he is the head, but on these we will not discourse
further at present.—President Charles W. Penrose, \textit{Liahona}, 6:999.

We now come to this vital point. My fellow believers in the mission of Jesus Christ, in
Joseph Smith and the restoration of the Gospel, as I have said, you are entitled through
faithfully keeping the commandments of the Lord, to individual guidance. It is your right
under these conditions to \textit{know the truth} which makes us free. You cannot be a true member
of the Church and reject Jesus Christ. You cannot be a faithful member and reject the
scriptures—Bible, Book of Mormon, Doctrine and Covenants and Pearl of Great Price—for
these are the standards of our faith. If you accept them you \textit{cannot} accept organic evolution,
for they are diametrically opposed. We will have more to consider in relation to man, his
origin and destiny, in chapters yet to come. Now let us reason together on what is here
presented:

1. Worlds without number have been created.

2. They have been created as habitations for the children of God.

3. The great work and glory of our Father is to bring to pass the immortality and eternal life
of man.

4. Inhabitants of other worlds are begotten sons and daughters of God.

5. When one earth passes away to its exaltation another comes.

6. The making of earths is a glorious work which has been carried on eternally.

This being true, then does it not appear to you that it is a foolish and ridiculous notion that
when God created this earth he had to begin with a speck of protoplasm, and take millions of
years, if not billions, to bring conditions to pass by which his sons and daughters might
obtain bodies made in his image? Why not the shorter route and \textit{transplant} them from
another earth as we are taught in the scriptures? Surely to any reasonable mind, the Lord
would not have to start with an amoeba, pass through the stage of lower fish to higher fish to
reptiles to apes and to man! When we stop to consider how perfect are the workings of God;
how thorough he is and orderly, surely these theories flatten out and are without substance.
Then we have this to think about. According to the revelations to Moses and Abraham, as
given to us in the Pearl of Great Price in clearness, and also stated in the Bible, does it not
seem rather out of harmony for a Latter-day Saint to believe that several billion years ago,
according to our reckoning, there was a council held in heaven at which we shouted for joy
because we were to have the opportunity of coming to the earth to receive bodies that we
might become, through faithfulness, like unto our Father, God? At that time many of the
great and noble spirits were chosen to become rulers. According to the theories of men, if we
believe the revelation of that pre-existence, we had to wait, some say several billions of
years, before that promise could be accomplished.

Also be it remembered, and all who accept the Gospel should remember it, every creature is
eternal. Evolutionists do not believe in the existence of spirit, for man or animal, but we, as
members of the Church do. The Lord has revealed it:

2. Q. What are we to understand by the four beasts, spoken of in the same verse [i. e.
Revelation 4:6.]

\textit{A. They are figurative expressions, used by the Revelator, John, in describing heaven, the
paradise of God, the happiness of man, and of beasts, and of creeping things, and of the
fowls of the air; that which is spiritual being in the likeness of that which is temporal; and
that which is temporal in the likeness of that which is spiritual; the spirit of man in the
likeness of his person, as also the spirit of the beast, and every other creature which God has
created.} (D. \& C. 77:2.)

Here we are informed that God created the beasts as well as man, creeping things, the fowls
of the air and placed in each a spirit in the exact likeness of its body, or more properly,
created every creature in the likeness of its spirit. Therefore they are living entities entitled to
the mercies of Jesus Christ and to receive the resurrection. Then again, the Lord revealed that
every creature shall receive a resurrection:

And again, verily, verily, I say unto you that when the thousand years are ended
[Millennium,] and men again begin to deny their God, then will I spare the earth but for a
little season;

And the end shall come, and the heaven and the earth shall be consumed and pass away, and
there shall be a new heaven and a new earth.

For all old things shall pass away, and all things shall become new, even the heaven and the
earth, and all the fulness thereof, both men and beasts, the fowls of the air, and the fishes of
the sea;

\textit{And not one hair, neither mote, shall be lost, for it is the workmanship of mine hand.} 20

So we learn that all things were created by our Eternal Father, and there is nothing which has
life that he did not create; moreover \textit{every thing shall live again receiving the benefit of the
resurrection.} This proves that \textit{every thing having life}, is endowed with a spirit, and \textit{had a fall.}
In other words became mortal following the transgression of Adam. This also applies to the
earth itself. It was not created a mortal, or temporal earth, but this was acquired when the
curse was placed upon it after Adam's transgression. We have previously referred to the word
of the Lord that—

Wherefore, verily I say unto you that all things unto me are spiritual, and not at any time
have I given unto you a law which was temporal; neither any man, nor the children of men;
neither Adam, your father, whom I created. (D. \& C. 29:34.)

Here again we find the Gospel—the revelations from the Lord—in conflict with geological
and evolutionary teachings. If there were creatures on the earth before Adam, especially men,
where did they obtain their mortal life? They could not be created mortal (temporal) by our
Eternal Father, for that would contradict his own word. If they were created first spiritual and
partook of a fall, who brought that fall upon them? If they were originally made mortal and
subject to death, then they were not entitled to a redemption from death, and since the Lord is
not the author of death, how did they obtain it? Further, how are they entitled to a redemption
and restoration to something that they never had? The whole thing is absurd. All life comes
from God and he did not create it temporally, that was achieved through the violation of a
law. There is no Redeemer other than Jesus Christ for this earth and since Adam could not
have brought death on pre-Adamite life, such life could not obtain the blessings of the
resurrection. Yet the Lord has declared that through the atonement all things partaking of the
fall will be redeemed. So there were no pre-Adamites.

Another thing I wish to say. A man cannot serve God and mammon. Organic evolution is
destructive of faith in God. It is rebellion against him. Those who accept this pernicious
doctrine cannot consistently believe in the fall of Adam. If they do not believe in the fall of
Adam they cannot believe in Jesus Christ, for if Adam had not transgressed the law under
which he was placed on this earth, there would have been no occasion for a redemption. How
could Adam be redeemed from something that never happened. We are taught that had not
Adam partaken of the forbidden fruit all things would have remained in the condition in
which they were before the fall. Here is the passage:

And now, behold, if Adam had not transgressed he would not have fallen, but he would have
remained in the garden of Eden. \textit{And all} things which were created must have remained in
the same state in which they were after they were created; and they must have remained for
ever and had no end. 21

Mr. Charles Darwin was first trained for the ministry. He accepted belief in God. After
making his research and reaching his deductions, he forsook belief in God. 22 Sir Arthur
Keith also was trained for the ministry and accepted a belief in Jesus Christ. After he joined
the ranks of Darwinism, he renounced his faith and rejected the Bible. So it has been with the
many scores of others. They had to renounce their faith in the atonement of Jesus Christ, for
they rejected their faith in the fall of Adam. So it was with Dr. Andrew D. White who
became a bitter opponent of the fall and atonement. Their theories are not compatible with
faith in the God of the scriptures. They consistently have to reject the Sonship of Jesus
Christ. They deny the resurrection of man, and even of Jesus Christ. Many of them say they
honor him as a teacher, a wonderful advocate of truth, but they cannot receive him as the
Messiah, the Savior of the world. Therefore, I appeal to all people everywhere, turn away
from these destructive teachings, for if you tamper with them they will eventually destroy
your soul.

\newpage
REFERENCES—CHAPTER TWELVE

Footnotes

1. West, Willis Mason, \textit{Ancient History}, pp. 1-2.

2. Westermann, William S., \textit{The Story of Ancient Nations}, Introduction.

3. John 8:31-32.

4. \textit{Ibid.}, 7:16-17.

5. Genesis 4:21-22.

6. Moses 5:5-8.

7. \textit{Ibid.}, 6:3-9.

8. Genesis 1:26; 5:1-2.

9. D. \& C. 76:23-24.

10. Acts 17:26-28.

11. D. \& C. 107:53-56.

12. Abraham 1:31.

13. Moses 7:67.

14. \textit{Ibid.}, 1:25-39.

15. 2 Nephi 2:20-21; 9:26-27.

16. D. \& C. 29:34.

17. 2 Peter 3:9-11.

18. Isaiah 51:6.

19. Psalms 102:25-26.

20. Matthew 24:35.

21. D. \& C. 29:22-25.

22. 2 Nephi 2:22.

23. Darwin, Francis, \textit{Life of Charles Darwin}, p. 63.

